# (PART) Inference {-} 

# Sampling {#sampling}

```{r setup_infer, include=FALSE, purl=FALSE}
chap <- 7
lc <- 0
rq <- 0
# **`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**
# **`r paste0("(RQ", chap, ".", (rq <- rq + 1), ")")`**

knitr::opts_chunk$set(
  tidy = FALSE, 
  out.width = '\\textwidth', 
  fig.height = 4,
  warning = FALSE
  )

# This bit of code is a bug fix on asis blocks, which we use to show/not show LC
# solutions, which are written like markdown text. In theory, it shouldn't be
# necessary for knitr versions <=1.11.6, but I've found I still need to for
# everything to knit properly in asis blocks. More info here: 
# https://stackoverflow.com/questions/32944715/conditionally-display-block-of-markdown-text-using-knitr
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})

# This controls which LC solutions to show. Options for solutions_shown: "ALL"
# (to show all solutions), or subsets of c('5-1', '5-2','5-3', '5-4'), including
# the null vector c('') to show no solutions.
solutions_shown <- c('')
show_solutions <- function(section){
  return(solutions_shown == "ALL" | section %in% solutions_shown)
  }
```

In this chapter we will introduce new concepts that will serve as the basis for the remainder of the text:  **sampling**.  We will see that the tools that you learned in the Data Exploration portion of this book (data visualization, "tidy" data format, and data wrangling) will also play an important role here.  As mentioned before, the concepts throughout this text all build into a culmination allowing you to create better stories with data.

We begin with some helpful definitions related to sampling that form the theoretical underpinnings of statical inference, which we will cover in the upcoming "Hypothesis Testing" and "Confidence Intervals", and when we revisit regression. 



### Needed packages {-}

Let's load all the packages needed for this chapter (this assumes you've already installed them). If needed, read Section \@ref(packages) for information on how to install and load R packages.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(broom)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Packages needed internally, but not in text.
library(knitr)
library(tidyverse)
```

---

## Sampling

INSERT PHOTOS OF TUB

```{r, echo=FALSE, message=FALSE, warning=FALSE}
responses <- read_csv("data/sampling_responses.csv") %>% 
  mutate(n=red+white+green) %>% 
  tibble::rownames_to_column(var="ID")
# kable(responses)

responses <- responses %>%
  slice(2:n()) %>%
  mutate(prop_red = red/50) %>% 
  mutate(ID=1:n())
# kable(responses)
```

Here are the 11 sets of counts based on samples of size $n=50$.

| |Group                     | red| white| green|  n|
|:------------|:-------------------------|---:|-----:|-----:|--:|
|1            |Hannah and Siya           |  15|    33|     0| 48|
|2            |Kathleen and Max          |  18|    32|     0| 50|
|3            |Sean, Jack, and CJ        |  18|    32|     0| 50|
|4            |Orianna and Judy          |  22|    28|     0| 50|
|5            |James and Jacob           |  21|    29|     0| 50|
|6            |Hannah and Siya           |  16|    34|     0| 50|
|7            |Niko, Sophie, and Caitlin |  14|    36|     0| 50|
|8            |Niko, Sophie, and Caitlin |  19|    31|     0| 50|
|9            |Aleja and Ray             |  20|    30|     0| 50|
|10           |Yaw and Drew              |  16|    34|     0| 50|
|11           |Yaw and Drew              |  21|    29|     0| 50|

The first row only had $n=48$, so let's scrap this one (sorry Hannah and Siya)!
So we have 10 samples of size $n=50$. Let's look at the sample proportion of the $n=50$
balls that were red.

| |Group                     | red| white| green|  n| prop_red|
|--:|:-------------------------|---:|-----:|-----:|--:|--------:|
|  1|Kathleen and Max          |  18|    32|     0| 50|     0.36|
|  2|Sean, Jack, and CJ        |  18|    32|     0| 50|     0.36|
|  3|Orianna and Judy          |  22|    28|     0| 50|     0.44|
|  4|James and Jacob           |  21|    29|     0| 50|     0.42|
|  5|Hannah and Siya           |  16|    34|     0| 50|     0.32|
|  6|Niko, Sophie, and Caitlin |  14|    36|     0| 50|     0.28|
|  7|Niko, Sophie, and Caitlin |  19|    31|     0| 50|     0.38|
|  8|Aleja and Ray             |  20|    30|     0| 50|     0.40|
|  9|Yaw and Drew              |  16|    34|     0| 50|     0.32|
| 10|Yaw and Drew              |  21|    29|     0| 50|     0.42|


Ehhhhh, this is hard to digest. Let's visualize:

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width = 1.5*9/2}
p <- 700/2400
sampling_distribution <- ggplot(responses, aes(x=prop_red)) +
  geom_histogram(binwidth = 0.05) +
  labs(x="Proportion red in sample of size n=50", y="Number of samples",
       title="Proportion red in 10 samples of size n=50")
sampling_distribution
```

What was the true proportion? I asked Professor X, who replied "From what I recall it was 2400 balls total, 1600 white, 100 green, so 700 red? But that's assuming we didn't lose any." In other words, the true proportion $p=\frac{700}{2400} = 0.2916$. Let's plot this.



### Definitions

1. Observational unit
1. Population
1. Population parameter
1. Sample
1. Sampling
1. Representative sampling
1. Generalizability
1. Bias
1. Statistic
1. Point estimates/sample statistics
1. Census

Moral:

* We use sampling when it is too expensive/tiring to conduct a *census*. Thus, we
need to guess at the value of the population parameter using a point estimate
based on a sample of size $n$.
* When sampling is done at random, or in other words in an unbiased fashion, the
sample is representative of the population
* When the sample is representative of the population, then any results based on 
the sample generalize to the entire population. 
* When any results based on the sample generalize to the entire population, the
point estimate/sample statistic is a "good" guess of the population parameter.


### Representative sampling

For the following examples:
  
1. Identify the population of interest
1. Comment on the representativeness of the sampling procedure


1. You go to the homepages of CNN and Fox News, and read 
1. The Royal Air Force wants to study how resistant their airplanes are to bullets. They study the bullet holes on all the airplanes on the tarmac after an air battle against the Luftwaffe (German Air Force).
1. You want to know the average income of graduates of your academic institution in the last 10 years.  So you get the records of 10 randomly chosen alumni.  They all answer and you take the average.
1. Imagine it's 1993 i.e. almost all households have landlines.  You want to know the average number of people in each household in your hometown.  You randomly pick out 500 phone numbers from the phone book and conduct a phone survey.
1. You want to know the prevalence of illegal downloading of TV shows among students at an academic institution.  You get the emails of 100 randomly chosen students and ask them "How many times did you download a pirated TV show last week?"







