# Confidence Intervals {#ci}
    
    
```{r setup_ci, include=FALSE}
chap <- 8
lc <- 0
rq <- 0
# **`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**
# **`r paste0("(RQ", chap, ".", (rq <- rq + 1), ")")`**
knitr::opts_chunk$set(tidy = FALSE, out.width='\\textwidth')#, fig.align = "center")
```

## Needed packages

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(okcupiddata)
library(mosaic)
library(knitr)
library(nycflights13)
```


**Definition: Confidence Interval**

A *confidence interval* gives a range of plausible values for a parameter.  It depends on a specified _confidence level_ with higher confidence levels corresponding to wider confidence intervals and lower confidence levels corresponding to narrower confidence intervals.  Common confidence levels include 90%, 95%, and 99%.


<!--
Standard error

Interpretation: was constructed using the sample data by a procedure that is 95% reliable. 
Lift/steal the OI diagram of horizontal lines, with those not capturing pop parameter in red
Sample heights from profiles in okcupiddata (Section 3.c) of outline)
There is only one CI 
  - Build a diagram
Fishing with a spear/net Analogy: Confidence intervals
Confidence interval MEGA LC: s/sqrt(n) form of SE
  - Population distribution very right skewed
  - Population distribution symmetric
  - Sample size small from very left skewed population
-->