# Inference for Regression {#inference-for-regression}

```{r setup_inference_regression, include=FALSE}
chap <- 11
lc <- 0
rq <- 0
# **`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**
# **`r paste0("(RQ", chap, ".", (rq <- rq + 1), ")")`**

knitr::opts_chunk$set(
  tidy = FALSE, 
  out.width = '\\textwidth', 
  fig.height = 4,
  warning = FALSE
  )

options(scipen = 99, digits = 3)

# Set random number generator see value for replicable pseudorandomness. Why 76?
# https://www.youtube.com/watch?v=xjJ7FheCkCU
set.seed(76)
```

### Needed packages {-}

Let's load all the packages needed for this chapter (this assumes you've already installed them). Read Section \@ref(packages) for information on how to install and load R packages.

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(moderndive)
library(infer)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Packages needed internally, but not in text.
library(knitr)
library(tidyr)
library(kableExtra)
library(patchwork)
# Add back in if needed later
#library(gapminder)
#library(ISLR)
```



***

## Regression refresher

Let's first head back to reminding ourselves of the University of Texas student evaluations analysis in Section \@ref(model1). Let's begin by looking at the `evals_ch6` data frame that selected a subset of variables from the `evals` data frame in the `moderndive` package.

```{r}
evals_ch6 <- evals %>%
  select(ID, score, bty_avg, age)
glimpse(evals_ch6)
```

In Chapter \@ref(regression), we investigated the relationship between student evaluation ratings (`score`) and the beauty rating (`bty_avg`). We saw there that there was a weakly positive correlation between the two variables as shown in the plot below. Note the use of the regression line as well in blue that attempts to summarize this linear relationship between the two variables.

```{r regline, fig.cap="Relationship with regression line"}
ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Relationship between teaching and beauty scores") +  
  geom_smooth(method = "lm", se = FALSE)
```

Looking at this plots again, you might be asking "But is that line really that positive in its slope?" It does increase from left to right as the `bty_avg` variable increases, but by how much? To get to this information, we can use the `lm()` function to determine the necessary values that comprise the defining regression equation. Recall the use of the `get_regression_table()` function in the `moderndive` package to better summarize the results of this regression fit as well.

```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch6)
# Get regression table:
get_regression_table(score_model)
```

```{r echo=FALSE}
slope_row <- get_regression_table(score_model) %>% 
  filter(term == "bty_avg")
b1 <- slope_row %>% pull(estimate)
se1 <- slope_row %>% pull(std_error)
t1 <- slope_row %>% pull(statistic)
lower1 <- slope_row %>% pull(lower_ci)
upper1 <- slope_row %>% pull(upper_ci)
```

## Interpretation of regression table results {#regression-interp}

In Chapter \@ref(regression) we focused on the left-most column of this resulting summary of `score_model` using `get_regression_table()`. Remember the interpretation here of the `r b1` value:

> For every increase of one unit in beauty rating, there is an associated increase, on average, of `r b1` units of evaluation score.

Thus, the slope value quantifies the relationship between the `y` variable of `score` and the `x` variable of `bty_avg` as a single statistic/point estimate from our sample of student evaluations.

We also discussed the intercept value of $b_0$ = `r get_regression_table(score_model) %>% filter(term == "intercept") %>% pull(estimate)` and its lack of practical interpretation since the range of possible beauty scores does not include 0. Let's now shift to focus on the remaining columns of `std_error`, `statistic`, `p_value`, and the two columns related to confidence intervals (`lower_ci` and `upper_ci`) that were not discussed in Chapter \@ref(regression). Given we've just mentioned the lack of interpretation for the intercept value, we'll focus on the second row of the table corresponding to the slope of the regression equation.

### Standard error

Recall the definition of **standard error** given in Chapter \@ref(sampling):

> The standard error is the standard deviation of any point estimate from a sampling scenario.

So what does this mean in this scenario focusing on the sample slope statistic of $b_1$ = `r b1`. This value is just one possible statistic from one sample of students evaluations at the University of Texas. If a different sample was collected for the two variables of `bty_avg` and `score`, we may see different results. If this was done over and over again, we could quantify just how much deviation one would expect with a standard deviation of the sampling distribution for the point estimate $b_1$. So in our case, we can expect about `r se1` units of deviation on the `bty_avg` slope variable. Recall that `estimate` and `std_error` play a key role in helping us to make an educated guess about possible values for the unknown population slope `\beta_1` relating **all** University of Texas student evaluation scores in the year of the study to the beauty score of all faculty.

### Statistic

The `statistic` column in the regression table corresponds to a standardized statistic. This is similar to what was done in Subsection \@ref(theory-hypo) when a $t$-score was computed. To determine the `statistic` entry for slope, we can divide the `estimate` value by its corresponding `std_error` value to create the $t$-score. (Note that there will be a bit of rounding error here if the calculation is done with these rounded table values.) This entry is useful in determining the statistical significance of the $b_1$ value by determining how extreme it is after standardization in the corresponding `p_value` that follows.

### p-value

The `p_value` entry here corresponds to how extreme the entry of `r t1` is when compared to the appropriate $t$ distribution. With the `p_value` being 0 with rounding, we can conclude that the sample slope of `r b1` is statistically greater than 0. In other words, we have evidence that the relationship between beauty score and student evaluation score is positive.

An extra caveat here is that this analysis of statistical significance via the hypothesis test using the $t$-statistic is only valid if some conditions are met. We'll see more about this in Section \@ref(regression-conditions).

### Confidence interval

The final rightmost entries in the table signify a range of possible values for the population slope `\beta_1` referring to the linear relationship between beauty score and student evaluation score in the population. Given that our hypothesis test above told us we had evidence of a positive slope, we should be happy to see the entries for `lower_ci` and `upper_ci` are both greater than 0. 

To obtain these two values, we can use an approach similar to what was done in Subsection \@ref(infer-se). Remember here we are using the theory-based method of a standardized statistic instead of a simulation-based method. We'll compare these two results in Section \@ref(infer-regression). Recall that if we want to compute a 95% confidence interval around our point estimate, we can use the following formula:

$$point\_estimate \pm 2 * standard\_error$$

This confidence interval also is only valid if the conditions discussed in Section \@ref(regression-conditions) are met. Doing the multiplication and subtraction prescribed in the formula by plugging in the other relevant entries from the slope row gives us the resulting confidence interval of $(`r lower1`, `r upper1`)$. 

## Conditions for inference with regression {#regression-conditions}

In order for inference to be valid with linear regression, we must check for four conditions to be met. Note the first four letters of the conditions as highlighted in bold below. This can serve as a nice reminder of what to check whenever linear regression is performed.

- **L**inearity of relationship between variables
- **I**ndependence of residuals
- **N**ormality of residuals
- **E**quality of variance

These conditions can mostly be checked via appropriate data visualizations. There are statistical tests that can also be done to further check for this, but here we will focus on the simpler style of evaluating plots.

### Linearity of relationship

We've already checked for this with our original scatterplot in Figure \@ref(fig:regline). While the plot didn't show a clear example of a linear relationship in that there was some clear deviation from a regression line fit throughout the plot, it also didn't show a different kind of non-linear relationship either. Let's dig further into the other conditions to see how they stack up.

### Independence of residuals

The next three conditions all refer to investigation of the residuals.  Investigating any such patterns is known as *residual analysis*. Recall that residuals can be thought of as the error or the "lack-of-fit" between the observed value $y$ and the fitted value $\widehat{y}$ on the blue regression line in Figure \@ref(fig:regline). Ideally when we fit a regression model, we'd like there to be *no systematic pattern* to these residuals. We'll be more specific as to what we mean by *no systematic pattern* when we see Figure \@ref(fig:regline), but let's keep this notion imprecise for now. 

We'll perform our residual analysis in two ways:

1. Creating a scatterplot with the residuals on the $y$-axis and the original explanatory variable $x$ on the $x$-axis.
1. Creating a histogram of the residuals, thereby showing the *distribution* of the residuals.

First, recall in Figure \@ref(fig:numxplot4) in Chapter \@ref(regression) we created a scatterplot. This is again shown below and has the following characteristics:

* on the vertical axis we had the teaching score $y$ and
* on the horizontal axis we had the beauty score $x$.

```{r numxplot-ch11, echo=FALSE, warning=FALSE, fig.cap="Example of observed value, fitted value, and residual"}
best_fit_plot <- ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Relationship of teaching and beauty scores") +
  geom_smooth(method = "lm", se = FALSE)
best_fit_plot
```


Instead, in Figure \@ref(fig:numxplot6) below, let's create a scatterplot where

* On the vertical axis we have the residual $y-\widehat{y}$ instead
* On the horizontal axis we have the beauty score $x$ as before:

```{r, eval=TRUE, echo=TRUE}
# Get data
evals_ch6 <- evals %>%
  select(score, bty_avg, age)
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch6)
# Get regression table:
get_regression_table(score_model)
# Get regression points
regression_points <- get_regression_points(score_model)
```

```{r, echo=FALSE}
index <- which(evals_ch6$bty_avg == 7.333 & evals_ch6$score == 4.9)
target_point <- score_model %>% 
  get_regression_points() %>% 
  slice(index)
x <- target_point$bty_avg
y <- target_point$score
y_hat <- target_point$score_hat
resid <- target_point$residual
```


```{r, eval=FALSE, echo=TRUE}
ggplot(regression_points, aes(x = bty_avg, y = residual)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1)
```
```{r numxplot6, echo=FALSE, warning=FALSE, fig.cap="Plot of residuals over beauty score"}
ggplot(regression_points, aes(x = bty_avg, y = residual)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1) +
  annotate("point", x = x, y = resid, col = "red", size = 3) +
  annotate("point", x = x, y = 0, col = "red", shape = 15, size = 3) +
  annotate("segment", x = x, xend = x, y = resid, yend = 0, color = "blue",
           arrow = arrow(type = "closed", length = unit(0.02, "npc")))
```

You can think of Figure \@ref(fig:numxplot6) as Figure \@ref(fig:numxplot5) but with the blue line flattened out to $y=0$. Does it seem like there is *no systematic pattern* to the residuals? This question is rather qualitative and subjective in nature, thus different people may respond with different answers to the above question. However, it can be argued that there isn't a *drastic* pattern in the residuals.

We do have extra information here that may lead us to question the independence of residuals further: The same professors exist multiple times in the data set in that multiple rows of the data set correspond to some of the same professors. Different courses that the professors taught were included in the analyses. Therefore, the residual in one observation will have a dependent relationship with another row breaking the independence assumption of the residuals.

**Note:** As is often the case with regression analysis in the real-world, it can be difficult for the assumptions to be met exactly. We as authors advocate for clarity of interpretation and letting the stakeholders of analyses know about any shortcomings of a model as needed. Here as a preliminary analysis teaching the topics, we will continue on with testing the other conditions. A more valid and precise analysis would incorporate these dependencies between rows and instructors. This topic can get to the advanced stages quickly, so we have decided to omit any further discussion here.


### Normality of residuals

Let's now get a little more precise in our definition of *no systematic pattern* in the residuals. Ideally, the residuals should behave *randomly*. In addition,

1. the residuals should be on average 0. In other words, sometimes the regression model will make a positive error in that $y - \widehat{y} > 0$, sometimes the regression model will make a negative error in that $y - \widehat{y} < 0$, but *on average* the error is 0. 
1. the residuals should have the characteristic normal shape.

The simplest way to check for this condition of the normality of the residuals is to look at a histogram.

```{r, eval=FALSE, echo=TRUE}
ggplot(regression_points, aes(x = residual)) +
  geom_histogram(binwidth = 0.25, color = "white") +
  labs(x = "Residual")
```
```{r model1residualshist, echo=FALSE, warning=FALSE, fig.cap="Histogram of residuals"}
ggplot(regression_points, aes(x = residual)) +
  geom_histogram(binwidth = 0.25, color = "white") +
  labs(x = "Residual")
```

This histogram seems to indicate that we have more positive residuals than negative. Since the residual $y-\widehat{y}$ is positive when $y > \widehat{y}$, it seems our fitted teaching score from the regression model tends to *underestimate* the true teaching score. This histogram has a slight *left-skew* in that there is a long tail on the left. Another way to say this is this data exhibits a *negative skew*. Is this a problem? Again, there is a certain amount of subjectivity in the response. In the authors' opinion, while there is a slight skew/pattern to the residuals, it isn't a large concern. On the other hand, others might disagree with our assessment. Here are examples of an ideal and less than ideal pattern to the residuals when viewed in a histogram:

```{r numxplot9, echo=FALSE, warning=FALSE, fig.cap="Examples of ideal and less than ideal residual patterns"}
resid_ex <- evals_ch6
resid_ex$`Ideal` <- rnorm(nrow(resid_ex), 0, sd = sd(regression_points$residual))
resid_ex$`Less than ideal` <-
  rnorm(nrow(resid_ex), 0, sd = sd(regression_points$residual))^2
resid_ex$`Less than ideal` <- resid_ex$`Less than ideal` - mean(resid_ex$`Less than ideal` )

resid_ex <- resid_ex %>%
  select(bty_avg, `Ideal`, `Less than ideal`) %>%
  gather(type, eps, -bty_avg)

ggplot(resid_ex, aes(x = eps)) +
  geom_histogram(binwidth = 0.25, color = "white") +
  labs(x = "Residual") +
  facet_wrap( ~ type, scales = "free")
```

### Equality of variance

Lastly, to test for the **E** part of **LINE** in our conditions check, the value and spread of the residuals should not depend on the value of $x$. 

In Figure \@ref(fig:numxplot7) below, we display some hypothetical examples where there are *drastic* patterns to the residuals. In Example 1, the value of the residual seems to depend on $x$: the residuals tend to be positive for small and large values of $x$ in this range, whereas values of $x$ more in the middle tend to have negative residuals. In Example 2, while the residuals seem to be on average 0 for each value of $x$, the spread of the residuals varies for different values of $x$; this situation is known as *heteroskedasticity*. 

```{r numxplot7, echo=FALSE, warning=FALSE, fig.cap="Examples of less than ideal residual patterns"}
resid_ex <- evals_ch6
resid_ex$ex_1 <- ((evals_ch6$bty_avg - 5) ^ 2 - 6 + rnorm(nrow(evals_ch6), 0, 0.5)) * 0.4
resid_ex$ex_2 <- (rnorm(nrow(evals_ch6), 0, 0.075 * evals_ch6$bty_avg ^ 2)) * 0.4
  
resid_ex <- resid_ex %>%
  select(bty_avg, ex_1, ex_2) %>%
  gather(type, eps, -bty_avg) %>% 
  mutate(type = ifelse(type == "ex_1", "Example 1", "Example 2"))

ggplot(resid_ex, aes(x = bty_avg, y = eps)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1) +
  facet_wrap(~type)
```

Do we see this kind of behavior in our student evaluation versus beauty score example? Let's look at the plot again in Figure \@ref(fig:numxplot6) removing the highlighting of a single residual.

```{r numxplot6-again, echo=FALSE, warning=FALSE, fig.cap="Plot of residuals over beauty score"}
ggplot(regression_points, aes(x = bty_avg, y = residual)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1)
```

As we look at this plot, we don't see the noticeable patterns shown in Figure \@ref(fig:numxplot7). It doesn't have the fan-like pattern of Example 2 or the bowl-like pattern of Example 1. The residuals are pretty much scattered in similar ways as we look from small beauty score values on the left to larger beauty score values on the right.

```{block, type='learncheck', purl=FALSE}
**_Learning check_**
```

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Continuing with our regression using `age` as the explanatory variable and teaching `score` as the outcome variable, use the `get_regression_points()` function to get the observed values, fitted values, and residuals for all `r nrow(evals_ch6)` instructors. Perform a residual analysis and look for any systematic patterns in the residuals. Ideally, there should be little to no pattern.

```{block, type='learncheck', purl=FALSE}
```

### Summary

The conditions for inference in regression problems are a key part of regression analysis that are of vital importance to the processes of confidence intervals and hypothesis tests. We also saw that there is a level of subjectivity in evaluating plots so you may be asking "What do I do when conditions aren't clearly met?". We as authors are of the opinion that discussion of these shortcomings goes a long way in helping others to understand your analyses. You won't always have ideal results but regression is a powerful technique that can do some wondrous things with data. But if conditions are violated badly, regression will just as much put you in the "garbage in, garbage out" situation. We recommend being vigilant about checking these conditions and you'll get better with practice on how to discuss the results of the analyses.

## Simulation-based inference for regression {#infer-regression}

### Formula for the standard error of the slope statistic

The discussion of the regression table results in Section \@ref(regression-interp) were done using the traditional theory-based analyses and a $t$-statistic. There are formulas to compute the standard error as well as given in the table there. Let's review that table again here:

```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch6)
# Get regression table:
get_regression_table(score_model)
```

Since we didn't do any bootstrapping or simulation-based procedures to get the value of `r se1` for the standard error, how was this number computed? What did R do behind the scenes? Well, there's a formula for the standard error. As we've seen with other standard error formulas $n$ and the square root function play a role. Memorizing this formula isn't the important part here. Seeing how the formula compares to the one given in the denominator of the $t$-statistic formula in Section \@ref(theory-hypo) is helpful here. From the theory-based perspective, the standard error of the slope statistic is denoted as $s_{b_1}$ and is given by

$$s_{b_1} = \dfrac{\dfrac{s_y}{s_x} \cdot \sqrt{1-r^2}}{\sqrt{n-2}}$$

<!-- Really like this breakdown: https://stats.stackexchange.com/questions/342632/how-to-understand-se-of-regression-slope-equation -->

As with many formulas in statistics, there's a lot going on here so let's first break down what each symbol represents:

- $s_y$ is the sample standard deviation for the response variable `score`
- $s_x$ is the sample standard deviation for the explanatory variable `bty_avg`
- $r$ is the sample correlation coefficient between `score` and `bty_avg`. This was computed as `r cor(evals_ch6$score, evals_ch6$bty_avg)` in Chapter \@ref(regression).
- $n$ is the number of pairs of points in the `evals_ch6` data frame (463).

To put into words, the standard error depends on the relationship between how the response variable varies and the explanatory variable varies in the $s_y / s_x$ term. Next it looks into the relationship of how the two variables relate to each other in the $\sqrt{1-r^2}$ term. Lastly, the standard error depends on the "sample size", which here corresponds to how many observations are in the data set.

### Hypothesis testing using infer

We can also use the concept of permuting to determine the standard error of our null distribution and conduct a hypothesis test for a population slope. Let's see how the results above using the theoretical procedures on the student evaluations of faculty data compares to if we used the `infer` package procedures instead.  

We'll begin in the basic regression setting to test to see if we have evidence that a statistically significant *positive* relationship exists between teaching and beauty scores for the University of Texas professors. As we did in Chapter \@ref(regression), teaching `score` will act as our outcome/response variable and `bty_avg` will be our explanatory variable. We will set up this hypothesis testing process as we have each before via the "There is Only One Test" diagram in Figure \@ref(fig:htdowney) using the `infer` package.

### Data

Our data is stored in `evals` and we are focused on the measurements of the `score` and `bty_avg` variables there. Note that we don't choose a subset of variables here since we will `specify()` the variables of interest using `infer`.

```{r}
evals %>% 
  specify(score ~ bty_avg)
```

### Test statistic $\delta$

Our test statistic here is the sample slope coefficient that we denote with $b_1$.

### Observed effect $\delta^*$

We can use the `specify() %>% calculate()` shortcut here to determine the slope value seen in our observed data:

```{r}
slope_obs <- evals %>% 
  specify(score ~ bty_avg) %>% 
  calculate(stat = "slope")
```

The calculated slope value from our observed sample is $b_1 = `r pull(slope_obs)`$.

### Model of $H_0$

We are looking to see if a positive relationship exists so $H_A: \beta_1 > 0$.  Our null hypothesis is always in terms of equality so we have $H_0: \beta_1 = 0$. In other words, when we assume the null hypothesis is true, we are assuming there is NOT a linear relationship between teaching and beauty scores for University of Texas professors.

### Simulated data

Now to simulate the null hypothesis being true and recreate how our sample was created, we need to think about what it means for $\beta_1$ to be zero. If $\beta_1 = 0$, we said earlier that there is no relationship between the teaching and beauty scores. If there is no relationship, then any one of the teaching score values could have just as likely occurred with any of the other beauty score values instead of the one that it actually did fall with. We, therefore, have another example of permuting in our simulating of data under the null hypothesis.

**Tactile simulation**

We could use a deck of `r nrow(evals) * 2` note cards to create a tactile simulation of this permuting process. We would write the `r nrow(evals)` different values of beauty scores on each of the `r nrow(evals)` cards, one per card. We would then do the same thing for the `r nrow(evals)` teaching scores putting them on one per card.

Next, we would lay out each of the `r nrow(evals)` beauty score cards and we would shuffle the teaching score deck. Then, after shuffling the deck well, we would disperse the cards one per each one of the beauty score cards. We would then enter these new values in for teaching score and compute a sample slope based on this permuting. We could repeat this process many times, keeping track of our sample slope after each shuffle.

### Distribution of $\delta$ under $H_0$

We can build our null distribution in much the same way we did in Chapter \@ref(hypothesis-testing) using the `generate()` and `calculate()` functions. Note also the addition of the `hypothesize()` function, which lets `generate()` know to perform the permuting instead of bootstrapping. 

```{r eval=FALSE}
null_slope_distn <- evals %>% 
  specify(score ~ bty_avg) %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 10000) %>% 
  calculate(stat = "slope")
```

```{r echo=FALSE}
if(!file.exists("rds/null_slope_distn.rds")){
  null_slope_distn <- evals %>% 
    specify(score ~ bty_avg) %>%
    hypothesize(null = "independence") %>% 
    generate(reps = 10000) %>% 
    calculate(stat = "slope")
   saveRDS(object = null_slope_distn, 
           "rds/null_slope_distn.rds")
} else {
   null_slope_distn <- readRDS("rds/null_slope_distn.rds")
}
```

```{r}
visualize(null_slope_distn) + 
  shade_p_value(obs_stat = slope_obs, direction = "greater")
```

In viewing the distribution above with shading to the right of our observed slope value of `r pull(slope_obs)`, we can see that we expect the p-value to be quite small. Let's calculate it next using a similar syntax to what was done with `shade_p_value()`.

### The p-value

```{r fig.cap="Shaded histogram to show p-value"}
null_slope_distn %>% 
  get_pvalue(obs_stat = slope_obs, direction = "greater")
```

Since `r pull(slope_obs)` falls far to the right of this plot beyond where any of the histogram bins have data, we can say that we have a $p$-value of 0.  We, thus, have evidence to reject the null hypothesis in support of there being a positive association between the beauty score and teaching score of University of Texas faculty members.

Recall that we came to a similar $p$-value and conclusion when looking at the theory-based approach given in the `get_regression_table(score_model)` table:

```{r echo=FALSE}
get_regression_table(score_model)
```

When the conditions are met and the null distribution has the characteristic bell shape, we are likely to see similar results between the simulation-based and theoretical results. In the next subsection, we'll compare the confidence interval results using bootstrapping and the theoretical results given in the regression table.

```{block lc9-5, type='learncheck', purl=FALSE}
**_Learning check_**
```

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Repeat the inference above but this time for the correlation coefficient instead of the slope. Note the implementation of `stat = "correlation"` in the `calculate()` function of the `infer` package.

```{block, type='learncheck', purl=FALSE}
```

### Bootstrapping for the regression slope

With the $p$-value calculated as 0 in the hypothesis test above, we can next determine just how strong of a positive slope value we might expect between the variables of teaching `score` and beauty score (`bty_avg`) for University of Texas faculty using bootstrapping. Recall the `infer` pipeline above to compute the null distribution. This assumes the null hypothesis is true that there is no relationship between teaching score and beauty score using the `hypothesize()` function.

```{r eval=FALSE}
null_slope_distn <- evals %>% 
  specify(score ~ bty_avg) %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 10000, type = "permute") %>% 
  calculate(stat = "slope")
```

To further reinforce the process being done in the pipeline, we've added the `type` argument to `generate()`. This is automatically added based on the entries for `specify()` and `hypothesize()` but it provides a useful way to check to make sure `generate()` is created the samples in the desired way. In this case, we `permute`d the values of one variable across the values of the other 10,000 times and `calculate`d a `"slope"` coefficient for each of these 10,000 `generate`d samples.

If instead we'd like to get a range of plausible values for the true slope value, we can use the process of bootstrapping:

```{r eval=FALSE}
bootstrap_slope_distn <- evals %>% 
  specify(score ~ bty_avg) %>%
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "slope")
```

```{r echo=FALSE}
if(!file.exists("rds/bootstrap_slope_distn.rds")){
  bootstrap_slope_distn <- evals %>% 
    specify(score ~ bty_avg) %>%
    generate(reps = 10000, type = "bootstrap") %>% 
    calculate(stat = "slope")
  saveRDS(object = bootstrap_slope_distn, 
           "rds/bootstrap_slope_distn.rds")
} else {
  bootstrap_slope_distn <- readRDS("rds/bootstrap_slope_distn.rds")
}
```

Here, bootstrapping is done on the row level as was first done in Subsection \@ref(ci-build). Thus, the orignal pairs of points are linked together and may be repeated with each resample.

```{r}
bootstrap_slope_distn %>% visualize()
```

Next we can use the `get_ci()` function to determine the confidence interval. Let's do this in two different ways obtaining 99% confidence intervals. Remember that these denote a range of plausible values for an unknown true population slope parameter regressing teaching score on beauty score.

```{r}
percentile_slope_ci <- bootstrap_slope_distn %>% 
  get_ci(level = 0.99, type = "percentile")
percentile_slope_ci
```

```{r}
se_slope_ci <- bootstrap_slope_distn %>% 
  get_ci(level = 0.99, type = "se", point_estimate = slope_obs)
se_slope_ci
```

With the bootstrap distribution being close to symmetric, it makes sense that the two resulting confidence intervals are similar. Note that both are a bit wider than the $(`r lower1`, `r upper1`)$ interval that theory would have predicted.

<!-- Albert will add content here.

## Conclusion


### Relating regression to other methods

To conclude this chapter, we'll be investigating how regression relates to two different statistical techniques. One of them was covered already in this book, the difference in sample means, and the other is new to the text but is related, ANOVA. We'll see how both can be represented in the regression framework. The hope is that this closing section helps you to tie together many of the concepts you've seen in the Data Modeling and Statistical Inference parts of this book.

#### Two sample difference in means

#### ANOVA

-->

### Additional resources

An R script file of all R code used in this chapter is available [here](scripts/11-inference-for-regression.R).

### What's to come

You've now concluded the last major part of the book on "Statistical Inference via infer." The closing chapter concludes with a case study on house prices in Seattle, Washington in the US. You'll see there how the principles in this book can apply to help you be a great storyteller with data!
