# Tips and Tricks {#appendixC}

## Needed packages {-}

Let's load all the packages needed for this chapter (this assumes you've already installed them). Recall from our discussion in Section \@ref(tidyverse-package) that loading the `tidyverse` package by running `library(tidyverse)` loads the following commonly used data science packages all at once:

* `ggplot2` for data visualization
* `dplyr` for data wrangling
* `tidyr` for converting data to "tidy" format
* `readr` for importing spreadsheet data into R
* As well as the more advanced `purrr`, `tibble`, `stringr`, and `forcats` packages

If needed, read Section \@ref(packages) for information on how to install and load R packages. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
```

## Data wrangling tips and tricks





***



## Interactive graphics

### Interactive linegraphs

Another useful tool for viewing linegraphs such as this is the `dygraph` function in the `dygraphs` package in combination with the `dyRangeSelector` function.  This allows us to zoom in on a selected range and get an interactive plot for us to work with:

```{r warning=FALSE, out.width="100%"}
library(dygraphs)
flights_day <- mutate(flights, date = as.Date(time_hour))
flights_summarized <- flights_day %>% 
  group_by(date) %>%
  summarize(median_arr_delay = median(arr_delay, na.rm = TRUE))
rownames(flights_summarized) <- flights_summarized$date
flights_summarized <- select(flights_summarized, -date)
dyRangeSelector(dygraph(flights_summarized))
```

<br>

The syntax here is a little different than what we have covered so far.  The `dygraph` function is expecting for the dates to be given as the `rownames` of the object.  We then remove the `date` variable from the `flights_summarized` data frame since it is accounted for in the `rownames`.  Lastly, we run the `dygraph` function on the new data frame that only contains the median arrival delay as a column and then provide the ability to have a selector to zoom in on the interactive plot via `dyRangeSelector`.  (Note that this plot will only be interactive in the HTML version of this book.)

<!--
**`paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Use the interactive linegraph to determine the highest median arrival delay for flights from NYC in 2013.  What date was it and what do you think contributed to it?


** ** What are three specific questions that can be more easily answered by looking at Figure 4.6 instead of Figure 4.5?

- Changing the labels of a plot (x-axis, y-axis)
- stat = "identity" for aggregated data and barplots
- Changing the theme for ggplots (`ggthemes` package too)
- Adding `code_folding` and `code_download` to YAML
- `kable` function from `knitr`
- Reading in data from files in different formats - Getting Used to R book reference
- Reshaping the data with `tidyr`

-->
