# Learning Check Solutions {#appendixD}

```{r setup_lc_solutions, include=FALSE, purl=FALSE}
opts_chunk$set(tidy = FALSE, out.width = "\\textwidth")
# This bit of code is a bug fix on asis blocks, which we use to show/not show LC solutions, which are written like markdown text. In theory, it shouldn't be necessary for knitr versions <=1.11.6, but I've found I still need to for everything to knit properly in asis blocks. More info here:
# https://stackoverflow.com/questions/32944715/conditionally-display-block-of-markdown-text-using-knitr
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```

If you detect any errors or have suggestions regarding these solutions, please [file an issue at the ModernDive GitHub repository](https://github.com/moderndive/ModernDive_book/issues) or, ideally, create a Pull Request on GitHub on the same repository making the suggested changes yourself (into the `v2` branch) for us to review. Thanks!

## Chapter 1 Solutions 

```{r child = "lc-answers/01_lc-answers.Rmd"}
```

***

## Chapter 2 Solutions 

```{r child = "lc-answers/02_lc-answers.Rmd"}
```

***

## Chapter 3 Solutions

```{r child = "lc-answers/03_lc-answers.Rmd"}
```

***

## Chapter 4 Solutions

```{r child = "lc-answers/04_lc-answers.Rmd"}
```

***

## Chapter 5 Solutions

```{r child = "lc-answers/05_lc-answers.Rmd"}
```

***

## Chapter 6 Solutions

```{r child = "lc-answers/06_lc-answers.Rmd"}
```

***

## Chapter 7 Solutions

```{r child = "lc-answers/07_lc-answers.Rmd"}
```

***

## Chapter 8 Solutions

```{r child = "lc-answers/08_lc-answers.Rmd"}
```

***

## Chapter 9 Solutions

```{r child = "lc-answers/09_lc-answers.Rmd"}
```

***

## Chapter 10 Solutions

```{r child = "lc-answers/10_lc-answers.Rmd"}
```

***

## Chapter 11 Solutions


```{r child = "lc-answers/11_lc-answers.Rmd"}
```

***

## Appendix A1 Solutions

```{r child = "lc-answers/A1_lc-answers.Rmd"}
```


