# Information about R Packages Used {#appendixE}

This book uses the following versions of R packages (and their dependent packages). If you are seeing results slightly different than what is shown in the book and you want to get a closer match, we recommend you install the particular version of the package we used. This can be done by first installing the `remotes` package via `install.packages("remotes")` and then the particular version of a package using syntax similar to the following replacing the `package` argument with the name of the package in quotes and the `version` argument with the particular number of the version to install.

```{r eval=FALSE}
remotes::install_version(package = "moderndive", version = "0.3.0")
```


```{r colophon, echo=FALSE}
devtools::session_info(needed_pkgs)$packages %>% 
  as_tibble() %>% 
  select(package, version = ondiskversion) %>% 
  knitr::kable(booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = ifelse(knitr:::is_latex_output(), 
                                               10, 16))
```
