# Versions of R Packages Used {#appendixE}

If you are seeing different results than what is in the book, we recommend installing the exact version of the packages we used. This can be done by first installing the `remotes` package via `install.packages("remotes")`. Then, use `install_version()` replacing the `package` argument with the package name in quotes and the `version` argument with the particular version number to install.

```{r eval=FALSE}
remotes::install_version(package = "skimr", version = "1.0.6")
```

<!--
\begin{multicols}{2}
\setbox\ltmcbox\vbox{
\makeatletter\col@number\@ne
-->

```{r colophon, echo=FALSE}
sessioninfo::package_info(needed_pkgs) %>% 
  as_tibble() %>% 
  filter(attached == TRUE | package %in% c("bookdown", "skimr")) %>% 
  select(package, version = ondiskversion) %>% 
  knitr::kable(
    booktabs = TRUE, 
    linesep = "",
    longtable = TRUE
  ) %>% 
  kableExtra::kable_styling(font_size = ifelse(knitr:::is_latex_output(), 
                                               9, 16))
```

\vspace{-2in}
<!--
% Wrap this after the table to get into multiple columns
\unskip
\unpenalty
\unpenalty}
\unvbox\ltmcbox

\end{multicols}
-->

