--- 
title: "ModernDive"
subtitle: "An Introduction to Statistical and Data Sciences via R"
author: "Chester Ismay and Albert Y. Kim"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [bib/books.bib, bib/packages.bib, bib/articles.bib]
biblio-style: apalike
link-citations: yes
github-repo: ismayc/moderndiver-book
description: "Combining statistical and computational thinking to make sense of data. Getting away from the traditional introductory statistics curriculum, more focused on reproducible research, data visualization, and modern data analysis techniques and tools including resampling and bootstrapping using R, RStudio, and R Markdown"
---

```{r set-options, include=FALSE}
options(width = 72)
knitr::opts_chunk$set(tidy = FALSE, out.width='\\textwidth', fig.align = "center")
version <- "0.1.0"
```

# Prerequisites {#prereqs}

This book is in beta testing and is currently at Version `r version`.  If you would like to receive updates on this book and other similar projects, please fill out [this Google Form](https://goo.gl/forms/IxiwBeEnk72NxMMx2) with your name, institution, and email address.

If you'd like to make a GitHub pull request and modify the book as you wish for your own needs (which we both encourage!), please do keep our names on your book.  Something like "YOU, Chester Ismay, and Albert Y. Kim" is perfectly fine.  We'd also appreciate if you let us now what changes you've made and how you've used the textbook.  We'd love some data on what's working well and what's not working so well.

***

This book was written using the **bookdown** R package from Yihui Xie.  In order to follow along and run the code in this book on your own, you'll need to have access to R and RStudio.  You can find more information on both of these with a simple Google search for "R" and for "RStudio."  An introduction to using R, RStudio, and R Markdown is also available in a free book [here](http://ismayc.github.io/rbasics-book) [@usedtor2016].  It is recommended that you refer back to this book frequently as it has GIF screen recordings that you can follow along with as you learn.

We will keep a running list of R packages you will need to have installed to complete the analysis as well here in the `needed_pkgs` character vector.  You can check if you have all of the needed packages installed by running all of the lines below.  The last lines including the `if` will install them as needed (i.e., download their needed files from the internet to your hard drive).

You can run the `library` function on them to load them into your current analysis.  Prior to each analysis where a package is needed, you will see the corresponding `library` function in the text.  Make sure to check the top of the chapter to see if a package was loaded there.

```{r pkg_setup, warning=FALSE, message=FALSE, results='hide'}
needed_pkgs <- c("nycflights13", "dplyr", "ggplot2", "knitr", 
  "okcupiddata", "dygraphs", "rmarkdown", "mosaic", "ggplot2movies")

new.pkgs <- needed_pkgs[!(needed_pkgs %in% installed.packages())]

if(length(new.pkgs)) {
  install.packages(new.pkgs, repos = "http://cran.rstudio.com")
}
```

```{r include=FALSE}
# Packages are needed to create the book, not for following
# the code in the book
needed_pkgs2 <- c("devtools", "webshot", "tufte")
new.pkgs2 <- needed_pkgs2[!(needed_pkgs2 %in% installed.packages())]

if(length(new.pkgs2)) {
  install.packages(new.pkgs2, repos = "http://cran.rstudio.com")
}
# Check that phantomjs is installed
if(is.null(webshot:::find_phantom()))
  webshot::install_phantomjs()
```

**Book was last updated:**

```{r time, echo=FALSE}
paste("By", Sys.info()[["user"]], "on", format(Sys.time(), "%A, %B %d, %Y %X %Z"))
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 
  'nycflights13', 'devtools', 'ggplot2', 'webshot', 
  'dygraphs', 'tufte', 'okcupiddata', 'mosaic',
   'dplyr', 'ggplot2movies'
), 'bib/packages.bib')
```

## Colophon {-}

The source of the book is available [here](https://github.com/ismayc/moderndiver-book) and was built with versions of R packages (and their dependent packages) given below.  This may not be of importance for initial readers of this book, but the hope is you can reproduce a duplicate of this book by installing these versions of the packages.

```{r colophon, echo=FALSE}
knitr::kable(devtools::session_info(needed_pkgs)$packages, 
             booktabs = TRUE,
             longtable = TRUE)
```