--- 
title: '`r ifelse(knitr:::is_latex_output(), "\\Large{A MODERN DIVE into Data with R}", "A MODERN DIVE into Data with R")`'
subtitle: "Modern Open-source Data-analysis Encouraging Reproducibility aNd Driven by Intuitive Visualization Efforts" #"Modern Open-source Data-analysis Encouraging Reproducibility and iNtuitive Data-Inspired Visualization Efforts"
author: "Chester Ismay"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [bib/books.bib, bib/packages.bib, bib/articles.bib]
biblio-style: apalike
link-citations: yes
github-repo: ismayc/moderndiver-source
description: "Getting away from the traditional introductory statistics curriculum, more focused on reproducible research and modern data analysis techniques and tools"
---

```{r set-options, include=FALSE}
options(width = 72)
knitr::opts_chunk$set(tidy = FALSE, fig.align = "center")
```

# Prerequisites {#prereqs}

This book was written using the **bookdown** R package from Yihui Xie.  In order to follow along and run the code in this book on your own, you'll need to have access to R and RStudio.  You can find more information on both of these with a simple Google search for "R" and for "RStudio."  An introduction to using R, RStudio, and R Markdown is also available in a free book [here](http://ismayc.github.io/rbasics-book) [@usedtor2016].  It is recommended that you refer back to this book frequently as it has GIF screen recordings that you can follow along with as you learn.

We will keep a running list of R packages you will need to have installed to complete the analysis as well here in the `needed_pkgs` character vector.  You can check if you have all of the needed packages installed by running all of the lines below.  The last lines including the `if` will install them as needed (i.e., download their needed files from the internet to your hard drive).

You can run the `library` function on them to load them into your current analysis.  Prior to each analysis where a package is needed, you will see the corresponding `library` function in the text.  Make sure to check the top of the chapter to see if a package was loaded there.

```{r include=FALSE, results='hide'}
# Not sure why it is pulling from a GitHub installation w/o this?
# remove.packages("rmarkdown")
```


```{r pkg_setup, warning=FALSE, message=FALSE, results='hide'}
needed_pkgs <- c("nycflights13", "dplyr", "ggplot2", "knitr", 
  "ggplot2movies", "dygraphs", "rmarkdown", "mosaic", "tibble")

new.pkgs <- needed_pkgs[!(needed_pkgs %in% installed.packages())]

if(length(new.pkgs)) {
  install.packages(new.pkgs, repos = "http://cran.rstudio.com")
}
```

```{r include=FALSE}
# Packages are needed to create the book, not for following
# the code in the book
needed_pkgs2 <- c("devtools", "webshot", "tufte")
new.pkgs2 <- needed_pkgs2[!(needed_pkgs2 %in% installed.packages())]

if(length(new.pkgs2)) {
  install.packages(new.pkgs2, repos = "http://cran.rstudio.com")
}
# Check that phantomjs is installed
if(is.null(webshot:::find_phantom()))
  webshot::install_phantomjs()

knitr::opts_chunk$set(tidy = FALSE, fig.align = "center", out.width='\\textwidth')
```

**Book was last updated:**

```{r time, echo=FALSE}
paste("By", Sys.info()[["user"]], "on", format(Sys.time(), "%A, %B %d, %Y %X %Z"))
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 
  'nycflights13', 'devtools', 'ggplot2', 'webshot', 
  'dygraphs', 'tufte', 'ggplot2movies', 'mosaic',
  'tibble'
), 'bib/packages.bib')
```

## Colophon

The source of the book is available [here](https://github.com/ismayc/moderndiver-source) and was built with versions of R packages (and their dependent packages) given below.  This may not be of importance for initial readers of this book, but the hope is you can reproduce a duplicate of this book by installing these versions of the packages.

```{r colophon, echo=FALSE}
knitr::kable(devtools::session_info(needed_pkgs)$packages, 
             booktabs = TRUE,
             longtable = TRUE)
```