
Needed packages

```{r message=FALSE}
library(tidyverse)
library(moderndive)
library(infer)
``` 

**(LC8.1)** What is the expected value of the sample mean weight of almonds in a large sample according to the sampling distribution theory?

**Solution:** **D.**
On average, the sample mean equals the population mean (unbiased), though any single sample can vary.

**(LC8.2)** What is a point estimate and how does it differ from an interval estimate?

**Solution:** **B.**
A point estimate is one number (e.g., $\bar x$). An interval estimate gives a plausible range that accounts for sampling variability.

**(LC8.3)** What does the population mean ($\mu$) represent in the almond activity?

**Solution:** **C.**
It’s the average weight of **all** 5,000 almonds (the full population), not just a sample.

**(LC8.4)** Which best describes the population standard deviation ($\sigma$) in the almond activity?

**Solution:** **B.**
$\sigma$ measures spread around the **population** mean, not the sample mean.

**(LC8.5)** Why use the sample mean to estimate the population mean?

**Solution:** **B.**
$\bar x$ is an unbiased estimator of $\mu$; across repeated samples it hits $\mu$ on average.

**(LC8.6)** How is the standard error of the sample mean calculated here?

**Solution:** **B.**
$SE(\bar x)=\sigma/\sqrt{n}$; larger $n$ shrinks variability of $\bar x$.

**(LC8.7)** What does a 95% confidence interval represent here?

**Solution:** **C.**
Before sampling, the procedure yields intervals that capture $\mu$ 95% of the time; numerically that’s $\bar x \pm 1.96\times SE$.

**(LC8.8)** Why does the $t$ distribution have thicker tails than the standard normal?

**Solution:** **D.**
Using $s$ instead of $\sigma$ adds uncertainty; $t$ accounts for this with heavier tails.

**(LC8.9)** Effect of increasing degrees of freedom on the $t$ distribution?

**Solution:** **B.**
Tails thin and the distribution approaches the standard normal as df increases.

**(LC8.10)** What is the chief difference between a bootstrap distribution and a sampling distribution?

**Solution:** A bootstrap distribution resamples **from the original sample** (with replacement), while a sampling distribution resamples **from the entire population**.

**(LC8.11)** Looking at the bootstrap distribution for the sample mean in Figure 8.14, between what two values would you say most values lie?

**Solution:** Roughly between **3.60 and 3.75 grams**.

**(LC8.12)** Which of the following is true about the confidence level when constructing a confidence interval?

**Solution:** **A.** The confidence level sets the interval’s width and affects how likely it is to contain the true parameter.

**(LC8.13)** How does increasing the sample size affect the width of a confidence interval for a given confidence level?

**Solution:** **C.** It decreases the width, making the estimate more precise.

**(LC8.14)** Construct a 95% confidence interval for the *median* weight of all almonds with the percentile method. Is it appropriate to also use the standard-error method?

**Solution:** The percentile method is appropriate for the median. 

```{r}
bootstrap_distribution <- almonds_sample |>
  specify(response = weight) |>
  generate(reps = 1000) |>
  calculate(stat = "median")
percentile_ci <- bootstrap_distribution |>
  get_confidence_interval(level = 0.95, type = "percentile")
percentile_ci
```

The standard-error method is not ideal, because SE-based formulas assume symmetry/normality, which may not hold for medians. The standard-error method is not appropriate here, because the bootstrap distribution is not bell-shaped:

```{r}
visualize(bootstrap_distribution)
```

**(LC8.15)** What are the advantages of using `infer` for building confidence intervals?

**Solution:** Clearer, step-by-step workflow; consistent syntax across CI and hypothesis testing; easier extension to multi-variable cases; and built-in visualization tools.

**(LC8.16)** What is the main purpose of bootstrapping in statistical inference?

**Solution:** **B.** To generate multiple samples from the original data for estimating parameters.

**(LC8.17)** Which function denotes the variables of interest for inference?

**Solution:** **C.** `specify()`.

**(LC8.18)** What is a key difference between the percentile method and the standard error method for constructing confidence intervals using bootstrap samples?

**Solution:** **B.** Percentile method takes the middle 95% of bootstrap statistics; the SE method centers on the original sample mean and adds/subtracts 1.96 × bootstrap SE.
