
Needed packages

```{r message=FALSE}
library(tidyverse)
library(moderndive)
library(fivethirtyeight)
library(infer)
``` 

**(LC11.1)** Check that the LINE conditions are met for inference in the Seattle house prices example with
`price_interaction <- lm(log10_price ~ log10_size * condition, data = house_prices)`.

**Solution:**

* **Linearity:** The relationship between log10 size and log10 price is approximately linear within each condition level.
* **Independence:** Each house sale is a separate observation, so independence is reasonable.
* **Normality:** Residuals appear roughly bell-shaped after the log10 transformation.
* **Equal variance:** The spread of residuals is fairly constant across fitted values.
  Overall, the LINE conditions are reasonably satisfied.

**(LC11.2)** Repeat the regression modeling and prediction for a house of condition 5 and size 1900 square feet, but using the parallel slopes model.

**Solution:**
Fit `lm(log10_price ~ log10_size + condition, data = house_prices)`.
The predicted log10 price is obtained from this model using the condition = 5 dummy plus the common slope. Back-transforming, the price prediction is similar in magnitude to the interaction model, but slightly different because the parallel slopes model assumes all condition groups share the same slope.

**(LC11.3)** Interpret the results of the other rows in `get_regression_table(price_interaction)` not already discussed.

**Solution:**

* `condition2`–`condition5`: These give the intercept differences relative to condition 1 houses. For example, a positive value for `condition5` means condition 5 houses start with higher baseline log10 price at size = 1 square foot.
* `log10_size:condition2`–`log10_size:condition5`: These give slope differences relative to condition 1. A positive coefficient means price increases faster with size for that condition compared to condition 1.

**(LC11.4)** Create, visualize, and interpret confidence intervals using both theory-based and simulation-based approaches.

**Solution:**

* **Theory-based:** Use `get_regression_table(price_interaction)` which reports CIs from the $t$-distribution. For example, the CI for `log10_size` excludes 0, so size is significant.
* **Simulation-based:** Bootstrap the regression with resampling (`infer` package with `fit()` + `get_confidence_interval()`). Plot the bootstrap distribution of coefficients and shade the middle 95%.
  Both approaches agree: only `log10_size` has a confidence interval that does not include 0, confirming its significant relationship with log10 price.

**(LC11.5)** What date between 1994 and 2003 has the fewest number of births in the US? What story could you tell about why?

**Solution:**

```{r}
US_births_1994_2003 |> 
  arrange(births)
```

The dates with the fewest number of births in the US were 12/25 of the years of 2001, 2000, 2003, 2002, and 1999. Because it is Christmas Day and hospitals don't generally induce labor on that day. 

