### Inference for multiple regression

### Refresher: Professor evaluations data

Let's revisit the professor evaluations data that we analyzed using multiple regression with one numerical and one categorical predictor. In particular

* $y$: outcome variable of instructor evaluation `score`
* predictor variables
    + $x_1$: numerical explanatory/predictor variable of `age`
    + $x_2$: categorical explanatory/predictor variable of `gender`

```{r, echo=FALSE}
library(tidyr)
```
    
```{r}
library(ggplot2)
library(dplyr)
library(moderndive)

evals_multiple <- evals %>%
  select(score, ethnicity, gender, language, age, bty_avg, rank)
```

First, recall that we had two competing potential models to explain professors'
teaching scores:

1. Model 1: No interaction term. i.e. both male and female profs have the same slope describing the associated effect of age on teaching score
1. Model 2: Includes an interaction term. i.e. we allow for male and female profs to have different slopes describing the associated effect of age on teaching score

### Refresher: Visualizations

Recall the plots we made for both these models:

```{r model1, echo=FALSE, warning=FALSE, fig.cap="Model 1: no interaction effect included"}
coeff <- lm(score ~ age + gender, data = evals_multiple) %>% coef() %>% as.numeric()
slopes <- evals_multiple %>%
  group_by(gender) %>%
  summarise(min = min(age), max = max(age)) %>%
  mutate(intercept = coeff[1]) %>%
  mutate(intercept = ifelse(gender == "male", intercept + coeff[3], intercept)) %>%
  gather(point, age, -c(gender, intercept)) %>%
  mutate(y_hat = intercept + age * coeff[2])
  
  ggplot(evals_multiple, aes(x = age, y = score, col = gender)) +
  geom_jitter() +
  labs(x = "Age", y = "Teaching Score", color = "Gender") +
  geom_line(data = slopes, aes(y = y_hat), size = 1)
```

```{r model2, echo=FALSE, warning=FALSE, fig.cap="Model 2: interaction effect included"}
ggplot(evals_multiple, aes(x = age, y = score, col = gender)) +
  geom_jitter() +
  labs(x = "Age", y = "Teaching Score", color = "Gender") +
  geom_smooth(method = "lm", se = FALSE)
```

### Refresher: Regression tables

Last, let's recall the regressions we fit. First, the regression with no 
interaction effect: note the use of `+` in the formula in Table \@ref(tab:modelmultireg).

```{r, eval=FALSE}
score_model_2 <- lm(score ~ age + gender, data = evals_multiple)
get_regression_table(score_model_2)
```
```{r modelmultireg, echo=FALSE}
score_model_2 <- lm(score ~ age + gender, data = evals_multiple)
get_regression_table(score_model_2) %>% 
  knitr::kable(
    digits = 3,
    caption = "Model 1: Regression table with no interaction effect included", 
    booktabs = TRUE
  ) %>% 
  kable_styling(font_size = ifelse(knitr:::is_latex_output(), 10, 16),
                latex_options = c("HOLD_position"))
```

Second, the regression with an interaction effect: note the use of `*` in the formula.

```{r, eval=FALSE}
score_model_3 <- lm(score ~ age * gender, data = evals_multiple)
get_regression_table(score_model_3)
```
```{r, echo=FALSE}
score_model_3 <- lm(score ~ age * gender, data = evals_multiple)
get_regression_table(score_model_3) %>% 
  knitr::kable(
    digits = 3,
    caption = "Model 2: Regression table with interaction effect included", 
    booktabs = TRUE
  ) %>% 
  kable_styling(font_size = ifelse(knitr:::is_latex_output(), 10, 16), 
                latex_options = c("HOLD_position"))
```

***



## Residual analysis

### Residual analysis {#model2residuals}

```{r, eval=TRUE, echo=TRUE}
# Get data:
gapminder2007 <- gapminder %>%
  filter(year == 2007) %>% 
  select(country, continent, lifeExp, gdpPercap)
# Fit regression model:
lifeExp_model <- lm(lifeExp ~ continent, data = gapminder2007)
# Get regression table:
get_regression_table(lifeExp_model)
# Get regression points
regression_points <- get_regression_points(lifeExp_model)
```

Recall our discussion on residuals from Section \@ref(model1residuals) where our goal was to investigate whether or not there was a *systematic pattern* to the residuals. Ideally since residuals can be thought of as error, there should be no such pattern. While there are many ways to do such residual analysis, we focused on two approaches based on visualizations. 

1. A plot with residuals on the vertical axis and the predictor (in this case continent) on the horizontal axis
1. A histogram of all residuals

First, let's plot the residuals versus continent in Figure \@ref(fig:catxplot7), but also let's plot all `r nrow(gapminder2007)` points with a little horizontal random jitter by setting the `width = 0.1` parameter in `geom_jitter()`:

```{r catxplot7, warning=FALSE, fig.cap="Plot of residuals over continent"}
ggplot(regression_points, aes(x = continent, y = residual)) +
  geom_jitter(width = 0.1) + 
  labs(x = "Continent", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue")
```

We observe

1. There seems to be a rough balance of both positive and negative residuals for all 5 continents.
1. However, there is one clear outlier in Asia, which has a residual with the largest deviation away from 0.

Let's investigate the 5 countries in Asia with the shortest life expectancy:

```{r, eval=FALSE}
gapminder2007 %>%
  filter(continent == "Asia") %>%
  arrange(lifeExp)
```
```{r, echo=FALSE}
gapminder2007 %>%
  filter(continent == "Asia") %>%
  arrange(lifeExp) %>%
  slice(1:5) %>%
  knitr::kable(
    digits = 3,
    caption = "Countries in Asia with shortest life expectancy",
    booktabs = TRUE
  ) %>% 
  kable_styling(font_size = ifelse(knitr:::is_latex_output(), 10, 16), 
                latex_options = c("HOLD_position"))
```

This was the earlier identified residual for Afghanistan of -26.9. Unfortunately 
given recent geopolitical turmoil, individuals who live in Afghanistan and, in particular in 2007, have a 
drastically lower life expectancy. 

Second, let's look at a histogram of all `r nrow(gapminder2007)` values of
residuals in Figure \@ref(fig:catxplot8). In this case, the residuals form a
rather nice bell-shape, although there are a couple of very low and very high
values at the tails. As we said previously, searching for patterns in residuals
can be somewhat subjective, but ideally we hope there are no "drastic" patterns.


```{r catxplot8, warning=FALSE, fig.cap="Histogram of residuals"}
ggplot(regression_points, aes(x = residual)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Residual")
```

```{block, type='learncheck', purl=FALSE}
**_Learning check_**
```

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Continuing with our regression using `gdpPercap` as the outcome variable and `continent` as the explanatory variable, use the `get_regression_points()` function to get the observed values, fitted values, and residuals for all `r nrow(gapminder2007)` countries in 2007 and perform a residual analysis to look for any systematic patterns in the residuals. Is there a pattern? Please keep in mind that these types of questions are somewhat subjective and different people will most likely have different answers. The focus should be on being able to justify the conclusions made.

```{block, type='learncheck', purl=FALSE}
```






### Residual analysis {#model3residuals}

Recall in Section \@ref(model1residuals), our first residual analysis plot investigated the presence of any systematic pattern in the residuals when we had a single numerical predictor: `bty_age`. For the `Credit` card dataset, since we have two numerical predictors, `Limit` and `Income`, we must perform this twice:

```{r, eval=TRUE, echo=TRUE}
# Get data:
Credit <- Credit %>%
  select(Balance, Limit, Income, Rating, Age)
# Fit regression model:
Balance_model <- lm(Balance ~ Limit + Income, data = Credit)
# Get regression table:
get_regression_table(Balance_model)
# Get regression points
regression_points <- get_regression_points(Balance_model)
```

```{r, eval=FALSE}
ggplot(regression_points, aes(x = Limit, y = residual)) +
  geom_point() +
  labs(x = "Credit limit (in $)", 
       y = "Residual", 
       title = "Residuals vs credit limit")
  
ggplot(regression_points, aes(x = Income, y = residual)) +
  geom_point() +
  labs(x = "Income (in $1000)", 
       y = "Residual", 
       title = "Residuals vs income")
```

```{r, echo=FALSE, fig.height=4, fig.cap="Residuals vs credit limit and income"}
model3_residual_vs_limit_plot <- ggplot(regression_points, aes(x = Limit, y = residual)) +
  geom_point() +
  labs(x = "Credit limit (in $)", y = "Residual", 
       title = "Residuals vs credit limit")
model3_residual_vs_income_plot <- ggplot(regression_points, aes(x = Income, y = residual)) +
  geom_point() +
  labs(x = "Income (in $1000)", y = "Residual", 
       title = "Residuals vs income")
model3_residual_vs_limit_plot + model3_residual_vs_income_plot
```

In this case, there **does** appear to be a systematic pattern to the residuals. As the scatter of the residuals around the line $y=0$ is definitely not consistent. This behavior of the residuals is further evidenced by the histogram of residuals in Figure \@ref(fig:model3-residuals-hist). We observe that the residuals have a slight right-skew (recall we say that data is right-skewed, or positively-skewed, if there is a tail to the right). Ideally, these residuals should be bell-shaped around a residual value of 0. 

```{r model3-residuals-hist, fig.height=4, fig.cap="Relationship between credit card balance and credit limit/income"}
ggplot(regression_points, aes(x = residual)) +
  geom_histogram(color = "white") +
  labs(x = "Residual")
```

Another way to interpret this histogram is that since the residual is computed as $y - \widehat{y}$ = `balance` - `balance_hat`, we have some values where the fitted value $\widehat{y}$ is very much lower than the observed value $y$. In other words, we are underestimating certain credit card holders' balances by a very large amount. 

```{block, type='learncheck', purl=FALSE}
**_Learning check_**
```

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Continuing with our regression using `Rating` and `Age` as the explanatory variables and credit card `Balance` as the outcome variable, use the `get_regression_points()` function to get the observed values, fitted values, and residuals for all `r nrow(Credit)` credit card holders. Perform a residual analysis and look for any systematic patterns in the residuals.

```{block, type='learncheck', purl=FALSE}
```





### Residual analysis {#model4residuals}

```{r, eval=TRUE, echo=TRUE}
# Get data:
evals_ch7 <- evals %>%
  select(score, age, gender)
# Fit regression model:
score_model_2 <- lm(score ~ age + gender, data = evals_ch7)
# Get regression table:
get_regression_table(score_model_2)
# Get regression points
regression_points <- get_regression_points(score_model_2)
```


As always, let's perform a residual analysis first with a histogram, which we can facet by `gender`:

```{r residual1, warning=FALSE, fig.cap="Interaction model histogram of residuals"}
ggplot(regression_points, aes(x = residual)) +
  geom_histogram(binwidth = 0.25, color = "white") +
  labs(x = "Residual") +
  facet_wrap(~gender)
```

Second, the residuals as compared to the predictor variables:

* $x_1$: numerical explanatory/predictor variable of `age`
* $x_2$: categorical explanatory/predictor variable of `gender`

```{r residual2, warning=FALSE, fig.cap="Interaction model residuals vs predictor"}
ggplot(regression_points, aes(x = age, y = residual)) +
  geom_point() +
  labs(x = "age", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1) +
  facet_wrap(~ gender)
```
