Solutions for the (short) end-of-chapter exercises in *Chapter 1 — Getting Started with Data in R*. These prompts focus on inspecting `olympic_athletes` from `olympicAthletes` — a gentle on-ramp to the larger exercise sets in *Chapter 2* onward.

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                          | # of exercises |
|---------------------------------------|---------------:|
| 1.1 What are R and RStudio?            | 1              |
| 1.2 How do I code in R?                | 1              |
| 1.3 What are R packages?               | 2              |
| 1.4 Explore your first datasets        | 8              |
| Forward references (Ch 3 pipe `\|>`)    | 1              |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                          | Exercises                          |
|-----------------------------------------------------|------------------------------------|
| 1.1 What are R and RStudio? (overview)               | EX1.1                              |
| 1.2.2 Errors, warnings, and messages                 | EX1.2                              |
| 1.3.2 Package installation                           | EX1.3                              |
| 1.3.3 Package loading                                | EX1.4                              |
| 1.4 Explore your first datasets (overview)           | EX1.5, EX1.6, EX1.7, EX1.12        |
| 1.4.4 Identification and measurement variables       | EX1.8, EX1.9                       |
| 1.4.5 Help files                                     | EX1.10, EX1.11                     |
| Forward reference: pipe operator `\|>` (Ch 3)         | EX1.13                             |

**A note on coverage.** Chapter 1 is intentionally a warm-up set — the larger exercise sets begin in *Chapter 2*. Sections like *1.1 Installing R and RStudio* are largely procedural (a one-time setup walkthrough), so they're covered by a single conceptual prompt rather than a battery of code-based ones. Exercises now follow the chapter's content order: R/RStudio (1.1) → coding & messages (1.2) → packages (1.3) → first datasets (1.4) → forward reference to the pipe (Ch 3).

Difficulty: ★ warm-up.
:::

```{r 01-ex-setup, message=FALSE}
library(dplyr)
library(olympicAthletes)
```

::: {.callout-note collapse="true" title="EX1.1 — show question"}
**EX1.1 (★)** A classmate says "R and RStudio are the same thing." Set them straight using an everyday analogy.
:::

**Solution** *(reference: Section 1.1 "What are R and RStudio?")*. *Answers will vary; an exemplar:*

> **R is the engine; RStudio is the dashboard.** R is the actual programming language and the program that *runs* your code — type `mean(x)` and R is what computes the answer. RStudio is the wraparound *interface* you click in: the editor where you write scripts, the console where output appears, the plot pane, the file browser, and the help viewer all stitched into one window. R works perfectly well without RStudio (you can run it from a bare terminal), but RStudio is useless without R installed underneath — it has no engine of its own.

Other analogies students reach for and that earn full credit: *R is the recipe ingredients; RStudio is the kitchen.* *R is the writing language (English); RStudio is the word processor (Microsoft Word).* *R is the gameplay code; RStudio is the game launcher.* The key idea — separating the *language/runtime* from the *interface* — is what matters; the metaphor is just a vehicle for it.

***

::: {.callout-note collapse="true" title="EX1.2 — show question"}
**EX1.2 (★)** Distinguish between an *error*, a *warning*, and a *message* in R. Which halts execution? One example each.
:::

**Solution** *(reference: Section 1.2.2 "Errors, warnings, and messages")*.

- **Error** — *halts* execution. R cannot continue. Example: `mean("hello")` (cannot take the mean of a string).
- **Warning** — execution *continues*; R flags something it computed but is unsure about. Example: `as.numeric("xyz")` returns `NA` plus the warning `NAs introduced by coercion`.
- **Message** — purely informational; nothing went wrong. Example: `library(dplyr)` prints attaching/conflict notices.

Rule of thumb: errors are red flags, warnings are yellow flags, messages are just narration.

***

::: {.callout-note collapse="true" title="EX1.3 — show question"}
**EX1.3 (★)** `library()` vs `install.packages()`.
:::

**Solution** *(reference: Section 1.3.2 "Package installation" + 1.3.3 "Package loading")*. `install.packages()` downloads and installs the package files **once** per machine (think: installing an app). `library()` **loads** the package into your current R session (think: opening the app) — required every session.

***

::: {.callout-note collapse="true" title="EX1.4 — show question"}
**EX1.4 (★)** Fresh R session, `glimpse(olympic_athletes)` with no `library()` calls — what error and what's the fix?
:::

**Solution** *(reference: Section 1.3.3 "Package loading")*. You'll get **`could not find function "glimpse"`** (because `dplyr` — which provides `glimpse()` — isn't loaded). Even if you reach for `glimpse()`'s namespace, you'd then hit **`object 'olympic_athletes' not found`** because `olympicAthletes` isn't loaded either.

**Fix:** load both packages at the top of the session:

```{r 01-ex-4, eval=FALSE}
library(dplyr)            # provides glimpse()
library(olympicAthletes)  # provides olympic_athletes
glimpse(olympic_athletes)
```

`install.packages()` is a *one-time* per-machine step; `library()` is a *once-per-session* step.

***

::: {.callout-note collapse="true" title="EX1.5 — show question"}
**EX1.5 (★)** `glimpse(olympic_athletes)` — rows and columns?
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets")*. **314,749 rows × 15 columns**.

```{r 01-ex-5}
dim(olympic_athletes)
```

***

::: {.callout-note collapse="true" title="EX1.6 — show question"}
**EX1.6 (★)** What does one row represent?
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets", observational units)*. One row represents **one athlete-event participation** — i.e., one athlete competing in one event at one Games. A single athlete who competed in three events at one Games has three rows; a swimmer who competed in two Games has rows for each.

***

::: {.callout-note collapse="true" title="EX1.7 — show question"}
**EX1.7 (★)** Pick three columns; describe each.
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets", inspecting variables)*. *Examples:*

- `Sex`: `"M"` or `"F"` — the athlete's sex as recorded.
- `Year`: the calendar year of the Games (e.g., 2024).
- `Medal`: `"Gold"`, `"Silver"`, `"Bronze"`, or `NA` for non-medalists.

***

::: {.callout-note collapse="true" title="EX1.8 — show question"}
**EX1.8 (★)** Two ID + three measurement variables.
:::

**Solution** *(reference: Section 1.4.4 "Identification and measurement variables")*. **Identifiers**: `ID` (athlete ID), `Name` (athlete name). **Measurements**: `Age`, `Height`, `Weight`, `Sport`, `Medal` — describe properties of each row.

***

::: {.callout-note collapse="true" title="EX1.9 — show question"}
**EX1.9 (★)** Classify columns as numerical or categorical.
:::

**Solution** *(reference: Section 1.4.4 "Identification and measurement variables")*. Numerical: `Age`, `Height`, `Year`. Categorical: `Sex`, `Sport`, `Medal`. **Year** is technically numerical but often *behaves* categorically in plots (boxplots by year, etc.).

***

::: {.callout-note collapse="true" title="EX1.10 — show question"}
**EX1.10 (★)** `?olympic_athletes` — what does the help file say about `Medal`?
:::

**Solution** *(reference: Section 1.4.5 "Help files")*. The help file documents `Medal` as `"Gold" / "Silver" / "Bronze" / NA`, with a note that **every member of a medal-winning team's roster gets a row with the team's medal**.

***

::: {.callout-note collapse="true" title="EX1.11 — show question"}
**EX1.11 (★)** What does `NOC` abbreviate?
:::

**Solution** *(reference: Section 1.4.5 "Help files", domain context)*. **National Olympic Committee** — a 3-letter IOC code (e.g., `"USA"`, `"GBR"`, `"NOR"`). The `Team` column shows the long-form name; `NOC` is the canonical short code.

***

::: {.callout-note collapse="true" title="EX1.12 — show question"}
**EX1.12 (★)** Describe one row of `medal_table`.
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets", observational units)*. One row of `medal_table` represents **one (Games × NOC) combination** — e.g., "USA at the 2024 Summer Games". The observational unit is *coarser* than `olympic_athletes` — there are 273 Games-NOC pairs vs ~315k athlete-rows.

```{r 01-ex-12}
glimpse(medal_table)
```

***

::: {.callout-note collapse="true" title="EX1.13 — show question"}
**EX1.13 (★)** Two ways to call `glimpse()`.
:::

**Solution** *(reference: Section 1.3.4 "Package use", function-call style — preview of *the pipe*)*. The two snippets produce the **same output**. `glimpse(olympic_athletes)` and `olympic_athletes |> glimpse()` are equivalent — the pipe just sends the left-hand side as the first argument of the right-hand function. We'll formally introduce the pipe in *Chapter 3 — Data Wrangling*.

***
