Solutions for the (short) end-of-chapter exercises in *Chapter 1 — Getting Started with Data in R*. These prompts focus on inspecting `olympic_athletes` from `olympicAthletes` — a gentle on-ramp to the larger exercise sets in *Chapter 2* onward.

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

| Section in chapter                                  | Exercises             |
|-----------------------------------------------------|-----------------------|
| 1.4 Explore your first datasets                      | EX1.1, EX1.2, EX1.3   |
| 1.4.4 Identification and measurement variables        | EX1.4, EX1.7          |
| 1.4.5 Help files                                      | EX1.5, EX1.6          |
| 1.3 Packages                                          | EX1.8                 |
| Forward references                                    | EX1.9, EX1.10         |

Difficulty: ★ warm-up.
:::

```{r 01-ex-setup, message=FALSE}
library(dplyr)
library(olympicAthletes)
```

::: {.callout-note collapse="true" title="EX1.1 — show question"}
**EX1.1 (★)** `glimpse(olympic_athletes)` — rows and columns?
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets")*. **314,749 rows × 15 columns**.

```{r 01-ex-1}
dim(olympic_athletes)
```

***

::: {.callout-note collapse="true" title="EX1.2 — show question"}
**EX1.2 (★)** What does one row represent?
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets", observational units)*. One row represents **one athlete-event participation** — i.e., one athlete competing in one event at one Games. A single athlete who competed in three events at one Games has three rows; a swimmer who competed in two Games has rows for each.

***

::: {.callout-note collapse="true" title="EX1.3 — show question"}
**EX1.3 (★)** Pick three columns; describe each.
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets", inspecting variables)*. *Examples:*

- `Sex`: `"M"` or `"F"` — the athlete's sex as recorded.
- `Year`: the calendar year of the Games (e.g., 2024).
- `Medal`: `"Gold"`, `"Silver"`, `"Bronze"`, or `NA` for non-medalists.

***

::: {.callout-note collapse="true" title="EX1.4 — show question"}
**EX1.4 (★)** Two ID + three measurement variables.
:::

**Solution** *(reference: Section 1.4.4 "Identification and measurement variables")*. **Identifiers**: `ID` (athlete ID), `Name` (athlete name). **Measurements**: `Age`, `Height`, `Weight`, `Sport`, `Medal` — describe properties of each row.

***

::: {.callout-note collapse="true" title="EX1.5 — show question"}
**EX1.5 (★)** `?olympic_athletes` — what does the help file say about `Medal`?
:::

**Solution** *(reference: Section 1.4.5 "Help files")*. The help file documents `Medal` as `"Gold" / "Silver" / "Bronze" / NA`, with a note that **every member of a medal-winning team's roster gets a row with the team's medal**.

***

::: {.callout-note collapse="true" title="EX1.6 — show question"}
**EX1.6 (★)** What does `NOC` abbreviate?
:::

**Solution** *(reference: Section 1.4.5 "Help files", domain context)*. **National Olympic Committee** — a 3-letter IOC code (e.g., `"USA"`, `"GBR"`, `"NOR"`). The `Team` column shows the long-form name; `NOC` is the canonical short code.

***

::: {.callout-note collapse="true" title="EX1.7 — show question"}
**EX1.7 (★)** Classify columns as numerical or categorical.
:::

**Solution** *(reference: Section 1.4.4 "Identification and measurement variables")*. Numerical: `Age`, `Height`, `Year`. Categorical: `Sex`, `Sport`, `Medal`. **Year** is technically numerical but often *behaves* categorically in plots (boxplots by year, etc.).

***

::: {.callout-note collapse="true" title="EX1.8 — show question"}
**EX1.8 (★)** `library()` vs `install.packages()`.
:::

**Solution** *(reference: Section 1.3.2 "Package installation" + 1.3.3 "Package loading")*. `install.packages()` downloads and installs the package files **once** per machine (think: installing an app). `library()` **loads** the package into your current R session (think: opening the app) — required every session.

***

::: {.callout-note collapse="true" title="EX1.9 — show question"}
**EX1.9 (★)** Two ways to call `glimpse()`.
:::

**Solution** *(reference: Section 1.3.4 "Package use", function-call style — preview of *the pipe*)*. The two snippets produce the **same output**. `glimpse(olympic_athletes)` and `olympic_athletes |> glimpse()` are equivalent — the pipe just sends the left-hand side as the first argument of the right-hand function. We'll formally introduce the pipe in *Chapter 3 — Data Wrangling*.

***

::: {.callout-note collapse="true" title="EX1.10 — show question"}
**EX1.10 (★)** Describe one row of `medal_table`.
:::

**Solution** *(reference: Section 1.4 "Explore your first datasets", observational units)*. One row of `medal_table` represents **one (Games × NOC) combination** — e.g., "USA at the 2024 Summer Games". The observational unit is *coarser* than `olympic_athletes` — there are 273 Games-NOC pairs vs ~315k athlete-rows.

```{r 01-ex-10}
glimpse(medal_table)
```

***
