Solutions for the end-of-chapter exercises in *Chapter 3 — Data Wrangling*. Datasets: `olympic_athletes` and `medal_table` from [`olympicAthletes`](https://github.com/moderndive/olympicAthletes), and `episodes` from [`steves`](https://github.com/ismayc/steves) — a tidied snapshot of every *Rick Steves' Europe* episode (159 rows × 38 columns).

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                          | # of exercises |
|---------------------------------------|---------------:|
| 3.1 The pipe operator `\|>`             | 1              |
| 3.2 `filter` rows                      | 5              |
| 3.3 `summarize` + 3.4 `group_by`       | 7              |
| 3.5 `mutate`                           | 4              |
| 3.6 `arrange`                          | 3              |
| 3.7 `join` data frames                 | 4              |
| 3.8 Other verbs                        | 4              |
| Setup / data exploration               | 3              |
| Critical thinking / synthesis          | 6              |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                  | Exercises                                                |
|---------------------------------------------|----------------------------------------------------------|
| Setup and data exploration                  | EX3.1, EX3.2, EX3.3                                      |
| 3.1 The pipe operator `\|>`                  | EX3.4                                                    |
| 3.2 `filter()`                              | EX3.5, EX3.6, EX3.7, EX3.8, EX3.9                        |
| 3.3 `summarize()` + 3.4 `group_by()`        | EX3.10, EX3.11, EX3.12, EX3.13, EX3.14, EX3.15, EX3.16   |
| 3.5 `mutate()`                              | EX3.17, EX3.18, EX3.19, EX3.20                           |
| 3.6 `arrange()`                             | EX3.21, EX3.22, EX3.23                                   |
| 3.7 `join()` family (inner/left/anti/semi)  | EX3.24, EX3.25, EX3.26, EX3.27                           |
| 3.8.1 `select()`                            | EX3.28, EX3.29, EX3.30                                   |
| 3.8.2 `relocate()` + 3.8.3 `rename()`        | EX3.31                                                   |
| Critical thinking / synthesis               | EX3.32, EX3.33, EX3.34, EX3.35, EX3.36, EX3.37           |

**A note on coverage.** Sections 3.8.2 (`relocate`) and 3.8.3 (`rename`) are short (≈10–15 lines each) — both verbs are *trivial reorderings* of an existing data frame, so a single combined exercise (EX3.31) exercises them alongside `select()` rather than padding three near-duplicate prompts. The pipe operator (3.1) is implicitly practiced in every other exercise; EX3.4 makes the *nested-vs-piped* equivalence explicit. Exercises now follow chapter content order: setup → pipe → filter → group_by/summarize → mutate → arrange → joins → other verbs → critical thinking.

Difficulty stars on each exercise: ★ warm-up, ★★ standard application, ★★★ critical thinking.
:::

```{r 03-ex-setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(olympicAthletes)
library(steves)
```

### Setup

::: {.callout-note collapse="true" title="EX3.1 — show question"}
**EX3.1 (★)** Load `steves` and run `glimpse(episodes)`. How many rows and columns? Name three numerical and three categorical variables.
:::

**Solution** *(reference: Chapter 1, "Identification and measurement variables")*. **159 × 38**. Numerical: `season`, `imdb_rating`, `imdb_votes`, `lat`, `long`, `best_runtime`. Categorical: `primary_country`, `region`, `episode_type`, `theme_tags`, `iso2`. Date: `original_air_date`.

```{r 03-ex-1}
glimpse(episodes)
```

***

::: {.callout-note collapse="true" title="EX3.2 — show question"}
**EX3.2 (★)** Predict, then verify: what is the range of `imdb_rating` in `episodes`? What is the range of `season`?
:::

**Solution** *(reference: Chapter 2 overview, exploring data)*. IMDB ratings span ~7 to ~9 (the show is well-loved). Seasons run 1–13.

```{r 03-ex-2}
range(episodes$imdb_rating, na.rm = TRUE)
range(episodes$season)
```

***

::: {.callout-note collapse="true" title="EX3.3 — show question"}
**EX3.3 (★)** `original_air_date` has type `Date`. Why does that matter for arithmetic compared to storing it as `character`?
:::

**Solution** *(reference: Section 3.5 "mutate existing variables", date arithmetic)*. With `Date`, you can compute differences directly (`date2 - date1` returns days), arrange chronologically without lexicographic surprises, and extract components with `lubridate::year()`/`month()`. Strings would need parsing every time.

***

### The pipe operator `|>`

::: {.callout-note collapse="true" title="EX3.4 — show question"}
**EX3.4 (★)** Rewrite a nested expression as a `|>`-piped chain — verify both produce the same output. Why is the piped version usually easier to *read*?
:::

**Solution** *(reference: Section 3.1 "The pipe operator `|>`")*. The two forms are equivalent. The pipe reads **left-to-right, top-to-bottom** — the same direction English readers move through text — whereas the nested form reads **inside-out** (innermost call first). For three or more steps the pipe wins decisively.

```{r 03-ex-4}
# Nested form
arrange(filter(olympic_athletes, Year == 2024, !is.na(Height)), desc(Height)) |>
  head(3)

# Piped form — same data, easier to scan
olympic_athletes |>
  filter(Year == 2024, !is.na(Height)) |>
  arrange(desc(Height)) |>
  head(3)
```

The piped chain also makes it trivial to *add* or *remove* a step (just insert/delete a `|> verb()` line); editing the nested form requires wrangling parentheses.

***

### `filter()`

::: {.callout-note collapse="true" title="EX3.5 — show question"}
**EX3.5 (★)** Filter `olympic_athletes` to athletes from `Sport == "Swimming"` who won a `Gold` medal.
:::

**Solution** *(reference: Section 3.2 "filter rows")*.

```{r 03-ex-5}
olympic_athletes |>
  filter(Sport == "Swimming", Medal == "Gold") |>
  nrow()
```

***

::: {.callout-note collapse="true" title="EX3.6 — show question"}
**EX3.6 (★)** Filter `episodes` to those where `primary_country == "Italy"`. How many?
:::

**Solution** *(reference: Section 3.2 "filter rows")*.

```{r 03-ex-6}
episodes |> filter(primary_country == "Italy") |> nrow()
```

Italy is the most-visited single country (~29 episodes).

***

::: {.callout-note collapse="true" title="EX3.7 — show question"}
**EX3.7 (★)** Filter `episodes` to **highly-rated AND well-watched**: `imdb_rating >= 8` AND `imdb_votes >= 50`.
:::

**Solution** *(reference: Section 3.2 "filter rows", boolean operators)*.

```{r 03-ex-7}
episodes |>
  filter(imdb_rating >= 8, imdb_votes >= 50) |>
  select(season, episode_in_season, title, imdb_rating, imdb_votes)
```

***

::: {.callout-note collapse="true" title="EX3.8 — show question"}
**EX3.8 (★)** Filter `episodes` to the union: `"Italy"` OR `"France"`. Use `%in%`.
:::

**Solution** *(reference: Section 3.2 "filter rows", `%in%`)*.

```{r 03-ex-8}
episodes |>
  filter(primary_country %in% c("Italy", "France")) |>
  count(primary_country)
```

***

::: {.callout-note collapse="true" title="EX3.9 — show question"}
**EX3.9 (★★)** Filter `olympic_athletes` to the most recent Summer Games. How many distinct athletes?
:::

**Solution** *(reference: Section 3.2 "filter rows")*.

```{r 03-ex-9}
olympic_athletes |>
  filter(Year == 2024, Season == "Summer") |>
  summarize(n_distinct_athletes = n_distinct(ID))
```

***

### `group_by` + `summarize`

::: {.callout-note collapse="true" title="EX3.10 — show question"}
**EX3.10 (★★)** Mean `Height` per `Sport`, top 10 tallest.
:::

**Solution** *(reference: Section 3.4 "group_by rows" + 3.3 "summarize variables")*.

```{r 03-ex-10}
olympic_athletes |>
  group_by(Sport) |>
  summarize(mean_height = mean(Height, na.rm = TRUE), .groups = "drop") |>
  slice_max(mean_height, n = 10)
```

***

::: {.callout-note collapse="true" title="EX3.11 — show question"}
**EX3.11 (★★)** Total medals per NOC across the five recent Games — top 10.
:::

**Solution** *(reference: Section 3.4 "group_by rows")*.

```{r 03-ex-11}
medal_table |>
  group_by(noc, country) |>
  summarize(total = sum(total), .groups = "drop") |>
  slice_max(total, n = 10)
```

***

::: {.callout-note collapse="true" title="EX3.12 — show question"}
**EX3.12 (★★)** Median IMDB rating per `region` in `episodes`.
:::

**Solution** *(reference: Section 3.4 "group_by rows")*.

```{r 03-ex-12}
episodes |>
  group_by(region) |>
  summarize(median_rating = median(imdb_rating, na.rm = TRUE),
            n_episodes    = n(),
            .groups = "drop") |>
  arrange(desc(median_rating))
```

***

::: {.callout-note collapse="true" title="EX3.13 — show question"}
**EX3.13 (★★)** Episodes per `season` of *Rick Steves' Europe* — barplot.
:::

**Solution** *(reference: Section 3.4 "group_by rows" + Section 2.8 "Barplots")*.

```{r 03-ex-13, fig.height=3}
episodes |>
  count(season) |>
  ggplot(aes(x = factor(season), y = n)) +
  geom_col() +
  labs(x = "Season", y = "Number of episodes")
```

***

::: {.callout-note collapse="true" title="EX3.14 — show question"}
**EX3.14 (★★)** Distinct *athletes* per `NOC`, top 10.
:::

**Solution** *(reference: Section 3.3 "summarize variables", `n_distinct()`)*.

```{r 03-ex-14}
olympic_athletes |>
  group_by(NOC) |>
  summarize(n_athletes = n_distinct(ID), .groups = "drop") |>
  slice_max(n_athletes, n = 10)
```

***

::: {.callout-note collapse="true" title="EX3.15 — show question"}
**EX3.15 (★★★)** Mean `Age` per `Sport`, only sports with ≥ 1,000 athlete-rows.
:::

**Solution** *(reference: Section 3.3 "summarize variables", `n()` for sample sizes)*. The post-summary `filter` keeps the comparison meaningful — averages computed on tiny samples are noisy.

```{r 03-ex-15}
olympic_athletes |>
  group_by(Sport) |>
  summarize(n = n(), mean_age = mean(Age, na.rm = TRUE),
            .groups = "drop") |>
  filter(n >= 1000) |>
  arrange(mean_age) |>
  slice_head(n = 5)
```

***

::: {.callout-note collapse="true" title="EX3.16 — show question"}
**EX3.16 (★★)** Distinct events per `Games` — has the number grown?
:::

**Solution** *(reference: Section 3.4 "group_by rows")*. Yes, the long-run trend is upward (more disciplines, more events per discipline).

```{r 03-ex-16, fig.height=3}
olympic_athletes |>
  group_by(Year, Season) |>
  summarize(n_events = n_distinct(Event), .groups = "drop") |>
  ggplot(aes(x = Year, y = n_events, color = Season)) +
  geom_line() + geom_point() +
  labs(y = "Distinct events")
```

***

### `mutate()`

::: {.callout-note collapse="true" title="EX3.17 — show question"}
**EX3.17 (★★)** Add a `height_in` column converting `Height` (cm) to inches.
:::

**Solution** *(reference: Section 3.5 "mutate existing variables")*.

```{r 03-ex-17}
olympic_athletes |>
  mutate(height_in = Height / 2.54) |>
  select(Name, Height, height_in) |>
  head()
```

***

::: {.callout-note collapse="true" title="EX3.18 — show question"}
**EX3.18 (★★)** Recompute `air_year` from `original_air_date`.
:::

**Solution** *(reference: Section 3.5 "mutate existing variables", date functions)*.

```{r 03-ex-18}
episodes |>
  mutate(air_year_recomputed = as.integer(format(original_air_date, "%Y"))) |>
  summarize(matches = all(air_year_recomputed == air_year, na.rm = TRUE))
```

***

::: {.callout-note collapse="true" title="EX3.19 — show question"}
**EX3.19 (★★)** Add `won_medal` (`TRUE`/`FALSE`). What proportion of athlete-rows won a medal?
:::

**Solution** *(reference: Section 3.5 "mutate existing variables", boolean coercion)*.

```{r 03-ex-19}
olympic_athletes |>
  mutate(won_medal = !is.na(Medal)) |>
  summarize(pct_with_medal = mean(won_medal))
```

About 14% of athlete-rows are part of a medal-winning entry. (Team events count each roster member.)

***

::: {.callout-note collapse="true" title="EX3.20 — show question"}
**EX3.20 (★★)** Add ordinal `age_bucket` using `cut()`.
:::

**Solution** *(reference: Section 3.5 "mutate existing variables", binning)*.

```{r 03-ex-20}
olympic_athletes |>
  mutate(age_bucket = cut(Age, breaks = c(0, 18, 22, 28, 35, 100),
                          right = FALSE)) |>
  count(age_bucket)
```

***

### `arrange()`

::: {.callout-note collapse="true" title="EX3.21 — show question"}
**EX3.21 (★★)** 10 oldest athletes by `Age`.
:::

**Solution** *(reference: Section 3.6 "arrange and sort rows")*.

```{r 03-ex-21}
olympic_athletes |>
  arrange(desc(Age)) |>
  select(Name, Age, Sport, Year) |>
  slice_head(n = 10)
```

***

::: {.callout-note collapse="true" title="EX3.22 — show question"}
**EX3.22 (★★)** 5 highest-rated *Rick Steves' Europe* episodes.
:::

**Solution** *(reference: Section 3.6 "arrange and sort rows")*.

```{r 03-ex-22}
episodes |>
  arrange(desc(imdb_rating)) |>
  select(season, episode_in_season, title, imdb_rating) |>
  slice_head(n = 5)
```

***

::: {.callout-note collapse="true" title="EX3.23 — show question"}
**EX3.23 (★★)** Sort by `season` asc, then `imdb_rating` desc.
:::

**Solution** *(reference: Section 3.6 "arrange and sort rows", multi-key)*.

```{r 03-ex-23}
episodes |>
  arrange(season, desc(imdb_rating)) |>
  select(season, episode_in_season, title, imdb_rating) |>
  slice_head(n = 10)
```

***

### `join()`

::: {.callout-note collapse="true" title="EX3.24 — show question"}
**EX3.24 (★★★)** Inner-join `olympic_athletes` with `editions` (5 rows). Resulting row count?
:::

**Solution** *(reference: Section 3.7 "join data frames", inner_join)*. `editions` only covers 2018–2026, so an inner join keeps only athletes from those Games. The original 314,749 rows shrink to the count of athletes in those 5 Games.

```{r 03-ex-24}
olympic_athletes |>
  inner_join(editions, by = c("Games" = "games")) |>
  nrow()
```

***

::: {.callout-note collapse="true" title="EX3.25 — show question"}
**EX3.25 (★★★)** Left-join instead. Why is the row count different?
:::

**Solution** *(reference: Section 3.7 "join data frames", left_join)*. `left_join` preserves *all* athlete rows; rows whose `Games` value isn't in `editions` get `NA`s for the joined columns. So the count matches the original 314,749.

```{r 03-ex-25}
olympic_athletes |>
  left_join(editions, by = c("Games" = "games")) |>
  nrow()
```

***

::: {.callout-note collapse="true" title="EX3.26 — show question"}
**EX3.26 (★★★)** Anti-join: athletes whose NOC didn't appear in `medal_table` for that Games.
:::

**Solution** *(reference: Section 3.7 "join data frames", anti_join + filtering joins)*. These are NOCs that participated but won no medals at that Games — the bulk of NOCs in any given Olympics.

```{r 03-ex-26}
olympic_athletes |>
  filter(Year >= 2018) |>
  anti_join(medal_table, by = c("Year" = "year", "NOC" = "noc")) |>
  count(Year, Season, NOC, sort = TRUE) |>
  slice_head(n = 10)
```

***

::: {.callout-note collapse="true" title="EX3.27 — show question"}
**EX3.27 (★★★)** Semi-join: keep only athletes from medal-winning NOCs.
:::

**Solution** *(reference: Section 3.7 "join data frames", semi_join)*. `semi_join()` filters `x` rows to those that match `y` — without bringing along `y`'s columns.

```{r 03-ex-27}
olympic_athletes |>
  filter(Year >= 2018) |>
  semi_join(medal_table, by = c("Year" = "year", "NOC" = "noc")) |>
  nrow()
```

***

### Other verbs (`select`, `relocate`, `rename`)

::: {.callout-note collapse="true" title="EX3.28 — show question"}
**EX3.28 (★)** Select identifier and medal-related columns from `olympic_athletes`.
:::

**Solution** *(reference: Section 3.8.1 "select variables")*.

```{r 03-ex-28}
olympic_athletes |>
  select(ID, Name, Medal, Event, Sport, Year) |>
  head()
```

***

::: {.callout-note collapse="true" title="EX3.29 — show question"}
**EX3.29 (★)** Use a `tidyselect` helper to keep only the IMDB-related columns of `episodes`.
:::

**Solution** *(reference: Section 3.8.1 "select variables", helpers)*.

```{r 03-ex-29}
episodes |> select(starts_with("imdb")) |> head()
```

***

::: {.callout-note collapse="true" title="EX3.30 — show question"}
**EX3.30 (★)** Drop `image_url` and `tvmaze_url` from `episodes`.
:::

**Solution** *(reference: Section 3.8.1 "select variables", negative selection)*.

```{r 03-ex-30}
episodes |> select(-image_url, -tvmaze_url) |> ncol()
```

***

::: {.callout-note collapse="true" title="EX3.31 — show question"}
**EX3.31 (★★)** Use `select()`, `relocate()`, and `rename()` together to clean up a `medal_table` view — keep 6 columns, rename `noc` → `iso_code`, put `country` first.
:::

**Solution** *(reference: Section 3.8 "Other verbs", `select` + `relocate` + `rename`)*.

```{r 03-ex-31}
medal_table |>
  select(noc, country, gold, silver, bronze, total) |>
  rename(iso_code = noc) |>
  relocate(country) |>
  head(5)
```

`select()` keeps the columns we want (and drops the rest), `rename()` swaps the name of one column without moving it, and `relocate()` reorders columns without changing names or count. The verbs are *composable* — each does one job, and the pipe lets you stack them in any order. (Order does matter: if you `rename()` first, you must reference the new name in subsequent calls.)

***

### Critical thinking and open exploration

::: {.callout-note collapse="true" title="EX3.32 — show question"}
**EX3.32 (★★)** Run `mean(olympic_athletes$Height)` without `na.rm`.
:::

**Solution** *(reference: Section 3.3 "summarize variables", `na.rm`)*. The result is `NA` — any `NA` in the input poisons the mean. Use `na.rm = TRUE` to ignore them.

```{r 03-ex-32}
mean(olympic_athletes$Height)
mean(olympic_athletes$Height, na.rm = TRUE)
```

***

::: {.callout-note collapse="true" title="EX3.33 — show question"}
**EX3.33 (★★★)** Compare a 3-step chain written with intermediate variables vs `|>`.
:::

**Solution** *(reference: Section 3.1 "The pipe operator |>")*. Piped style is easier to *modify* (insert/swap a step in one line) and easier to *read top-to-bottom*. Intermediate-variable style is easier to *debug step-by-step* (you can `View(step1)` between calls).

```{r 03-ex-33, eval=FALSE}
# Intermediate variables
step1 <- filter(olympic_athletes, Sport == "Athletics")
step2 <- group_by(step1, Year)
step3 <- summarize(step2, mean_h = mean(Height, na.rm = TRUE))

# Piped (same)
olympic_athletes |>
  filter(Sport == "Athletics") |>
  group_by(Year) |>
  summarize(mean_h = mean(Height, na.rm = TRUE))
```

***

::: {.callout-note collapse="true" title="EX3.34 — show question"}
**EX3.34 (★★★)** Two countries (USA & GBR) — total medals across the five recent Games as a single linegraph.
:::

**Solution** *(reference: Sections 3.4 "group_by rows" + 2.4 "Linegraphs")*.

```{r 03-ex-34, fig.height=3}
medal_table |>
  filter(noc %in% c("USA", "GBR")) |>
  group_by(noc, year) |>
  summarize(total = sum(total), .groups = "drop") |>
  ggplot(aes(x = year, y = total, color = noc)) +
  geom_line() + geom_point() +
  labs(y = "Total medals (Summer + Winter)")
```

***

::: {.callout-note collapse="true" title="EX3.35 — show question"}
**EX3.35 (★★★)** Episodes per primary country, then again with `geo_match == "full"`. Does the leader change?
:::

**Solution** *(reference: Section 3.2 "filter rows" + 3.4 "group_by rows")*. Italy is typically the leader in both — the geo-match filter mostly removes "compilation" or multi-country episodes (`geo_match == "country"` etc.).

```{r 03-ex-35}
episodes |> count(primary_country, sort = TRUE) |> slice_head(n = 5)

episodes |>
  filter(geo_match == "full") |>
  count(primary_country, sort = TRUE) |> slice_head(n = 5)
```

***

::: {.callout-note collapse="true" title="EX3.36 — show question"}
**EX3.36 (★★★)** Country diversity per Games — distinct `NOC`s per `Year`/`Season`.
:::

**Solution** *(reference: Section 3.4 "group_by rows", time-series summary)*. The Olympic movement has expanded substantially since 1896.

```{r 03-ex-36, fig.height=3}
olympic_athletes |>
  group_by(Year, Season) |>
  summarize(n_nocs = n_distinct(NOC), .groups = "drop") |>
  ggplot(aes(x = Year, y = n_nocs, color = Season)) +
  geom_line() + geom_point() +
  labs(y = "Distinct NOCs")
```

***

::: {.callout-note collapse="true" title="EX3.37 — show question"}
**EX3.37 (★★★)** Open exploration.
:::

**Solution** *(reference: Chapter 3 — synthesis)*. *Answers will vary.* Example: "Among `Rick Steves' Europe` episodes, *British Isles* episodes have the highest median IMDB rating among regions with at least 10 episodes."

```{r 03-ex-37}
episodes |>
  group_by(region) |>
  summarize(n = n(), median_rating = median(imdb_rating, na.rm = TRUE),
            .groups = "drop") |>
  filter(n >= 10) |>
  arrange(desc(median_rating))
```

***
