Solutions for the end-of-chapter exercises in *Chapter 4 — Data Importing and Tidy Data*. The dataset is `bob_ross` from the [`fivethirtyeight`](https://fivethirtyeight-r.netlify.app/reference/bob_ross.html) package: 403 episodes × 71 columns (4 identifiers + 67 binary indicator columns).

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                                      | # of exercises |
|---------------------------------------------------|---------------:|
| 4.1 Importing data                                 | 2              |
| 4.2 Tidy data                                      | 23             |
| 4.3 Case study: democracy in Guatemala             | 1              |
| 4.4 `tidyverse` package                            | 1              |
| Setup / chapter overview                           | 1              |
| Critical thinking / open exploration               | 4              |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                        | Exercises                                              |
|---------------------------------------------------|--------------------------------------------------------|
| Setup and first look                               | EX4.1                                                  |
| 4.1 Importing data (`read_csv`, type-coercion)     | EX4.2, EX4.3                                           |
| 4.2.1 Definition of tidy data                      | EX4.4, EX4.5, EX4.6, EX4.7, EX4.8, EX4.9, EX4.12       |
| 4.2.2 Converting to tidy data — `pivot_longer()`   | EX4.10, EX4.11, EX4.13                                 |
| 4.2.2 Applied summaries on long-form               | EX4.14, EX4.15, EX4.16, EX4.17, EX4.18, EX4.19, EX4.20, EX4.21 |
| 4.2.2 Applied visualizations on long-form          | EX4.22, EX4.23, EX4.24                                 |
| 4.2.2 Pivoting wider / round-trips                 | EX4.25, EX4.26                                         |
| 4.3 Case study: democracy in Guatemala             | EX4.27                                                 |
| 4.4 `tidyverse` package                            | EX4.28                                                 |
| Critical thinking / open exploration               | EX4.29, EX4.30, EX4.31, EX4.32                         |

**A note on coverage.** Section 4.4 is short (~40 lines) and conceptually straightforward (one umbrella `library()` call vs many) — a single exercise (EX4.28) covers it cleanly. Section 4.3 (the Guatemala case study) is a worked example rather than new methodology; EX4.27 acts as a *comprehension* check on the pivoting structure used there. Exercises now follow chapter content order: setup → importing (4.1) → recognizing tidy structure (4.2.1) → pivoting longer (4.2.2) → applied summaries/viz → pivoting wider → case study (4.3) → tidyverse umbrella (4.4) → critical thinking.

Difficulty stars on each exercise: ★ warm-up, ★★ standard application, ★★★ critical thinking. Click *show question* on any callout to reveal the prompt; the solution is always visible underneath.
:::

```{r 04-ex-setup, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(fivethirtyeight)
```

### Setup and first look

::: {.callout-note collapse="true" title="EX4.1 — show question"}
**EX4.1 (★)** Load `fivethirtyeight` and the tidyverse, then run `glimpse(bob_ross)`. How many rows does the data frame have, and how many columns?
:::

**Solution** *(reference: Chapter 4, "Data Importing and Tidy Data")*. `bob_ross` has **403 rows** (one per episode) and **71 columns** (4 identifiers + 67 indicator columns).

```{r 04-ex-1}
glimpse(bob_ross)
```

***

### Importing data

::: {.callout-note collapse="true" title="EX4.2 — show question"}
**EX4.2 (★)** Suppose someone published a CSV version of the dataset at `https://example.com/bob-ross.csv`. Write the single line of code you would use — with `readr` — to read it into a tibble called `br`.
:::

**Solution** *(reference: Section 4.1 "Importing data")*.

```{r 04-ex-2, eval=FALSE}
br <- read_csv("https://example.com/bob-ross.csv")
```

***

::: {.callout-note collapse="true" title="EX4.3 — show question"}
**EX4.3 (★★)** `read_csv()` reads `age` as character (`"unknown"` mixed in). Two fixes — at-import vs after-import. Why does at-import win?
:::

**Solution** *(reference: Section 4.1 "Importing data", `readr` / `read_csv()`)*.

**(a) At-import fix** — tell `read_csv()` what `"unknown"` means *as you read*:

```r
read_csv("athletes.csv", na = c("", "NA", "unknown"))
```

`readr` then parses `age` as **numeric** automatically (because the `"unknown"` strings became `NA` before type-detection ran), and you don't carry a string-typed column forward.

**(b) After-import fix** — coerce in a `mutate()`:

```r
read_csv("athletes.csv") |>
  mutate(age = na_if(age, "unknown"),
         age = as.numeric(age))
```

This works, but you've already paid the cost of a character-typed column propagating through the pipeline (and silenced the chance to spot *other* sentinel values like `"N/A"` or `"missing"` at the same place).

**Why at-import wins:** the fix lives next to the *cause*. The dataset's "missing" convention is documented at the boundary — anyone reading the script can see exactly which strings got translated to `NA`. Subsequent code can then assume clean numeric data. This is the classic "fail at the boundary, trust the interior" principle.

***

### Recognizing tidy structure

::: {.callout-note collapse="true" title="EX4.4 — show question"}
**EX4.4 (★)** The first four columns of `bob_ross` (`episode`, `season`, `episode_num`, `title`) are *identification* variables, not *measurement* variables. Explain in one sentence why.
:::

**Solution** *(reference: "Definition of tidy data")*. They serve to **uniquely identify** which episode each row refers to (e.g., `S01E01` titled "A Walk in the Woods"); they describe *which* observation we're looking at, not *what* was painted in it.

***

::: {.callout-note collapse="true" title="EX4.5 — show question"}
**EX4.5 (★)** Without running any code, predict: what are the only two possible values that any of the 67 indicator columns (e.g., `mountain`, `tree`) can take? Then verify by running `summary(bob_ross$tree)`.
:::

**Solution** *(reference: "Definition of tidy data")*. The columns are **0/1 indicators** — `1` means the element appears in that episode, `0` means it does not.

```{r 04-ex-5}
summary(bob_ross$tree)
range(bob_ross$tree)
```

***

::: {.callout-note collapse="true" title="EX4.6 — show question"}
**EX4.6 (★)** Is `bob_ross` in tidy format? Justify by referring to the three rules of tidy data.
:::

**Solution** *(reference: "Definition of tidy data")*. **No.** Two of Wickham's three rules are violated:

1. *Each variable is a column.* ✗ — there are really only two variables encoded across the 67 indicator columns: the **element name** (currently a column name) and **whether it appeared** (currently a cell value).
2. *Each observation is a row.* ✗ — the natural observation is "this element in this episode," but right now one row holds 67 such observations.
3. *Each observational unit has its own table.* ✓.

The fix is `pivot_longer()` (see EX4.10).

***

::: {.callout-note collapse="true" title="EX4.7 — show question"}
**EX4.7 (★)** Two variables are encoded in the *names* of the 67 indicator columns. What are they?
:::

**Solution** *(reference: "Definition of tidy data")*. The column **names** encode the variable **`element`** (e.g., `mountain`, `cabin`, `tree`), and the **cell values** encode the variable **`present`** (0 or 1). Both should be their own columns in tidy form.

***

::: {.callout-note collapse="true" title="EX4.8 — show question"}
**EX4.8 (★★)** The `episode` column contains values like `"S01E01"`, `"S01E02"`. How many distinct pieces of information are packed into this single column? Suggest tidier columns to replace it.
:::

**Solution** *(reference: "Definition of tidy data")*. Two: the **season number** and the **episode-within-season number**. They are already separated into the `season` and `episode_num` columns, so `episode` is technically redundant — though useful as a human-readable label. A `tidyr::separate()` call would split `"S01E01"` into the two pieces if the columns weren't already present:

```{r 04-ex-8}
tibble(episode = c("S01E01", "S02E13")) |>
  separate_wider_position(episode, widths = c(s = 1, season = 2, e = 1, ep = 2))
```

***

::: {.callout-note collapse="true" title="EX4.9 — show question"}
**EX4.9 (★)** What is the *observational unit* of `bob_ross`? Phrase your answer as "one row of `bob_ross` represents one ____."
:::

**Solution** *(reference: "Definition of tidy data")*. One row of `bob_ross` represents one **episode** of *The Joy of Painting*.

***

### Pivoting longer

::: {.callout-note collapse="true" title="EX4.10 — show question"}
**EX4.10 (★★)** Use `pivot_longer()` to reshape `bob_ross` so that all 67 indicator columns become two columns: one called `element` (the old name) and one called `present` (the 0/1 value). Save the result to `bob_long`.
:::

**Solution** *(reference: "tidyverse package")*.

```{r 04-ex-10}
bob_long <- bob_ross |>
  pivot_longer(
    cols      = -c(episode, season, episode_num, title),
    names_to  = "element",
    values_to = "present"
  )

bob_long
```

***

::: {.callout-note collapse="true" title="EX4.11 — show question"}
**EX4.11 (★★)** `bob_ross` has 403 rows and 67 indicator columns. **Predict** the number of rows in `bob_long` *before* you check. Then verify with `nrow()`.
:::

**Solution** *(reference: "tidyverse package")*. Predicted: $403 \times 67 = 27{,}001$ rows. Verified:

```{r 04-ex-11}
nrow(bob_long)
403 * 67
```

***

::: {.callout-note collapse="true" title="EX4.12 — show question"}
**EX4.12 (★★)** After pivoting, each row of `bob_long` represents one ___ in one ___. Fill in the blanks.
:::

**Solution** *(reference: "Definition of tidy data")*. Each row represents **one element in one episode** — e.g., "was `mountain` present in `S01E03`?".

***

::: {.callout-note collapse="true" title="EX4.13 — show question"}
**EX4.13 (★★)** Why does using `cols = -c(episode, season, episode_num, title)` work *better* than listing all 67 element columns by name in `pivot_longer()`?
:::

**Solution** *(reference: "tidyverse package")*. *Negative selection* (`-c(...)`) means "everything except these four." It's **shorter**, **less error-prone** (no risk of typos in 67 column names), and **future-proof**: if FiveThirtyEight ever adds a new element column, the code keeps working without changes.

***

### Summarizing the long-form data

::: {.callout-note collapse="true" title="EX4.14 — show question"}
**EX4.14 (★)** Count how many episodes feature a `cabin`. Hint: an indicator column of 0/1 sums to the number of 1s.
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-14}
sum(bob_ross$cabin)
```

Roughly 18% of episodes — cabins are surprisingly common.

***

::: {.callout-note collapse="true" title="EX4.15 — show question"}
**EX4.15 (★★)** Filter `bob_long` to rows where `present == 1` and call the result `bob_present`. How many rows does it have, and what does that count tell you about the show as a whole?
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-15}
bob_present <- bob_long |> filter(present == 1)
nrow(bob_present)
nrow(bob_present) / 403   # average elements per episode
```

The total — the count of element-episode pairs where `present == 1` — divided by 403 episodes gives the **mean number of elements per painting** (≈10).

***

::: {.callout-note collapse="true" title="EX4.16 — show question"}
**EX4.16 (★★)** Using `bob_long` (or `bob_present`), find the **top 10 most-painted elements** across all episodes.
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-16}
bob_long |>
  group_by(element) |>
  summarize(n_episodes = sum(present)) |>
  arrange(desc(n_episodes)) |>
  slice_head(n = 10)
```

***

::: {.callout-note collapse="true" title="EX4.17 — show question"}
**EX4.17 (★★)** How many distinct elements appear in the very first episode of season 1?
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-17}
bob_long |>
  filter(episode == "S01E01", present == 1) |>
  nrow()
```

***

::: {.callout-note collapse="true" title="EX4.18 — show question"}
**EX4.18 (★★)** Compute the *average number of elements per painting* across the full series. Is this larger or smaller than you would have guessed?
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-18}
bob_long |>
  group_by(episode) |>
  summarize(n_elements = sum(present)) |>
  summarize(mean_elements = mean(n_elements))
```

About 10 elements per painting — Bob's "happy little" components add up.

***

::: {.callout-note collapse="true" title="EX4.19 — show question"}
**EX4.19 (★★)** Did Bob Ross ever paint a `barn`? In how many episodes? In what fraction of his run does a barn appear?
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-19}
bob_long |>
  filter(element == "barn") |>
  summarize(
    n_episodes  = sum(present),
    pct_of_show = mean(present)
  )
```

***

::: {.callout-note collapse="true" title="EX4.20 — show question"}
**EX4.20 (★★)** Which episode features the *largest* number of distinct elements? Print its `episode`, `title`, and total element count.
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-20}
bob_long |>
  group_by(episode, title) |>
  summarize(n_elements = sum(present), .groups = "drop") |>
  slice_max(n_elements, n = 1)
```

***

::: {.callout-note collapse="true" title="EX4.21 — show question"}
**EX4.21 (★★★)** For each season, find the **single most-painted element**. Are there elements that dominate one season but not another?
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-21}
bob_long |>
  group_by(season, element) |>
  summarize(n_episodes = sum(present), .groups = "drop_last") |>
  slice_max(n_episodes, n = 1, with_ties = FALSE) |>
  arrange(season)
```

`tree` and `tree_s` (plural) appear at the top across nearly every season — Bob really did love trees.

***

### Visualizing the long-form data

::: {.callout-note collapse="true" title="EX4.22 — show question"}
**EX4.22 (★★)** Build a horizontal bar chart of the top 15 elements by total appearance count across the full series.
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-22, fig.height=4}
top15 <- bob_long |>
  group_by(element) |>
  summarize(n = sum(present)) |>
  slice_max(n, n = 15)

ggplot(top15, aes(x = n, y = reorder(element, n))) +
  geom_col() +
  labs(x = "Episodes featuring element", y = NULL,
       title = "Top 15 elements across The Joy of Painting")
```

***

::: {.callout-note collapse="true" title="EX4.23 — show question"}
**EX4.23 (★★)** Plot the number of episodes per season featuring `tree` as a line graph (season on the x-axis). Was Bob's tree habit consistent?
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-23, fig.height=3}
bob_long |>
  filter(element == "tree") |>
  group_by(season) |>
  summarize(n = sum(present)) |>
  ggplot(aes(x = season, y = n)) +
  geom_line() +
  geom_point() +
  labs(x = "Season", y = "Episodes with at least one tree",
       title = "Bob Ross trees per season")
```

The line is nearly flat near the maximum — trees appeared in almost every episode, every season.

***

::: {.callout-note collapse="true" title="EX4.24 — show question"}
**EX4.24 (★★★)** Build a faceted bar chart showing the **top 5 elements within each season** (one panel per season). Comment on whether the rankings shift over time.
:::

**Solution** *(reference: "Tidy data")*.

```{r 04-ex-24, fig.height=8, fig.width=8}
bob_long |>
  group_by(season, element) |>
  summarize(n = sum(present), .groups = "drop_last") |>
  slice_max(n, n = 5, with_ties = FALSE) |>
  ggplot(aes(x = n, y = reorder(element, n))) +
  geom_col() +
  facet_wrap(~ season, scales = "free_y") +
  labs(x = "Episodes", y = NULL,
       title = "Top 5 elements per season")
```

Trees, mountains, and clouds dominate every season; rarer items (waterfalls, structures) shuffle around the edges.

***

### Pivoting wider

::: {.callout-note collapse="true" title="EX4.25 — show question"}
**EX4.25 (★★★)** Starting from `bob_long`, use `pivot_wider()` to reconstruct a wide data frame and confirm its dimensions match the original `bob_ross`.
:::

**Solution** *(reference: "tidyverse package")*.

```{r 04-ex-25}
bob_wide <- bob_long |>
  pivot_wider(names_from = element, values_from = present)

dim(bob_wide)
dim(bob_ross)
```

Same dimensions (403 × 71). The pivot is a true *inverse* when no information is lost.

***

::: {.callout-note collapse="true" title="EX4.26 — show question"}
**EX4.26 (★★★)** Build a *season × element* count matrix using `pivot_wider()` with `values_fn = sum` so each cell contains the count of episodes that season featuring that element.
:::

**Solution** *(reference: "tidyverse package")*.

```{r 04-ex-26}
season_element_matrix <- bob_long |>
  pivot_wider(
    id_cols     = season,
    names_from  = element,
    values_from = present,
    values_fn   = sum
  )

season_element_matrix
```

***

### Case study: democracy in Guatemala

::: {.callout-note collapse="true" title="EX4.27 — show question"}
**EX4.27 (★★)** `dem_score` (one column per year): name the variable hidden in the column names, sketch the `pivot_longer()` call.
:::

**Solution** *(reference: Section 4.3 "Case study: democracy in Guatemala")*. **(a)** The column names — `1952`, `1957`, `1962`, …, `1992` — encode the variable **`year`**. The values inside those columns encode a second variable, **`democracy_score`**. Two variables hiding in a wide layout — the textbook tidy-data violation.

**(b)** The pivot looks like:

```{r 04-ex-27, message=FALSE}
library(tidyr)

dem_score <- readr::read_csv("https://moderndive.com/data/dem_score.csv")

dem_score |>
  pivot_longer(
    cols      = -country,         # everything except country becomes year/value
    names_to  = "year",
    values_to = "democracy_score"
  ) |>
  head(6)
```

Each row of the result represents **one country in one year** — the right granularity for plotting trajectories or running models.

***

### `tidyverse` package

::: {.callout-note collapse="true" title="EX4.28 — show question"}
**EX4.28 (★)** What is the difference between `library(tidyverse)` and `library(tidyr)`? Which one would suffice for the pivots in this set of exercises, and which is more convenient?
:::

**Solution** *(reference: "tidyverse package")*. `library(tidyr)` loads only the `tidyr` package — which is enough for `pivot_longer()` and `pivot_wider()`. `library(tidyverse)` is the **umbrella** package that loads `tidyr`, `dplyr`, `ggplot2`, `readr`, and several others in a single call. The umbrella is more convenient when you want all the tidyverse tools at once, as we did throughout this exercise set.

***

### Critical thinking and open exploration

::: {.callout-note collapse="true" title="EX4.29 — show question"}
**EX4.29 (★★★)** The wide format is sometimes *more* convenient than the long format. Give one analytical question that is easier to answer in `bob_ross` (wide) than in `bob_long` (long).
:::

**Solution** *(reference: "Tidy data")*. "In how many episodes did Bob paint **both** a `mountain` AND a `cabin`?" In the wide data this is a single filter:

```{r 04-ex-29}
bob_ross |>
  filter(mountain == 1, cabin == 1) |>
  nrow()
```

In the long data you'd need to pivot back to wide, or do a self-join — both more work.

***

::: {.callout-note collapse="true" title="EX4.30 — show question"}
**EX4.30 (★★★)** Pick one element whose frequency *surprises* you (high or low). Write 2–3 sentences interpreting why that frequency might make sense given what you know — or suspect — about the show.
:::

**Solution** *(reference: "Tidy data")*. *Answers will vary.* One worked example: **`aurora_borealis`** appears in only a handful of episodes — far fewer than I'd guessed for a painter known for atmospheric night scenes. Painting an aurora is technically demanding (multiple wet-into-wet color glazes), so Bob may have reserved it for "special" episodes rather than the standard half-hour format.

```{r 04-ex-30}
bob_long |>
  filter(element == "aurora_borealis") |>
  summarize(n = sum(present))
```

***

::: {.callout-note collapse="true" title="EX4.31 — show question"}
**EX4.31 (★★★)** Name **two** questions about Bob Ross's painting style that this dataset *cannot* answer. For each, explain what additional data you would need.
:::

**Solution** *(reference: "Tidy data")*. *Answers will vary.* Examples:

1. *"Which colors did Bob use most often?"* — the dataset records **what objects** appeared, not paint pigments or palette choices. You'd need each episode's full color list.
2. *"How long did each painting take to complete?"* — the dataset has no runtime or stroke-count information. You'd need video timing data or the show's production records.
3. *"Did the public prefer mountain paintings to ocean paintings?"* — there are no popularity, ratings, or sales data tied to individual episodes.

***

::: {.callout-note collapse="true" title="EX4.32 — show question"}
**EX4.32 (★★★)** Open exploration: find one pattern the chapter examples did *not* uncover and create a chart that communicates it clearly.
:::

**Solution** *(reference: "Tidy data")*. *Answers will vary.* One example: **complexity over time** — does the average number of elements per painting trend up or down across seasons?

```{r 04-ex-32, fig.height=3}
bob_long |>
  group_by(season, episode) |>
  summarize(n_elements = sum(present), .groups = "drop_last") |>
  summarize(mean_elements = mean(n_elements)) |>
  ggplot(aes(x = season, y = mean_elements)) +
  geom_line() + geom_point() +
  labs(x = "Season", y = "Mean # elements / painting",
       title = "Did Bob's paintings get more elaborate over time?")
```

The series is fairly flat — Bob's compositional density is remarkably consistent.

***
