Solutions for the end-of-chapter exercises in *Chapter 5 — Basic Regression*. Dataset: `olympic_athletes` from [`olympicAthletes`](https://github.com/moderndive/olympicAthletes).

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                                  | # of exercises |
|-----------------------------------------------|---------------:|
| 5.1 One numerical explanatory variable         | 7              |
| 5.2 One categorical explanatory variable       | 7              |
| 5.3 Related topics                             | 5              |
| Setup / EDA                                    | 5              |
| Residual diagnostics (cross-cuts 5.1.3/5.2.3)  | 6              |
| Critical thinking / synthesis                  | 7              |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                          | Exercises                                          |
|-----------------------------------------------------|----------------------------------------------------|
| Setup and exploratory data analysis                  | EX5.1, EX5.2, EX5.3, EX5.4, EX5.5                  |
| 5.1.1 EDA / 5.1.2 Simple linear regression           | EX5.6, EX5.7, EX5.8, EX5.9                         |
| 5.1.3 Observed/fitted values & residuals             | EX5.10, EX5.11, EX5.12                             |
| 5.2.1 EDA / 5.2.2 Linear regression (categorical)    | EX5.13, EX5.14, EX5.15, EX5.16, EX5.17, EX5.18, EX5.19 |
| Residual diagnostics & model fit (5.1.3 + 5.2.3)     | EX5.20, EX5.21, EX5.22, EX5.23, EX5.24, EX5.25     |
| 5.3.1 Correlation is not necessarily causation       | EX5.26, EX5.27, EX5.28                             |
| 5.3.2 Best-fitting line (least-squares)              | EX5.29                                             |
| 5.3.3 `get_regression_*()` companion functions       | EX5.30                                             |
| Critical thinking / synthesis                        | EX5.31, EX5.32, EX5.33, EX5.34, EX5.35, EX5.36, EX5.37 |

**A note on coverage.** Exercises now follow chapter content order: setup/EDA → 5.1 numerical SLR → 5.1.3 observed/fitted/residuals → 5.2 categorical SLR → residual diagnostics → 5.3.1 correlation/causation → 5.3.2 best-fitting line → 5.3.3 helper functions → critical thinking. The Setup/EDA bucket reviews dplyr+ggplot fundamentals before any modeling — these spill across 5.1.1 and 5.2.1 EDA subsections.

Difficulty stars: ★ warm-up, ★★ standard application, ★★★ critical thinking.
:::

```{r 05-ex-setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(moderndive)
library(olympicAthletes)

bball <- olympic_athletes |>
  filter(Sport == "Basketball", !is.na(Height), !is.na(Weight))
```

### Setup and EDA

::: {.callout-note collapse="true" title="EX5.1 — show question"}
**EX5.1 (★)** Filter to basketball with no NAs and call it `bball`. How many rows?
:::

**Solution** *(reference: Section 5.1.1 "Exploratory data analysis")*.

```{r 05-ex-1}
nrow(bball)
```

***

::: {.callout-note collapse="true" title="EX5.2 — show question"}
**EX5.2 (★)** Summary stats for `Height` and `Weight` in `bball`.
:::

**Solution** *(reference: Section 5.1.1 "Exploratory data analysis")*.

```{r 05-ex-2}
bball |>
  summarize(
    mean_h = mean(Height), sd_h = sd(Height),
    min_h = min(Height),  max_h = max(Height),
    mean_w = mean(Weight), sd_w = sd(Weight),
    min_w = min(Weight),  max_w = max(Weight)
  )
```

***

::: {.callout-note collapse="true" title="EX5.3 — show question"}
**EX5.3 (★)** Scatter Weight vs Height. Describe the pattern.
:::

**Solution** *(reference: Section 5.1.1 "Exploratory data analysis", Section 2.3 "Scatterplots")*. Strong positive linear association — taller basketball athletes weigh more, on average.

```{r 05-ex-3, message=FALSE, fig.height=3}
ggplot(bball, aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE)
```

***

::: {.callout-note collapse="true" title="EX5.4 — show question"}
**EX5.4 (★★)** `cor(bball$Height, bball$Weight)`. Interpret.
:::

**Solution** *(reference: Section 5.1.1 "Exploratory data analysis", correlation)*. A correlation around 0.7–0.8 indicates a strong positive linear relationship — most of the variation in Weight is captured by Height alone.

```{r 05-ex-4}
cor(bball$Height, bball$Weight)
```

***

::: {.callout-note collapse="true" title="EX5.5 — show question"}
**EX5.5 (★★)** Why is the bivariate scatter the right *first step* before fitting a regression line?
:::

**Solution** *(reference: Section 5.1.1 "Exploratory data analysis", EDA-first principle)*. Regression coefficients are summary statistics — they hide outliers, non-linearity, and clustering. A scatterplot lets you *see* the relationship and decide whether a straight line is even appropriate.

***

### Simple linear regression: one numerical predictor

::: {.callout-note collapse="true" title="EX5.6 — show question"}
**EX5.6 (★★)** Fit `model_b <- lm(Weight ~ Height, data = bball)` and print `get_regression_table()`.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression")*.

```{r 05-ex-6}
model_b <- lm(Weight ~ Height, data = bball)
get_regression_table(model_b)
```

***

::: {.callout-note collapse="true" title="EX5.7 — show question"}
**EX5.7 (★★)** Interpret the slope.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", slope interpretation)*. For each additional centimeter of height, predicted weight increases by approximately *β* kilograms (read the slope estimate from your table — typically ~1.0 kg/cm for basketball).

***

::: {.callout-note collapse="true" title="EX5.8 — show question"}
**EX5.8 (★★)** Interpret the intercept. Why is the literal interpretation misleading?
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", extrapolation)*. The intercept is the predicted Weight at Height = 0 cm — far outside the observed range (~150–230 cm). Using it as a literal prediction is **extrapolation**, which the chapter warns against.

***

::: {.callout-note collapse="true" title="EX5.9 — show question"}
**EX5.9 (★★★)** Repeat for `Sport == "Gymnastics"`. Compare slopes.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression")*. Gymnasts' slope is typically *steeper* (kg per cm) — a small height range maps to a larger weight range relative to basketball, where many tall players cluster around average BMI.

```{r 05-ex-9}
gym <- olympic_athletes |>
  filter(Sport == "Gymnastics", !is.na(Height), !is.na(Weight))
model_g <- lm(Weight ~ Height, data = gym)
get_regression_table(model_g)
```

***

### Observed/fitted values and residuals

::: {.callout-note collapse="true" title="EX5.10 — show question"}
**EX5.10 (★★)** First 5 fitted values and residuals.
:::

**Solution** *(reference: Section 5.1.3 "Observed/fitted values and residuals")*.

```{r 05-ex-10}
get_regression_points(model_b) |> slice_head(n = 5)
```

***

::: {.callout-note collapse="true" title="EX5.11 — show question"}
**EX5.11 (★★)** R² from `get_regression_summaries()`. Interpret in context.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", R² interpretation)*. With R² ≈ 0.6, roughly 60% of the variation in basketball-athlete Weight is explained by Height alone — a strong but imperfect predictor.

```{r 05-ex-11}
get_regression_summaries(model_b)
```

***

::: {.callout-note collapse="true" title="EX5.12 — show question"}
**EX5.12 (★★)** Predict weight at Height = 200 cm.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", prediction)*.

```{r 05-ex-12}
coef(model_b)
predict(model_b, newdata = data.frame(Height = 200))
```

***

### Simple linear regression: one categorical predictor

::: {.callout-note collapse="true" title="EX5.13 — show question"}
**EX5.13 (★★)** Boxplots of `Height` by `Sex` on `bball`.
:::

**Solution** *(reference: Section 5.2.1 "Exploratory data analysis")*. Male basketball players are taller on average than female basketball players.

```{r 05-ex-13, fig.height=3}
ggplot(bball, aes(x = Sex, y = Height)) + geom_boxplot()
```

***

::: {.callout-note collapse="true" title="EX5.14 — show question"}
**EX5.14 (★★)** Fit `lm(Height ~ Sex)` on `bball`. Which sex is the baseline?
:::

**Solution** *(reference: Section 5.2.2 "Linear regression")*. R uses **alphabetical order** for factor levels by default, so `F` is the baseline. The coefficient `SexM` is the *difference* M − F.

```{r 05-ex-14}
model_sex <- lm(Height ~ Sex, data = bball)
get_regression_table(model_sex)
```

***

::: {.callout-note collapse="true" title="EX5.15 — show question"}
**EX5.15 (★★)** Interpret the intercept of `model_sex`.
:::

**Solution** *(reference: Section 5.2.2 "Linear regression", baseline interpretation)*. The intercept is the **mean Height of female basketball players** in the data.

***

::: {.callout-note collapse="true" title="EX5.16 — show question"}
**EX5.16 (★★)** Interpret the `SexM` coefficient.
:::

**Solution** *(reference: Section 5.2.2 "Linear regression", offset coefficients)*. `SexM` = *mean Height of male players − mean Height of female players*. So adding it to the intercept gives the male group mean.

***

::: {.callout-note collapse="true" title="EX5.17 — show question"}
**EX5.17 (★★★)** Confirm with group means via `dplyr`.
:::

**Solution** *(reference: Section 5.2.2 "Linear regression", equivalence to group means)*.

```{r 05-ex-17}
bball |>
  group_by(Sex) |>
  summarize(mean_h = mean(Height), .groups = "drop")
coef(model_sex)
```

The difference between the two group means equals the `SexM` coefficient.

***

::: {.callout-note collapse="true" title="EX5.18 — show question"}
**EX5.18 (★★★)** `lm(Height ~ Season)` on the full data. Which season is taller?
:::

**Solution** *(reference: Section 5.2.2 "Linear regression")*. Summer's mean Height is slightly higher (driven by basketball, volleyball, rowing, etc.).

```{r 05-ex-18}
all_h <- olympic_athletes |> filter(!is.na(Height))
get_regression_table(lm(Height ~ Season, data = all_h))
```

***

::: {.callout-note collapse="true" title="EX5.19 — show question"}
**EX5.19 (★★★)** `lm(Height ~ Sport)` on the 5 most-populous sports. Interpret one coefficient.
:::

**Solution** *(reference: Section 5.2.2 "Linear regression", multi-level categorical)*.

```{r 05-ex-19}
top5 <- olympic_athletes |>
  count(Sport, sort = TRUE) |> slice_head(n = 5) |> pull(Sport)

dat <- olympic_athletes |>
  filter(Sport %in% top5, !is.na(Height))
get_regression_table(lm(Height ~ Sport, data = dat))
```

Each non-baseline sport's coefficient is its **mean Height minus the baseline sport's mean Height**.

***

### Residuals & diagnostics

::: {.callout-note collapse="true" title="EX5.20 — show question"}
**EX5.20 (★★★)** Residuals vs fitted plot for `model_b`.
:::

**Solution** *(reference: Section 5.1.3 "Observed/fitted values and residuals", residual diagnostics)*. Look for trumpet shapes (heteroscedasticity) or curvature (mis-specification). The basketball data is mostly clean; modest fan-out is normal at the high-Weight tail.

```{r 05-ex-20, fig.height=3}
get_regression_points(model_b) |>
  ggplot(aes(x = Weight_hat, y = residual)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = 2)
```

***

::: {.callout-note collapse="true" title="EX5.21 — show question"}
**EX5.21 (★★★)** Histogram of residuals from `model_b`.
:::

**Solution** *(reference: Section 5.1.3 "Observed/fitted values and residuals")*. Roughly bell-shaped, centered at 0 — the linear-model assumption is reasonable.

```{r 05-ex-21, fig.height=3}
get_regression_points(model_b) |>
  ggplot(aes(x = residual)) +
  geom_histogram(bins = 30)
```

***

::: {.callout-note collapse="true" title="EX5.22 — show question"}
**EX5.22 (★★★)** Residuals vs Height. Heteroscedasticity?
:::

**Solution** *(reference: Section 5.1.3 "Observed/fitted values and residuals")*. Some fan-out at higher Heights is visible — not extreme, but enough to note. A weighted regression or a transformation could address it.

```{r 05-ex-22, fig.height=3}
get_regression_points(model_b) |>
  ggplot(aes(x = Height, y = residual)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = 2)
```

***

::: {.callout-note collapse="true" title="EX5.23 — show question"}
**EX5.23 (★★)** Plain-English R² sentence.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", R²)*. "Knowing how tall a basketball athlete is gets us about 60% of the way to predicting their weight — the rest depends on factors we haven't measured (build, body composition, sport-specific training)."

***

::: {.callout-note collapse="true" title="EX5.24 — show question"}
**EX5.24 (★★★)** Top-5 outliers in basketball Weight given Height.
:::

**Solution** *(reference: Section 5.1.3 "Observed/fitted values and residuals")*.

```{r 05-ex-24}
get_regression_points(model_b) |>
  arrange(desc(abs(residual))) |>
  slice_head(n = 5)
```

***

::: {.callout-note collapse="true" title="EX5.25 — show question"}
**EX5.25 (★★★)** What evidence shows that one straight line is *not* a good model on the *full* dataset?
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", model assumptions)*. The full scatterplot shows multiple distinct *clusters* (one per sport), each with its own slope and intercept. A single line averages over a structurally heterogeneous population — visible curvature or banding, plus an R² well below the per-sport models.

```{r 05-ex-25, message=FALSE, fig.height=3}
olympic_athletes |>
  filter(!is.na(Height), !is.na(Weight)) |>
  ggplot(aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.02) +
  geom_smooth(method = "lm", se = FALSE)
```

***

### Correlation is not necessarily causation

::: {.callout-note collapse="true" title="EX5.26 — show question"}
**EX5.26 (★★★)** Regress `Height ~ Year` on basketball.
:::

**Solution** *(reference: Section 5.3.1 "Correlation is not necessarily causation")*. A positive slope (cm/year) shows the secular trend.

```{r 05-ex-26}
get_regression_table(lm(Height ~ Year, data = bball))
```

***

::: {.callout-note collapse="true" title="EX5.27 — show question"}
**EX5.27 (★★★)** Refute "Olympic basketball *causes* people to grow taller every year."
:::

**Solution** *(reference: Section 5.3.1 "Correlation is not necessarily causation")*. The slope reflects **selection** — over the years the Olympics has progressively recruited *taller* athletes — not biology. Confounders include rule changes, professional NBA pipeline, and global trends in human height. Regression detects association; causation needs an experiment or a careful identification strategy.

***

::: {.callout-note collapse="true" title="EX5.28 — show question"}
**EX5.28 (★★★)** Simple way to control for sport without multiple regression.
:::

**Solution** *(reference: Section 5.3.1 "Correlation is not necessarily causation", stratification)*. Fit the regression *separately within each sport* (stratification). Each model's slope is sport-specific. The pooled slope mixes within-sport and between-sport variation; the per-sport slopes isolate the within-sport effect.

***

### Best-fitting line

::: {.callout-note collapse="true" title="EX5.29 — show question"}
**EX5.29 (★★)** On a tiny toy dataset, fit `lm()`, compute fitted values, residuals, and SSE. Then check that the lm() slope minimizes SSE.
:::

**Solution** *(reference: Section 5.3.2 "Best-fitting line", least-squares)*.

```{r 05-ex-29}
toy <- data.frame(x = c(1, 2, 3, 4), y = c(2, 4, 5, 9))

m_toy <- lm(y ~ x, data = toy)
coef(m_toy)
```

The `lm()` slope is **2.2** and intercept **−0.5** (rounded). Fitted values, residuals, and SSE:

```{r 05-ex-29b}
toy$y_hat <- predict(m_toy)
toy$resid <- toy$y - toy$y_hat
toy
sse_lm <- sum(toy$resid^2)
sse_lm
```

**Now try a "wrong" slope of 1.5 (keep the lm() intercept):**

```{r 05-ex-29c}
toy$y_hat_alt <- coef(m_toy)[1] + 1.5 * toy$x
sse_alt <- sum((toy$y - toy$y_hat_alt)^2)
sse_alt
```

`sse_alt` (≈8.0) is **larger** than `sse_lm` (≈1.8). **Why must this hold for any slope you try?** Because the least-squares fit is *defined* as the unique line that **minimizes** the sum of squared residuals. Any deviation from the lm() coefficients — slope, intercept, or both — moves you off the minimum, so the SSE can only stay equal (if you happened to land back on lm()'s answer) or grow. That's the entire content of the words "least squares".

***

### `get_regression_*()` companion functions

::: {.callout-note collapse="true" title="EX5.30 — show question"}
**EX5.30 (★★)** What unit of information does each of `get_regression_table()`, `get_regression_points()`, and `get_regression_summaries()` return?
:::

**Solution** *(reference: Section 5.3.3 "`get_regression_*()` functions")*.

```{r 05-ex-30}
model_b <- lm(Weight ~ Height, data = bball)

get_regression_table(model_b)       # one row per coefficient
get_regression_summaries(model_b)   # one row — model-level
head(get_regression_points(model_b), 5)  # one row per observation
```

| Function                       | One row per ____      | What you read off |
|--------------------------------|-----------------------|--------------------|
| `get_regression_table()`        | **coefficient**        | estimate, SE, *t*, *p*-value, CI |
| `get_regression_points()`       | **observation** (row of data) | observed, fitted, residual |
| `get_regression_summaries()`    | **model**              | R², σ̂, # observations, df  |

A useful mnemonic: the *table* tells you about the *parameters*, the *points* tell you about the *data*, the *summaries* tell you about the *fit as a whole*.

***

### Critical thinking and synthesis

::: {.callout-note collapse="true" title="EX5.31 — show question"}
**EX5.31 (★★★)** `Weight ~ Height` for two sports. Compare slopes.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression")*. *Answers vary by choice.* Different slopes ⇒ pooling distorts the relationship. If slopes are similar, pooling gains power without bias.

```{r 05-ex-31}
sports_two <- olympic_athletes |>
  filter(Sport %in% c("Athletics", "Swimming"),
         !is.na(Height), !is.na(Weight))

sports_two |>
  group_by(Sport) |>
  summarize(slope = coef(lm(Weight ~ Height))[2],
            intercept = coef(lm(Weight ~ Height))[1])
```

***

::: {.callout-note collapse="true" title="EX5.32 — show question"}
**EX5.32 (★★★)** Why baseline matters for *interpretation* but not *prediction*.
:::

**Solution** *(reference: Section 5.2.2 "Linear regression", model parametrization)*. Predictions are invariant to which level is baseline — the same fitted means come out either way. But the coefficients have different *names* and *meanings* depending on baseline (intercept = mean of whichever level is baseline; offsets are *differences* from that baseline). Pick a baseline that makes interpretation natural for your audience.

***

::: {.callout-note collapse="true" title="EX5.33 — show question"}
**EX5.33 (★★★)** Two situations where linear regression is a poor choice.
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", model assumptions)*. (1) **Curved relationships** — e.g., diminishing returns or U-shapes; a line systematically misses both ends. (2) **Binary or count outcomes** — `Medal won (Y/N)`; logistic or count regressions are appropriate, not OLS.

***

::: {.callout-note collapse="true" title="EX5.34 — show question"}
**EX5.34 (★★★)** Multi-level categorical regression with interpretation.
:::

**Solution** *(reference: Section 5.2.2 "Linear regression")*. *Answers vary.* Example: `lm(imdb_rating ~ region, data = steves::episodes)` gives one coefficient per non-baseline region; each is the difference in mean IMDB rating from the baseline region.

***

::: {.callout-note collapse="true" title="EX5.35 — show question"}
**EX5.35 (★★)** Why doesn't `lm()` break on `NA`s?
:::

**Solution** *(reference: Section 5.1.2 "Simple linear regression", `na.action`)*. `lm()`'s default is `na.action = na.omit` — it silently drops rows with `NA` in any variable used in the formula. The summary's `n` reflects the kept rows. This is convenient, but you should always check how many rows were dropped.

```{r 05-ex-35}
summary(lm(Weight ~ Height, data = olympic_athletes))$df[2] + 2
```

***

::: {.callout-note collapse="true" title="EX5.36 — show question"}
**EX5.36 (★★★)** Two questions this dataset can't answer with regression alone.
:::

**Solution** *(reference: Chapter 5 — synthesis)*. *Answers vary.* (1) "Does specific *training* cause athletic improvement?" — needs longitudinal training data. (2) "Why has female participation grown?" — needs policy/institutional data outside athlete demographics.

***

::: {.callout-note collapse="true" title="EX5.37 — show question"}
**EX5.37 (★★★)** Open exploration.
:::

**Solution** *(reference: Chapter 5 — synthesis)*. *Answers will vary.* Example: "Among Olympic swimmers, **Height predicts about 40% of variation in Weight, with each cm associated with ~0.9 kg.**"

```{r 05-ex-37}
swim <- olympic_athletes |>
  filter(Sport == "Swimming", !is.na(Height), !is.na(Weight))
get_regression_summaries(lm(Weight ~ Height, data = swim))
```

***
