Solutions for the end-of-chapter exercises in *Chapter 6 — Multiple Regression*. Dataset: `planets` from [`exoplanets`](https://github.com/moderndive/exoplanets).

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                                          | # of exercises |
|-------------------------------------------------------|---------------:|
| 6.1 One numerical + one categorical                    | 10             |
| 6.2 Two numerical explanatory variables                | 7              |
| Setup / EDA (chapter overview)                         | 4              |
| Model comparison / selection (cross-section)           | 3              |
| Critical thinking / synthesis / open exploration       | 12             |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                             | Exercises                                                |
|--------------------------------------------------------|----------------------------------------------------------|
| Setup and chapter overview                              | EX6.1, EX6.2, EX6.3, EX6.4                               |
| 6.1.1 EDA for numerical + categorical                   | EX6.5, EX6.6                                             |
| 6.1.2 Model with interactions                           | EX6.7, EX6.8, EX6.9, EX6.10                              |
| 6.1.3 Model without interactions (parallel slopes)      | EX6.11, EX6.12, EX6.13                                   |
| 6.1.4 Observed/fitted/residuals (interaction model)     | EX6.14                                                   |
| 6.2.1 EDA for two numerical predictors                  | EX6.15                                                   |
| 6.2.2 Multiple regression — two numerical regressors    | EX6.16, EX6.17, EX6.18, EX6.19, EX6.20                   |
| 6.2.3 Observed/fitted/residuals (two-numerical model)   | EX6.21                                                   |
| Model comparison / selection                            | EX6.22, EX6.23, EX6.24                                   |
| Critical thinking / synthesis / open exploration        | EX6.25, EX6.26, EX6.27, EX6.28, EX6.29, EX6.30, EX6.31, EX6.32, EX6.33, EX6.34, EX6.35, EX6.36 |

**A note on coverage.** *Chapter 6 is long (≈1,290 lines)*; the dozen critical-thinking prompts deliberately *cross-cut* the subsections (selection bias, log transforms, R² interpretation, residual standard error). Section 6.1.4 (residual diagnostics for the interaction model) was the one truly absent subsection — EX6.14 fills that gap. Exercises now follow chapter content order: setup → 6.1 (num+cat) → 6.2 (two numerical) → model selection → critical thinking.

Difficulty: ★ warm-up · ★★ standard · ★★★ critical thinking.
:::

```{r 06-ex-setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(moderndive)
library(exoplanets)

planets_lite <- planets |>
  filter(!is.na(radius_earth), !is.na(mass_earth),
         !is.na(discovery_method), !is.na(eq_temp_k)) |>
  mutate(log_mass = log10(mass_earth),
         log_radius = log10(radius_earth))

top2 <- planets_lite |> count(discovery_method, sort = TRUE) |>
  slice_head(n = 2) |> pull(discovery_method)
pl2 <- planets_lite |> filter(discovery_method %in% top2)

planets_temp <- planets |>
  filter(!is.na(eq_temp_k), !is.na(star_temp_k), !is.na(insolation_earth)) |>
  mutate(log_insol = log10(insolation_earth))
```

### Setup and EDA

::: {.callout-note collapse="true" title="EX6.1 — show question"}
**EX6.1 (★)** `glimpse(planets)`. How many rows / cols?
:::

**Solution** *(reference: Chapter 6 overview)*. **6,278 × 28**.

```{r 06-ex-1}
dim(planets)
```

***

::: {.callout-note collapse="true" title="EX6.2 — show question"}
**EX6.2 (★)** Proportion of `NA`s; why is mass missing more often than radius?
:::

**Solution** *(reference: Chapter 6 — EDA)*. Most planets are detected via the **transit method**, which directly yields a *radius* (size of the dip in starlight). Mass requires *radial velocity follow-up*, which is more expensive and not always done.

```{r 06-ex-2}
planets |>
  summarize(across(c(mass_earth, radius_earth, eq_temp_k),
                   ~ mean(is.na(.x))))
```

***

::: {.callout-note collapse="true" title="EX6.3 — show question"}
**EX6.3 (★★)** Build `planets_lite`. Row count?
:::

**Solution** *(reference: Section 6.1.1 "Exploratory data analysis", complete-case filtering)*.

```{r 06-ex-3}
nrow(planets_lite)
```

***

::: {.callout-note collapse="true" title="EX6.4 — show question"}
**EX6.4 (★★)** Mass–radius scatter (log–log). Clusters?
:::

**Solution** *(reference: Section 6.1.1 "Exploratory data analysis")*. Two big clusters separated by a transition near ~2 Earth radii: **rocky/sub-Neptune** worlds vs. **gas giants** dominate. The gap reflects real planetary structure.

```{r 06-ex-4, message=FALSE, fig.height=3}
ggplot(planets_lite, aes(x = radius_earth, y = mass_earth)) +
  geom_point(alpha = 0.3) +
  scale_x_log10() + scale_y_log10()
```

***

### One numerical + one categorical: EDA

::: {.callout-note collapse="true" title="EX6.5 — show question"}
**EX6.5 (★★)** Filter to top-2 discovery methods.
:::

**Solution** *(reference: Section 6.1 "One numerical and one categorical explanatory variable")*.

```{r 06-ex-5}
count(pl2, discovery_method)
```

The two dominant methods in `planets_lite` are **Transit** and **Radial Velocity**.

***

::: {.callout-note collapse="true" title="EX6.6 — show question"}
**EX6.6 (★★)** Mass vs Radius (log–log) colored by method, with per-group regression line.
:::

**Solution** *(reference: Section 6.1.1 "Exploratory data analysis")*.

```{r 06-ex-6, message=FALSE, fig.height=3}
ggplot(pl2, aes(x = radius_earth, y = mass_earth, color = discovery_method)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() + scale_y_log10()
```

***

### Model with interactions

::: {.callout-note collapse="true" title="EX6.7 — show question"}
**EX6.7 (★★★)** Interaction model.
:::

**Solution** *(reference: Section 6.1.2 "Model with interactions")*.

```{r 06-ex-7}
m_int <- lm(log_mass ~ log_radius * discovery_method, data = pl2)
get_regression_table(m_int)
```

***

::: {.callout-note collapse="true" title="EX6.8 — show question"}
**EX6.8 (★★★)** Interpret the interaction coefficient.
:::

**Solution** *(reference: Section 6.1.2 "Model with interactions", interpreting interactions)*. The interaction quantifies how much **steeper or shallower** the radius–mass slope is for the non-baseline method compared to the baseline. A non-zero interaction means the two methods sample populations with *different* mass-radius relations — likely because each method preferentially detects planets in different mass-radius regions.

***

::: {.callout-note collapse="true" title="EX6.9 — show question"}
**EX6.9 (★★★)** Interaction model — diverging lines.
:::

**Solution** *(reference: Section 6.1.2 "Model with interactions")*. **Different slopes**: the two lines fan apart (or together) as `radius_earth` grows.

```{r 06-ex-9, message=FALSE, fig.height=3}
pl2 |>
  mutate(fit_int = predict(m_int)) |>
  ggplot(aes(x = log_radius, y = log_mass, color = discovery_method)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(y = fit_int), linewidth = 1)
```

***

::: {.callout-note collapse="true" title="EX6.10 — show question"}
**EX6.10 (★★★)** Visualize the interaction.
:::

**Solution** *(reference: Section 6.1.2 "Model with interactions")*.

```{r 06-ex-10, message=FALSE, fig.height=3}
pl2 |>
  ggplot(aes(x = radius_earth, y = mass_earth, color = discovery_method)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10() + scale_y_log10() +
  labs(title = "Mass–radius relation differs by detection method",
       x = "Radius (Earth radii)", y = "Mass (Earth masses)")
```

***

### Model without interactions (parallel slopes)

::: {.callout-note collapse="true" title="EX6.11 — show question"}
**EX6.11 (★★★)** Parallel-slopes vs interaction R².
:::

**Solution** *(reference: Section 6.1.3 "Model without interactions")*.

```{r 06-ex-11}
m_par <- lm(log_mass ~ log_radius + discovery_method, data = pl2)
c(parallel    = get_regression_summaries(m_par)$r_squared,
  interaction = get_regression_summaries(m_int)$r_squared)
```

If the gap is small, the interaction is **mostly noise** and parallel slopes is preferred (simpler, easier to interpret).

***

::: {.callout-note collapse="true" title="EX6.12 — show question"}
**EX6.12 (★★★)** When R² barely changes, what's the right choice?
:::

**Solution** *(reference: Section 6.1.4 "Choosing between models", parsimony)*. **Parsimony**: the simpler parallel-slopes model. Equal explanatory power and simpler interpretation — the slope is a *single number* you can quote, rather than two slopes you must contrast.

***

::: {.callout-note collapse="true" title="EX6.13 — show question"}
**EX6.13 (★★)** Parallel-slopes lines visually.
:::

**Solution** *(reference: Section 6.1.3 "Model without interactions", parallel slopes visually)*. **Same slope, shifted intercepts**: the two lines have the same tilt, but one sits above or below the other.

```{r 06-ex-13, message=FALSE, fig.height=3}
pl2 |>
  mutate(fit_par = predict(m_par)) |>
  ggplot(aes(x = log_radius, y = log_mass, color = discovery_method)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(y = fit_par), linewidth = 1)
```

***

### Observed/fitted values and residuals (interaction model)

::: {.callout-note collapse="true" title="EX6.14 — show question"}
**EX6.14 (★★★)** Diagnostics for the interaction model `m_int`: residuals vs fitted by method, plus per-method mean residual.
:::

**Solution** *(reference: Section 6.1.4 "Observed values, fitted values, and residuals" — interaction model)*.

```{r 06-ex-14, message=FALSE, fig.height=3}
pts_int <- get_regression_points(m_int)

# (a) residuals vs fitted, colored by method
pts_int |>
  ggplot(aes(x = log_mass_hat, y = residual,
             color = discovery_method)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x = "Fitted log_mass", y = "Residual",
       color = "Method")

# (b) per-method mean residual — should be ≈ 0
pts_int |>
  group_by(discovery_method) |>
  summarize(mean_resid = mean(residual), .groups = "drop")
```

**(a)** The two clouds **should overlap** — both centered near zero, with similar spread. If you see one method's residuals systematically positive and the other systematically negative, the chosen interaction structure is missing something (e.g., a non-linear relationship within one group).

**(b) Mean residuals are ≈ 0 by construction.** Why? An interaction model fits a *separate intercept and slope per method* — and the OLS first-order conditions force the residuals *within each level of the categorical variable* to sum to zero. Equivalently: each per-method line passes through the per-method means, so the centered residuals must average to zero. (This *fails* for a parallel-slopes model that didn't include the per-method intercepts.)

***

### Two numerical predictors: EDA

::: {.callout-note collapse="true" title="EX6.15 — show question"}
**EX6.15 (★★)** `eq_temp_k` vs `insolation_earth` scatter.
:::

**Solution** *(reference: Section 6.2.1 "Exploratory data analysis")*. The relationship is roughly linear in *log* insolation — physics: $T_\text{eq}^4 \propto \text{insolation}$ ⇒ $\log T_\text{eq} \propto \frac{1}{4} \log \text{insolation}$.

```{r 06-ex-15, fig.height=3}
ggplot(planets_temp, aes(x = insolation_earth, y = eq_temp_k)) +
  geom_point(alpha = 0.3) + scale_x_log10()
```

***

### Multiple regression with two numerical regressors

::: {.callout-note collapse="true" title="EX6.16 — show question"}
**EX6.16 (★★★)** Fit `m_two`.
:::

**Solution** *(reference: Section 6.2.2 "Multiple regression with two numerical regressors")*.

```{r 06-ex-16}
m_two <- lm(eq_temp_k ~ log_insol + star_temp_k, data = planets_temp)
get_regression_table(m_two)
```

***

::: {.callout-note collapse="true" title="EX6.17 — show question"}
**EX6.17 (★★★)** Interpret insolation slope (holding star temp).
:::

**Solution** *(reference: Section 6.2.2, "holding others constant")*. For each **factor-of-10 increase** in insolation (one unit on the log scale), predicted equilibrium temperature rises by *β* Kelvin — *for stars of the same temperature*.

***

::: {.callout-note collapse="true" title="EX6.18 — show question"}
**EX6.18 (★★★)** Interpret star temperature slope (holding insolation).
:::

**Solution** *(reference: Section 6.2.2 "Multiple regression with two numerical regressors")*. For each additional Kelvin of host star temperature, predicted planet equilibrium temperature rises by *γ* Kelvin — *holding incoming flux fixed*. Captures the spectral-shape effect (hotter stars emit more in shorter wavelengths, slightly increasing absorption).

***

::: {.callout-note collapse="true" title="EX6.19 — show question"}
**EX6.19 (★★★)** Why "holding constant" matters in multiple regression but not simple.
:::

**Solution** *(reference: Section 6.2 "Two numerical explanatory variables", partial effect)*. In simple regression there's only one predictor — nothing to hold constant. In multiple regression each coefficient is a *partial* effect: how the outcome changes with that predictor *while the other predictors are fixed*. Skipping the qualifier inflates the apparent effect when predictors are correlated.

***

::: {.callout-note collapse="true" title="EX6.20 — show question"}
**EX6.20 (★★★)** Earth-like prediction.
:::

**Solution** *(reference: Section 6.2.2 "Multiple regression with two numerical regressors", prediction)*. The model's prediction is in the same ballpark as 255 K, but coverage is sample-dependent — most planets in the catalog are *hotter* than Earth (transit selection biases toward close-in, hot planets), so the model's intercept calibration may pull predictions warm.

```{r 06-ex-20}
predict(m_two, newdata = data.frame(log_insol = log10(1), star_temp_k = 5800))
```

***

### Observed/fitted values and residuals (two-numerical model)

::: {.callout-note collapse="true" title="EX6.21 — show question"}
**EX6.21 (★★★)** Residuals vs fitted for `m_two`.
:::

**Solution** *(reference: Section 6.2.3 "Observed/fitted values and residuals")*. Some heteroscedasticity at the high-temp end is expected — hot exoplanets near their stars are over-represented and under-modeled by a two-predictor linear fit.

```{r 06-ex-21, fig.height=3}
get_regression_points(m_two) |>
  ggplot(aes(x = eq_temp_k_hat, y = residual)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = 2)
```

***

### Comparing models / model selection

::: {.callout-note collapse="true" title="EX6.22 — show question"}
**EX6.22 (★★★)** Three R² for `eq_temp_k` models.
:::

**Solution** *(reference: Section 6.2.4 "Choosing between models")*.

```{r 06-ex-22}
m0 <- lm(eq_temp_k ~ 1,                       data = planets_temp)
m1 <- lm(eq_temp_k ~ log_insol,               data = planets_temp)
m2 <- m_two
c(intercept_only = get_regression_summaries(m0)$r_squared,
  one_predictor  = get_regression_summaries(m1)$r_squared,
  two_predictors = get_regression_summaries(m2)$r_squared)
```

The big jump comes from adding `insolation` (physics-driven). `star_temp_k` adds a smaller, secondary contribution.

***

::: {.callout-note collapse="true" title="EX6.23 — show question"}
**EX6.23 (★★★)** Push back on "more predictors are always better."
:::

**Solution** *(reference: Section 6.2.4 "Choosing between models", overfitting + interpretability)*. Each added predictor risks **overfitting** (capturing noise) and **multicollinearity** (entangled coefficients). Simpler models with similar R² are preferred for interpretation and out-of-sample reliability.

***

::: {.callout-note collapse="true" title="EX6.24 — show question"}
**EX6.24 (★★★)** Context-dependent "high" R².
:::

**Solution** *(reference: Section 6.2.4 "Choosing between models")*. R² has no universal threshold. In physics where one variable largely *causes* another, R² > 0.9 is common; in social science with many unmeasured confounders, R² ≈ 0.2 may be impressive. Always read R² in the context of your field's typical signal-to-noise.

***

### Critical thinking and synthesis

::: {.callout-note collapse="true" title="EX6.25 — show question"}
**EX6.25 (★★★)** Method-selection effect on the discovery-method coefficient.
:::

**Solution** *(reference: Section 6.1.2 "Model with interactions", selection bias)*. The coefficient mostly reflects **detection bias**, not intrinsic mass differences in the underlying planet population. We're not comparing apples to apples — each method samples a different mass-radius region of parameter space.

***

::: {.callout-note collapse="true" title="EX6.26 — show question"}
**EX6.26 (★★★)** Two ways selection distorts conclusions.
:::

**Solution** *(reference: Chapter 6 — generalization caveats)*. (1) **Brightness/closeness bias** — only nearby/bright stars get follow-up, so the sample is geometrically skewed. (2) **Mass-radius gap** — small-mass small-radius planets near distant faint stars are systematically *missing*, distorting any fitted slope.

***

::: {.callout-note collapse="true" title="EX6.27 — show question"}
**EX6.27 (★★★)** Re-fit on `planets_neighbors` (`distance_pc < 100`).
:::

**Solution** *(reference: Chapter 6 — robustness check)*. Coefficients should be similar in direction but with different magnitudes, since restricting to nearby stars selects more transit-friendly systems.

```{r 06-ex-27}
planets_neighbors <- planets |>
  filter(distance_pc < 100,
         !is.na(radius_earth), !is.na(mass_earth), !is.na(discovery_method)) |>
  mutate(log_mass = log10(mass_earth), log_radius = log10(radius_earth))
top2_n <- planets_neighbors |> count(discovery_method, sort = TRUE) |>
  slice_head(n = 2) |> pull(discovery_method)
plN <- planets_neighbors |> filter(discovery_method %in% top2_n)
get_regression_table(
  lm(log_mass ~ log_radius + discovery_method, data = plN)
)
```

***

::: {.callout-note collapse="true" title="EX6.28 — show question"}
**EX6.28 (★★★)** Open exploration: numerical outcome + 2 predictors.
:::

**Solution** *(reference: Chapter 6 — synthesis)*. *Answers will vary.* Example: predict `eq_temp_k` from `log10(insolation_earth)` interacted with `discovery_method`. Compare to parallel-slopes; pick the simpler one if R² differences are small.

***

::: {.callout-note collapse="true" title="EX6.29 — show question"}
**EX6.29 (★★)** Why log-transform mass/radius but not temperature?
:::

**Solution** *(reference: Chapter 6 — variable transformations)*. Mass and radius span **many orders of magnitude** (Earth-mass to multi-Jupiter). Logs compress that range, linearize a power-law relationship, and stabilize variance. Temperature spans a single order of magnitude in Kelvin, and its physical relationship is closer to linear in `log(insolation)` and `star_temp_k` *individually*, so transforming the response gains little.

***

::: {.callout-note collapse="true" title="EX6.30 — show question"}
**EX6.30 (★★★)** Biggest limitation of regression on this dataset.
:::

**Solution** *(reference: Chapter 6 — generalization caveats)*. **Selection bias**: confirmed exoplanets are not a random sample of all exoplanets. Regression coefficients describe the *catalog*, not the *true population* — extrapolating to "all exoplanets" is unjustified.

***

::: {.callout-note collapse="true" title="EX6.31 — show question"}
**EX6.31 (★★★)** Slope on a tiny method subset is less trustworthy. Why?
:::

**Solution** *(reference: Chapter 6 — sample size)*. Small *n* means the estimated slope has a large **standard error**; the line could swing wildly with one or two added points. The pooled estimate uses the full dataset and is more stable.

```{r 06-ex-31}
small_methods <- planets_lite |>
  count(discovery_method) |> filter(n < 50)
small_methods
```

***

::: {.callout-note collapse="true" title="EX6.32 — show question"}
**EX6.32 (★★★)** What chart shows mass–radius isn't a single line?
:::

**Solution** *(reference: Section 6.1.1 "Exploratory data analysis")*. A log–log scatter where two distinct populations (rocky vs gas giant) cluster separately, OR a residuals-vs-fitted plot from `lm(log10(mass) ~ log10(radius))` showing a clear curved/banded pattern.

***

::: {.callout-note collapse="true" title="EX6.33 — show question"}
**EX6.33 (★★★)** Slope as exponent in `M = k R^p`.
:::

**Solution** *(reference: Section 6.2.2, log–log linearization)*. Take logs: $\log M = \log k + p \log R$. So `lm(log10(mass) ~ log10(radius))` reads off slope = `p`. A slope near 1 ⇒ approximately constant density (fixed composition); slopes > 1 ⇒ density rising with radius (gas giants); < 1 ⇒ falling.

```{r 06-ex-33}
get_regression_table(lm(log_mass ~ log_radius, data = planets_lite))
```

***

::: {.callout-note collapse="true" title="EX6.34 — show question"}
**EX6.34 (★★★)** RSE of `m_two` in temperature units.
:::

**Solution** *(reference: Section 6.2.3 "Observed/fitted values and residuals", residual SE)*.

```{r 06-ex-34}
sigma_hat <- sqrt(deviance(m_two) / df.residual(m_two))
sigma_hat
```

A typical prediction is off by about ±`r` Kelvin from the true equilibrium temperature — the model is good for trends, not single-planet temperatures.

***

::: {.callout-note collapse="true" title="EX6.35 — show question"}
**EX6.35 (★★★)** Why physics-informed?
:::

**Solution** *(reference: Section 6.2 "Two numerical explanatory variables")*. `log10(insolation_earth)` reflects the **black-body relationship** $T \propto S^{1/4}$; `star_temp_k` captures the **spectral shape** of incident radiation (hotter stars deposit more usable photons per unit flux). The two together approximate basic radiative-equilibrium physics.

***

::: {.callout-note collapse="true" title="EX6.36 — show question"}
**EX6.36 (★★★)** Open exploration #2.
:::

**Solution** *(reference: Chapter 6 — synthesis)*. *Answers vary.* Example: model `orbital_period_days` from `semi_major_axis_au` + `star_mass_solar` — Kepler's third law gives $P^2 \propto a^3 / M_\star$, so `log10(period) ≈ 1.5 log10(a) − 0.5 log10(M_*)`.

```{r 06-ex-36}
keplerian <- planets |>
  filter(!is.na(orbital_period_days), !is.na(semi_major_axis_au),
         !is.na(star_mass_solar)) |>
  mutate(log_period = log10(orbital_period_days),
         log_a      = log10(semi_major_axis_au),
         log_M      = log10(star_mass_solar))
get_regression_table(lm(log_period ~ log_a + log_M, data = keplerian))
```

***
