Solutions for the end-of-chapter exercises in *Chapter 7 — Sampling*. Primary dataset: `volcanoes` from [`volcanoes`](https://github.com/moderndive/volcanoes); secondary: `eruptions` (same package) and `planets` from [`exoplanets`](https://github.com/moderndive/exoplanets).

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                                    | # of exercises |
|-------------------------------------------------|---------------:|
| 7.1 First activity: red balls                    | 12             |
| 7.2 Sampling framework                           | 1              |
| 7.3 The Central Limit Theorem                    | 8              |
| 7.4 Second activity: chocolate-covered almonds   | (mirrored — see note) |
| 7.5 Sampling distribution in other scenarios     | 4              |
| Critical thinking / synthesis / open exploration | 9              |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                              | Exercises                                              |
|---------------------------------------------------------|--------------------------------------------------------|
| 7.1.1 Population statistics ($\mu$, $\sigma$, $p$)       | EX7.1, EX7.2, EX7.3, EX7.4                             |
| 7.1.2 Manual sampling / a single sample                  | EX7.5, EX7.6, EX7.7                                    |
| 7.1.3 Virtual sampling / sampling distribution           | EX7.8, EX7.9, EX7.10, EX7.11, EX7.12                   |
| 7.2 Sampling framework (terminology and notation)        | EX7.13                                                 |
| 7.3 Central Limit Theorem (sample mean)                  | EX7.14, EX7.15, EX7.16, EX7.17                         |
| 7.3.4 Sampling distribution of the sample proportion     | EX7.18, EX7.19, EX7.20, EX7.21                         |
| 7.5 Other populations (one-sample mean — eruptions, planets) | EX7.22, EX7.23, EX7.24                            |
| 7.5.1 Sampling distribution for two samples (difference) | EX7.25                                                 |
| Critical thinking / synthesis / open exploration         | EX7.26, EX7.27, EX7.28, EX7.29, EX7.30, EX7.31, EX7.32, EX7.33, EX7.34 |

**A note on coverage.** *Section 7.4 (the chocolate-almond activity for the sample mean) appears 0 here, but the volcano-elevation exercises EX7.5–EX7.17 deliberately mirror the same structure on a different population* — single sample, virtual replicates, sampling distribution, CLT — so 7.4 is *implicitly* exercised throughout the chapter. EX7.13 makes the **terminology** of Section 7.2 (parameters vs. statistics, populations vs. random variables) explicit, and EX7.25 covers the **two-sample** subsection (7.5.1). Exercises now follow chapter content order: setup → manual sampling → virtual sampling → sampling framework → CLT → proportion → other populations → two-sample → critical thinking.

Difficulty: ★ warm-up · ★★ standard · ★★★ critical thinking.
:::

```{r 07-ex-setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(moderndive)
library(infer)
library(volcanoes)

mu     <- mean(volcanoes$elevation, na.rm = TRUE)
sigma  <- sd(volcanoes$elevation, na.rm = TRUE)
p_strato <- mean(volcanoes$primary_volcano_type == "Stratovolcano",
                 na.rm = TRUE)
```

### Setup and population statistics

::: {.callout-note collapse="true" title="EX7.1 — show question"}
**EX7.1 (★)** Glimpse `volcanoes`. Treat as the *population*.
:::

**Solution** *(reference: Section 7.1 "First activity: red balls", population concept)*. ~1,200 rows × 13 columns; this is our population for the rest of the chapter.

```{r 07-ex-1}
dim(volcanoes)
```

***

::: {.callout-note collapse="true" title="EX7.2 — show question"}
**EX7.2 (★)** Population $\mu$ and $\sigma$ of elevation.
:::

**Solution** *(reference: Section 7.4.3 "The center of the distribution: the expected value")*.

```{r 07-ex-2}
c(mu = mu, sigma = sigma)
```

***

::: {.callout-note collapse="true" title="EX7.3 — show question"}
**EX7.3 (★)** Population proportion `Stratovolcano`.
:::

**Solution** *(reference: Section 7.1.1 "The proportion of red balls in the bowl", population proportion)*.

```{r 07-ex-3}
p_strato
```

About 60–65% of Holocene volcanoes are stratovolcanoes (the iconic conical type).

***

::: {.callout-note collapse="true" title="EX7.4 — show question"}
**EX7.4 (★★)** Histogram of elevation — shape?
:::

**Solution** *(reference: Section 2.5 "Histograms", connecting EDA to sampling)*. **Right-skewed** — most volcanoes are at modest elevations, with a long tail of high-altitude stratovolcanoes (Andes, Japan).

```{r 07-ex-4, fig.height=3}
ggplot(volcanoes, aes(x = elevation)) +
  geom_histogram(bins = 30) +
  labs(x = "Elevation (m)")
```

***

### A single sample

::: {.callout-note collapse="true" title="EX7.5 — show question"}
**EX7.5 (★★)** A single $n = 30$ sample.
:::

**Solution** *(reference: Section 7.1.2 "Manual sampling")*.

```{r 07-ex-5}
set.seed(76)
volcanoes |>
  rep_sample_n(size = 30, reps = 1) |>
  summarize(xbar = mean(elevation, na.rm = TRUE))
```

The sample mean is *near* but not exactly equal to $\mu$.

***

::: {.callout-note collapse="true" title="EX7.6 — show question"}
**EX7.6 (★★)** Different seed → different sample → different mean. Why?
:::

**Solution** *(reference: Section 7.4.4 "Sampling variation", concept of sampling variation)*. Each sample is a *random subset*; different draws yield different sample means. This is **sampling variation** — and it's what motivates the sampling distribution.

```{r 07-ex-6}
set.seed(2026)
volcanoes |>
  rep_sample_n(size = 30, reps = 1) |>
  summarize(xbar = mean(elevation, na.rm = TRUE))
```

***

::: {.callout-note collapse="true" title="EX7.7 — show question"}
**EX7.7 (★)** "$\bar{x} \neq \mu$ ⇒ biased sample" — refute.
:::

**Solution** *(reference: Section 7.2 "Sampling framework", bias vs variance)*. Disagree. **Bias** is a *systematic* tendency — present when the sampling *procedure* favors certain values. A single $\bar{x}$ deviating from $\mu$ is *sampling variation*, not bias. With many random samples, the mean of $\bar{x}$ values *does* equal $\mu$ — even though any one sample's mean does not.

***

### Building the sampling distribution

::: {.callout-note collapse="true" title="EX7.8 — show question"}
**EX7.8 (★★)** 1,000 samples × $n=30$, plot histogram.
:::

**Solution** *(reference: Section 7.1.3 "Virtual sampling")*.

```{r 07-ex-8, fig.height=3}
set.seed(76)
samp_dist_30 <- volcanoes |>
  rep_sample_n(size = 30, reps = 1000) |>
  group_by(replicate) |>
  summarize(xbar = mean(elevation, na.rm = TRUE), .groups = "drop")

ggplot(samp_dist_30, aes(x = xbar)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = mu, color = "red", linetype = 2) +
  labs(x = expression(bar(x)), title = "Sampling distribution, n = 30")
```

***

::: {.callout-note collapse="true" title="EX7.9 — show question"}
**EX7.9 (★★)** Center and SE.
:::

**Solution** *(reference: Section 7.4.4 "Sampling variation: standard deviation and standard error")*. Center ≈ $\mu$ (CLT); SD of $\bar{x}$ is the **standard error**.

```{r 07-ex-9}
samp_dist_30 |>
  summarize(center = mean(xbar), SE = sd(xbar), mu = mu)
```

***

::: {.callout-note collapse="true" title="EX7.10 — show question"}
**EX7.10 (★★)** Compare sampling dists for $n = 30$ vs $n = 100$.
:::

**Solution** *(reference: Section 7.4.4 "Sampling variation: standard deviation and standard error", $n$ effect)*. The $n=100$ distribution is **narrower** — bigger samples give more precise estimates.

```{r 07-ex-10, fig.height=3}
set.seed(76)
samp_dist_100 <- volcanoes |>
  rep_sample_n(size = 100, reps = 1000) |>
  group_by(replicate) |>
  summarize(xbar = mean(elevation, na.rm = TRUE), .groups = "drop")

bind_rows(
  samp_dist_30  |> mutate(n = "n = 30"),
  samp_dist_100 |> mutate(n = "n = 100")
) |>
  ggplot(aes(x = xbar)) +
  geom_histogram(bins = 30) +
  facet_wrap(~ n, ncol = 1)
```

***

::: {.callout-note collapse="true" title="EX7.11 — show question"}
**EX7.11 (★★)** SE shrinkage factor; compare to $\sqrt{30/100}$.
:::

**Solution** *(reference: Section 7.4 "Central Limit Theorem", CLT scaling)*.

```{r 07-ex-11}
SE_30  <- sd(samp_dist_30$xbar)
SE_100 <- sd(samp_dist_100$xbar)
c(SE_30 = SE_30, SE_100 = SE_100,
  ratio = SE_100 / SE_30,
  theory = sqrt(30 / 100))
```

The simulated ratio matches $\sqrt{n_1/n_2}$ — exactly the CLT prediction.

***

::: {.callout-note collapse="true" title="EX7.12 — show question"}
**EX7.12 (★★★)** SE vs $n$ on log–log; slope ≈ $-1/2$.
:::

**Solution** *(reference: Section 7.4 "Central Limit Theorem")*.

```{r 07-ex-12, fig.height=3}
set.seed(76)
ns <- c(10, 30, 100, 300)
ses <- sapply(ns, function(nn) {
  volcanoes |>
    rep_sample_n(size = nn, reps = 500) |>
    group_by(replicate) |>
    summarize(xbar = mean(elevation, na.rm = TRUE), .groups = "drop") |>
    summarize(SE = sd(xbar)) |>
    pull(SE)
})

ggplot(data.frame(n = ns, SE = ses), aes(x = n, y = SE)) +
  geom_point() + geom_line() +
  scale_x_log10() + scale_y_log10() +
  labs(title = "log SE vs log n: slope ≈ −1/2 (CLT)")
```

***

### Sampling framework

::: {.callout-note collapse="true" title="EX7.13 — show question"}
**EX7.13 (★)** Match terms to "fixed number" vs "random variable" + state the symbol.
:::

**Solution** *(reference: Section 7.2 "Sampling framework", terminology and notation)*.

| Term                                                | Type                                            | Symbol          |
|-----------------------------------------------------|-------------------------------------------------|-----------------|
| (a) Population mean of elevations                    | **fixed number** (a *parameter*)                 | $\mu$           |
| (b) Sample mean from a fresh draw of 30 volcanoes    | **random variable** (a *statistic*)              | $\bar{x}$       |
| (c) Population proportion of stratovolcanoes         | **fixed number** (a *parameter*)                 | $p$             |
| (d) Sample proportion from a fresh draw              | **random variable** (a *statistic*)              | $\hat{p}$       |
| (e) Standard error of the sample mean                | **fixed number** *for a given $n$* — it is the *standard deviation of the random variable in (b)* | $\text{SE}(\bar{x}) = \sigma/\sqrt{n}$ |

The key dichotomy: **populations have parameters (Greek letters, fixed); samples have statistics (Roman letters with hats or bars, random across draws).** The sampling distribution is the *probability distribution of those random statistics* across all possible samples.

***

### Central Limit Theorem

::: {.callout-note collapse="true" title="EX7.14 — show question"}
**EX7.14 (★★★)** Skewed population, near-normal sampling distribution. Why?
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem")*. The **CLT** says: regardless of the population's shape, the sampling distribution of $\bar{x}$ approaches normality as $n$ grows. By $n = 30$, even a fairly skewed population yields a roughly bell-shaped sampling distribution.

***

::: {.callout-note collapse="true" title="EX7.15 — show question"}
**EX7.15 (★★★)** Population vs sampling distribution side-by-side.
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem")*. The population is right-skewed; the sampling distribution of $\bar{x}$ is roughly bell-shaped — the CLT in action.

```{r 07-ex-15, fig.height=4, message=FALSE}
library(patchwork)

p_pop <- ggplot(volcanoes, aes(x = elevation)) +
  geom_histogram(bins = 30) + labs(title = "Population")

p_sd  <- ggplot(samp_dist_30, aes(x = xbar)) +
  geom_histogram(bins = 30) + labs(title = "Sampling dist of x̄, n = 30")

p_pop / p_sd
```

***

::: {.callout-note collapse="true" title="EX7.16 — show question"}
**EX7.16 (★★★)** Theoretical SE = $\sigma / \sqrt{n}$ vs simulated.
:::

**Solution** *(reference: Section 7.4.4 "Sampling variation: standard deviation and standard error")*.

```{r 07-ex-16}
c(theoretical = sigma / sqrt(30),
  simulated   = sd(samp_dist_30$xbar))
```

Match within simulation noise.

***

::: {.callout-note collapse="true" title="EX7.17 — show question"}
**EX7.17 (★★★)** Smallest $n$ that already looks bell-shaped.
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem", convergence)*. *Answers vary.* For elevation (modestly skewed), even $n = 10$ looks roughly bell-shaped at the center, with thicker tails. By $n = 30$ the bell is clean.

```{r 07-ex-17, fig.height=3}
set.seed(76)
sd_5 <- volcanoes |>
  rep_sample_n(size = 5, reps = 1000) |>
  group_by(replicate) |>
  summarize(xbar = mean(elevation, na.rm = TRUE), .groups = "drop")

ggplot(sd_5, aes(x = xbar)) + geom_histogram(bins = 30) +
  labs(title = "n = 5 — clearly skewed still")
```

***

### Sampling distribution for a proportion

::: {.callout-note collapse="true" title="EX7.18 — show question"}
**EX7.18 (★★)** Sampling dist of $\hat{p}$ for stratovolcano, $n = 50$.
:::

**Solution** *(reference: Section 7.1.3 "Virtual sampling", proportion variant)*.

```{r 07-ex-18, fig.height=3}
set.seed(76)
samp_p <- volcanoes |>
  filter(!is.na(primary_volcano_type)) |>
  rep_sample_n(size = 50, reps = 1000) |>
  group_by(replicate) |>
  summarize(phat = mean(primary_volcano_type == "Stratovolcano"),
            .groups = "drop")

ggplot(samp_p, aes(x = phat)) +
  geom_histogram(bins = 25) +
  geom_vline(xintercept = p_strato, color = "red", linetype = 2)
```

***

::: {.callout-note collapse="true" title="EX7.19 — show question"}
**EX7.19 (★★)** Simulated vs theoretical SE for $\hat{p}$.
:::

**Solution** *(reference: Section 7.4.4 "Sampling variation: standard deviation and standard error", $\sqrt{p(1-p)/n}$)*.

```{r 07-ex-19}
c(simulated   = sd(samp_p$phat),
  theoretical = sqrt(p_strato * (1 - p_strato) / 50))
```

***

::: {.callout-note collapse="true" title="EX7.20 — show question"}
**EX7.20 (★★★)** What if $n = 5$?
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem", small-$n$ failure)*. With $n = 5$, $\hat{p}$ can only take values $\{0, 0.2, 0.4, 0.6, 0.8, 1\}$ — the histogram has visible spikes (no longer continuous-looking) and the CLT approximation is poor.

```{r 07-ex-20, fig.height=3}
set.seed(76)
samp_p_5 <- volcanoes |>
  filter(!is.na(primary_volcano_type)) |>
  rep_sample_n(size = 5, reps = 1000) |>
  group_by(replicate) |>
  summarize(phat = mean(primary_volcano_type == "Stratovolcano"),
            .groups = "drop")
ggplot(samp_p_5, aes(x = phat)) + geom_histogram(bins = 6)
```

***

::: {.callout-note collapse="true" title="EX7.21 — show question"}
**EX7.21 (★★★)** Rare type — Caldera — at $n = 50$.
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem", small $p$ regime)*. When $p$ is near 0 or 1, even moderately large $n$ gives a skewed sampling distribution — the CLT requires *both* large $n$ AND $np(1-p)$ not too small.

```{r 07-ex-21, fig.height=3}
set.seed(76)
samp_p_caldera <- volcanoes |>
  filter(!is.na(primary_volcano_type)) |>
  rep_sample_n(size = 50, reps = 1000) |>
  group_by(replicate) |>
  summarize(phat = mean(primary_volcano_type == "Caldera"),
            .groups = "drop")
ggplot(samp_p_caldera, aes(x = phat)) + geom_histogram(bins = 25)
```

***

### Other populations

::: {.callout-note collapse="true" title="EX7.22 — show question"}
**EX7.22 (★★★)** Population mean `vei`.
:::

**Solution** *(reference: Section 7.1.1 "The proportion of red balls in the bowl", different population)*.

```{r 07-ex-22}
mean(eruptions$vei, na.rm = TRUE)
```

***

::: {.callout-note collapse="true" title="EX7.23 — show question"}
**EX7.23 (★★★)** Sampling dist of mean `vei`, $n = 50$.
:::

**Solution** *(reference: Section 7.1.3 "Virtual sampling")*.

```{r 07-ex-23, fig.height=3}
set.seed(76)
samp_vei <- eruptions |>
  filter(!is.na(vei)) |>
  rep_sample_n(size = 50, reps = 1000) |>
  group_by(replicate) |>
  summarize(xbar = mean(vei, na.rm = TRUE), .groups = "drop")

ggplot(samp_vei, aes(x = xbar)) + geom_histogram(bins = 30)
sd(samp_vei$xbar)
```

***

::: {.callout-note collapse="true" title="EX7.24 — show question"}
**EX7.24 (★★★)** Sampling dist of `radius_earth` for two $n$ values.
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem", $n$ effect on heavy-skew populations)*. The exoplanet radius distribution has *much* heavier tails than volcano elevations, so $n = 30$ may still look skewed. By $n = 200$, it's much closer to bell-shaped.

```{r 07-ex-24, fig.height=4, message=FALSE}
library(exoplanets)
set.seed(76)

samp_30  <- planets |> filter(!is.na(radius_earth)) |>
  rep_sample_n(30,  reps = 500) |>
  group_by(replicate) |>
  summarize(xbar = mean(radius_earth), .groups = "drop") |>
  mutate(n = "n = 30")
samp_200 <- planets |> filter(!is.na(radius_earth)) |>
  rep_sample_n(200, reps = 500) |>
  group_by(replicate) |>
  summarize(xbar = mean(radius_earth), .groups = "drop") |>
  mutate(n = "n = 200")

bind_rows(samp_30, samp_200) |>
  ggplot(aes(x = xbar)) + geom_histogram(bins = 30) +
  facet_wrap(~ n, scales = "free_y", ncol = 1)
```

***

### Sampling distribution for two samples

::: {.callout-note collapse="true" title="EX7.25 — show question"}
**EX7.25 (★★★)** Sampling distribution of the difference $\bar{x}_{SA} - \bar{x}_{MCA}$, $n=30$ each, 1,000 replicates.
:::

**Solution** *(reference: Section 7.5.1 "Sampling distribution for two samples — difference in sample means")*.

```{r 07-ex-25, message=FALSE, fig.height=3}
set.seed(76)
sa  <- volcanoes |> filter(region == "South America Volcanic Regions")
mca <- volcanoes |> filter(region == "Middle America-Caribbean Volcanic Regions")

reps <- 1000

draw_diff <- function() {
  mean(sample(sa$elevation,  30, replace = TRUE), na.rm = TRUE) -
  mean(sample(mca$elevation, 30, replace = TRUE), na.rm = TRUE)
}

diffs <- replicate(reps, draw_diff())

ggplot(data.frame(diff = diffs), aes(x = diff)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = mean(sa$elevation, na.rm = TRUE) -
                          mean(mca$elevation, na.rm = TRUE),
             color = "red", linetype = 2) +
  labs(x = "Sample mean elevation: South America − Mexico/C. America")
```

**(a) Center.** The histogram's center estimates the *population* difference $\mu_{SA} - \mu_{MCA}$ — the fixed truth your simulation is hovering around. (The red dashed line marks that population value.)

**(b) Spread.** The variance of a difference of *independent* sample means equals the **sum** of the individual variances:

$$\text{Var}(\bar{x}_{SA} - \bar{x}_{MCA}) = \tfrac{\sigma_{SA}^2}{n} + \tfrac{\sigma_{MCA}^2}{n}$$

so the SE of the *difference* is **larger** than either single-region SE — typically by a factor of $\sqrt{2}$ when the two region SDs are comparable. Combining two independent estimators adds noise; the CLT still applies, so the histogram is approximately normal.

***

### Critical thinking and synthesis

::: {.callout-note collapse="true" title="EX7.26 — show question"}
**EX7.26 (★★★)** Biased sampling design.
:::

**Solution** *(reference: Section 7.2 "Sampling framework", representative samples)*. **No** — over-sampling tall volcanoes systematically pushes $\bar{x}$ above $\mu$. This is **bias from non-representative sampling**, distinct from *random* sampling variation. No amount of replication fixes a biased procedure.

***

::: {.callout-note collapse="true" title="EX7.27 — show question"}
**EX7.27 (★★★)** SD vs SE.
:::

**Solution** *(reference: Section 7.4.4 "Sampling variation: standard deviation and standard error", terminology)*. **SD** = "spread of *individuals* in the population." Example: SD of volcano elevations = how spread-out individual volcano altitudes are. **SE** = "spread of an *estimator* over hypothetical resamples." Example: SE of $\bar{x}$ for $n = 30$ = how much the sample mean would vary if we redrew the sample many times.

***

::: {.callout-note collapse="true" title="EX7.28 — show question"}
**EX7.28 (★★★)** "n = 30 always works" — critique.
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem", scope of CLT)*. The CLT requires *enough* sample size relative to the population's skewness. For symmetric populations $n = 5$ may suffice; for highly skewed populations or rare-event proportions, $n = 30$ is far too small. Always verify with a sampling-distribution simulation when in doubt.

***

::: {.callout-note collapse="true" title="EX7.29 — show question"}
**EX7.29 (★★★)** SE halves ⇒ $n$ × ?
:::

**Solution** *(reference: Section 7.4.4 "Sampling variation: standard deviation and standard error", $1/\sqrt{n}$ scaling)*. Halving SE requires $n$ to **quadruple** ($\sqrt{4} = 2$). Doubling SE precision quadruples sampling cost — diminishing returns.

***

::: {.callout-note collapse="true" title="EX7.30 — show question"}
**EX7.30 (★★★)** Replicates vs $n$.
:::

**Solution** *(reference: Section 7.1.3 "Virtual sampling", reps vs n)*. Increasing **reps** smooths the *histogram* of the sampling distribution but does **not** change the SE — the SE is a property of the population and $n$. Increasing **n** narrows the sampling distribution itself (smaller SE), regardless of how many replicates we use.

***

::: {.callout-note collapse="true" title="EX7.31 — show question"}
**EX7.31 (★★★)** With vs without replacement, small $n$ vs population.
:::

**Solution** *(reference: Section 7.1.3 "Virtual sampling", finite-population correction)*. With $N \approx 1{,}200$ and $n = 30$ ($n / N \approx 2.5\%$), the difference between with-replacement and without-replacement sampling is negligible. The finite-population correction $\sqrt{(N-n)/(N-1)}$ is only ~0.987.

```{r 07-ex-31}
set.seed(76)
sd_wo <- volcanoes |>
  rep_sample_n(30, reps = 1000, replace = FALSE) |>
  group_by(replicate) |> summarize(xbar = mean(elevation, na.rm = TRUE), .groups = "drop")
sd_wr <- volcanoes |>
  rep_sample_n(30, reps = 1000, replace = TRUE) |>
  group_by(replicate) |> summarize(xbar = mean(elevation, na.rm = TRUE), .groups = "drop")
c(SE_without_repl = sd(sd_wo$xbar), SE_with_repl = sd(sd_wr$xbar))
```

***

::: {.callout-note collapse="true" title="EX7.32 — show question"}
**EX7.32 (★★★)** Whole-chapter visualization.
:::

**Solution** *(reference: Chapter 7 — synthesis)*. *Answers vary.* A 4-panel figure: (1) population histogram with $\mu$ marked; (2) one sample's histogram; (3) sampling distribution with $\mu$ marked; (4) SE-vs-$n$ on log-log.

```{r 07-ex-32, fig.height=3}
ggplot(samp_dist_30, aes(x = xbar)) +
  geom_histogram(bins = 30, fill = "steelblue") +
  geom_vline(xintercept = mu, color = "red", linewidth = 1) +
  labs(title = "Sampling distribution centers on μ — that's the CLT",
       x = expression(bar(x)), y = NULL)
```

***

::: {.callout-note collapse="true" title="EX7.33 — show question"}
**EX7.33 (★★★)** Sampling dist of the *median*.
:::

**Solution** *(reference: Section 7.4 "The Central Limit Theorem", non-mean estimators)*. The median's sampling distribution is also approximately normal under regularity conditions, but with a **larger SE** than the mean (typically by a factor ≈ 1.25 for normal-ish populations).

```{r 07-ex-33, fig.height=3}
set.seed(76)
samp_med <- volcanoes |>
  rep_sample_n(30, reps = 1000) |>
  group_by(replicate) |>
  summarize(med = median(elevation, na.rm = TRUE), .groups = "drop")
ggplot(samp_med, aes(x = med)) + geom_histogram(bins = 25)
sd(samp_med$med)
```

***

::: {.callout-note collapse="true" title="EX7.34 — show question"}
**EX7.34 (★★★)** Open exploration.
:::

**Solution** *(reference: Chapter 7 — synthesis)*. *Answers vary.* Example: "For `eruptions$vei`, the population is heavily right-skewed, but the sampling distribution of $\bar{x}$ at $n = 50$ is bell-shaped with simulated SE ≈ theoretical $\sigma/\sqrt{n}$ — the CLT works even on a count variable."

```{r 07-ex-34}
c(theoretical_SE = sd(eruptions$vei, na.rm = TRUE) / sqrt(50),
  simulated_SE   = sd(samp_vei$xbar))
```

***
