Solutions for the end-of-chapter exercises in *Chapter 9 — Hypothesis Testing*. Datasets: `bob_ross` (`fivethirtyeight`), `volcanoes` and `eruptions` (`volcanoes`), `olympic_athletes` (`olympicAthletes`).

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                                                | # of exercises |
|-------------------------------------------------------------|---------------:|
| 9.1 Tying CIs to hypothesis testing (incl. one-sample mean)  | 9              |
| 9.2 Music popularity activity (permutation / two-sample)     | 8              |
| 9.3 Understanding hypothesis tests                           | 1              |
| 9.4 Conducting hypothesis tests (`infer` workflow)           | 4              |
| 9.5 Interpreting hypothesis tests (errors, alpha)            | 3              |
| 9.6 Case study (mirrored)                                    | (mirrored — see note) |
| Misuses / critical thinking (cross-section)                  | 8              |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                                  | Exercises                                                    |
|-------------------------------------------------------------|--------------------------------------------------------------|
| 9.1.1 One-sample test for a mean                             | EX9.1, EX9.2, EX9.3                                          |
| One-sample test for a proportion (parallel to 9.1.1)         | EX9.4, EX9.5, EX9.6, EX9.7                                   |
| 9.1.2 Hypothesis tests and confidence intervals              | EX9.8, EX9.9                                                 |
| 9.2 Music popularity (permutation, two-sample mean)          | EX9.10, EX9.11, EX9.12, EX9.13, EX9.14                       |
| Two-sample proportion test (parallel to 9.2)                 | EX9.15, EX9.16, EX9.17                                       |
| 9.3 Understanding hypothesis tests (vocabulary)              | EX9.18                                                       |
| 9.4.1 `infer` workflow practice                              | EX9.19, EX9.20                                               |
| 9.4.3 "There is only one test" / theory-vs-simulation        | EX9.21, EX9.22                                               |
| 9.5.2 Types of errors (Type I / Type II)                     | EX9.23, EX9.24, EX9.25                                       |
| Misuses / critical thinking                                  | EX9.26, EX9.27, EX9.28, EX9.29, EX9.30, EX9.31, EX9.32, EX9.33 |

**A note on coverage.** *Section 9.6 — the IMDb action-vs-romance case study — appears 0 here, but it is **structurally** mirrored by EX9.10–EX9.14 (a complete two-sample permutation workflow on Olympic athletes' Height by Sex)*; running EX9.10 and writing it up in the case-study narrative form is the natural way to exercise 9.6's pedagogy. EX9.18 gives Section 9.3 — the *short conceptual* section that introduces the vocabulary of hypothesis testing — a dedicated comprehension prompt. Exercises now follow chapter content order: one-sample mean → one-sample proportion → CI/HT linking → two-sample (mean, proportion) → vocabulary → infer workflow → errors → critical thinking.

Difficulty: ★ warm-up · ★★ standard · ★★★ critical thinking.
:::

```{r 09-ex-setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(infer)
library(moderndive)
library(volcanoes)
library(olympicAthletes)
library(fivethirtyeight)
library(tidyr)
```

### One-sample mean test

::: {.callout-note collapse="true" title="EX9.1 — show question"}
**EX9.1 (★★)** Two-sided test of $\mu = 1500$ on `volcanoes$elevation`.
:::

**Solution** *(reference: Section 9.1.1 "One-sample hypothesis test for the population mean")*.

```{r 09-ex-1}
set.seed(76)
volc_clean <- volcanoes |> filter(!is.na(elevation))

null_mean <- volc_clean |>
  specify(response = elevation) |>
  hypothesize(null = "point", mu = 1500) |>
  generate(reps = 500, type = "bootstrap") |>
  calculate(stat = "mean")

obs_mean <- mean(volc_clean$elevation)
get_p_value(null_mean, obs_stat = obs_mean, direction = "two-sided")
```

***

::: {.callout-note collapse="true" title="EX9.2 — show question"}
**EX9.2 (★★)** Observed sample mean and p-value.
:::

**Solution** *(reference: Section 9.1.1 "One-sample hypothesis test for the population mean")*.

```{r 09-ex-2}
obs_mean
```

***

::: {.callout-note collapse="true" title="EX9.3 — show question"}
**EX9.3 (★★★)** One-sided $\mu > 1500$.
:::

**Solution** *(reference: Section 9.1.1 "One-sample hypothesis test for the population mean")*. The one-sided p-value is **half** the two-sided one (when the observed statistic falls on the side of the alternative).

```{r 09-ex-3}
get_p_value(null_mean, obs_stat = obs_mean, direction = "greater")
```

***

### One-sample proportion test

::: {.callout-note collapse="true" title="EX9.4 — show question"}
**EX9.4 (★★)** Tree proportion test, $H_0: p = 0.5$ vs $H_A: p > 0.5$.
:::

**Solution** *(reference: Section 9.5.1 "infer package workflow")*. The observed proportion is well above 0.5; the null distribution at p = 0.5 sits *much* lower; p-value is essentially 0.

```{r 09-ex-4}
set.seed(76)
br_tree <- bob_ross |>
  mutate(has_tree = if_else(tree == 1, "yes", "no"))

null_dist <- br_tree |>
  specify(response = has_tree, success = "yes") |>
  hypothesize(null = "point", p = 0.5) |>
  generate(reps = 1000, type = "draw") |>
  calculate(stat = "prop")

obs_p <- mean(bob_ross$tree == 1)
obs_p
get_p_value(null_dist, obs_stat = obs_p, direction = "greater")
```

***

::: {.callout-note collapse="true" title="EX9.5 — show question"}
**EX9.5 (★★)** Visualize null + p-value shading.
:::

**Solution** *(reference: Section 9.5.1 "infer package workflow", `shade_p_value()`)*.

```{r 09-ex-5, fig.height=3}
visualize(null_dist) +
  shade_p_value(obs_stat = obs_p, direction = "greater")
```

***

::: {.callout-note collapse="true" title="EX9.6 — show question"}
**EX9.6 (★★)** Interpret in context.
:::

**Solution** *(reference: Section 9.5.1 "infer package workflow", p-value interpretation)*. **If** Bob really painted trees in only half of episodes, **then** the chance of observing a sample with as many or more tree-episodes as we saw is essentially zero. We reject $H_0$ — Bob painted trees in *significantly more* than half of episodes.

***

::: {.callout-note collapse="true" title="EX9.7 — show question"}
**EX9.7 (★★★)** Repeat for mountain, cabin, barn.
:::

**Solution** *(reference: Section 9.5.1 "infer package workflow")*. Mountain ≈ 60%+ → reject. Cabin ≈ 18% → fail to reject "more than half". Barn small → fail to reject.

```{r 09-ex-7}
test_one <- function(col_name) {
  obs <- mean(bob_ross[[col_name]] == 1)
  bob_ross |>
    mutate(has = if_else(.data[[col_name]] == 1, "yes", "no")) |>
    specify(response = has, success = "yes") |>
    hypothesize(null = "point", p = 0.5) |>
    generate(reps = 500, type = "draw") |>
    calculate(stat = "prop") |>
    get_p_value(obs_stat = obs, direction = "greater") |>
    mutate(element = col_name, observed = obs)
}

set.seed(76)
purrr::map_dfr(c("mountain", "cabin", "barn"), test_one)
```

***

### Linking CIs and tests

::: {.callout-note collapse="true" title="EX9.8 — show question"}
**EX9.8 (★★★)** 95% CI for $\mu_M - \mu_F$ Athletics Height.
:::

**Solution** *(reference: Section 9.1.2 "Hypothesis tests and confidence intervals")*. CI well above 0 ⇒ same conclusion as the test (reject $H_0$).

```{r 09-ex-8}
set.seed(76)
ath <- olympic_athletes |>
  filter(Sport == "Athletics", !is.na(Height))
ath |>
  specify(response = Height, explanatory = Sex) |>
  generate(reps = 500, type = "bootstrap") |>
  calculate(stat = "diff in means", order = c("M", "F")) |>
  get_confidence_interval(level = 0.95, type = "percentile")
```

***

::: {.callout-note collapse="true" title="EX9.9 — show question"}
**EX9.9 (★★★)** Why CI ↔ test equivalence.
:::

**Solution** *(reference: Section 9.1.2 "Hypothesis tests and confidence intervals")*. A two-sided test rejects iff the null value is outside the corresponding CI — both procedures use the same standard error and critical value. They differ only in *framing*: a CI tells you the *range* of plausible values; a test tells you whether *one specific* null value is plausible.

***

### Two-sample mean test (music popularity, permutation)

::: {.callout-note collapse="true" title="EX9.10 — show question"}
**EX9.10 (★★★)** Permutation test of $\mu_M = \mu_F$ on Athletics Height.
:::

**Solution** *(reference: Section 9.2 "Music popularity activity")*.

```{r 09-ex-10}
set.seed(76)
null_diff <- ath |>
  specify(response = Height, explanatory = Sex) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "diff in means", order = c("M", "F"))

obs_diff <- ath |>
  specify(response = Height, explanatory = Sex) |>
  calculate(stat = "diff in means", order = c("M", "F"))

get_p_value(null_diff, obs_stat = obs_diff, direction = "greater")
```

***

::: {.callout-note collapse="true" title="EX9.11 — show question"}
**EX9.11 (★★)** Visualize.
:::

**Solution** *(reference: Section 9.5.1 "infer package workflow", `shade_p_value()`)*.

```{r 09-ex-11, fig.height=3}
visualize(null_diff) +
  shade_p_value(obs_stat = obs_diff, direction = "greater")
```

***

::: {.callout-note collapse="true" title="EX9.12 — show question"}
**EX9.12 (★★★)** Interpret.
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests")*. The observed M−F difference is far in the right tail of the permutation null — extremely inconsistent with random shuffling. We reject $H_0$ and conclude male athletics athletes are taller on average.

***

::: {.callout-note collapse="true" title="EX9.13 — show question"}
**EX9.13 (★★★)** Repeat for `Weight`.
:::

**Solution** *(reference: Section 9.2 "Music popularity activity")*. The Weight test will likewise reject — the M-vs-F gap is at least as large in weight terms.

```{r 09-ex-13}
set.seed(76)
ath_w <- olympic_athletes |>
  filter(Sport == "Athletics", !is.na(Weight))
null_w <- ath_w |>
  specify(response = Weight, explanatory = Sex) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "diff in means", order = c("M", "F"))
obs_w <- ath_w |>
  specify(response = Weight, explanatory = Sex) |>
  calculate(stat = "diff in means", order = c("M", "F"))
get_p_value(null_w, obs_stat = obs_w, direction = "greater")
```

***

::: {.callout-note collapse="true" title="EX9.14 — show question"}
**EX9.14 (★★★)** Mean VEI Asia vs South America.
:::

**Solution** *(reference: Section 9.2 "Music popularity activity")*.

```{r 09-ex-14}
set.seed(76)
joined <- eruptions |>
  inner_join(volcanoes |> select(volcano_number, region),
             by = "volcano_number") |>
  filter(region %in% c("Eastern Asia Volcanic Regions",
                       "South America Volcanic Regions"),
         !is.na(vei)) |>
  mutate(region = factor(region))

null_v <- joined |>
  specify(response = vei, explanatory = region) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "diff in means",
            order = levels(joined$region))
obs_v <- joined |>
  specify(response = vei, explanatory = region) |>
  calculate(stat = "diff in means",
            order = levels(joined$region))
get_p_value(null_v, obs_stat = obs_v, direction = "two-sided")
```

***

### Two-proportion test

::: {.callout-note collapse="true" title="EX9.15 — show question"}
**EX9.15 (★★★)** Mountain vs cabin proportion.
:::

**Solution** *(reference: Section 9.2 "Music popularity activity", two-proportion variant)*.

```{r 09-ex-15}
set.seed(76)
br_long <- bob_ross |>
  select(episode, mountain, cabin) |>
  tidyr::pivot_longer(c(mountain, cabin),
                      names_to = "element", values_to = "present") |>
  mutate(present = if_else(present == 1, "yes", "no"))

null_p <- br_long |>
  specify(response = present, explanatory = element, success = "yes") |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "diff in props", order = c("mountain", "cabin"))

obs_p_diff <- br_long |>
  specify(response = present, explanatory = element, success = "yes") |>
  calculate(stat = "diff in props", order = c("mountain", "cabin"))

get_p_value(null_p, obs_stat = obs_p_diff, direction = "two-sided")
```

***

::: {.callout-note collapse="true" title="EX9.16 — show question"}
**EX9.16 (★★★)** What can / can't we conclude?
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests")*. **Can:** reject the null that mountain and cabin are equally common. **Cannot:** infer *why* — Bob may simply have preferred mountain scenes; we have no causal evidence.

***

::: {.callout-note collapse="true" title="EX9.17 — show question"}
**EX9.17 (★★★)** Independence cabin vs barn.
:::

**Solution** *(reference: Section 9.2 "Music popularity activity", independence test)*.

```{r 09-ex-17}
set.seed(76)
cb <- bob_ross |>
  mutate(has_cabin = if_else(cabin == 1, "yes", "no"),
         has_barn  = if_else(barn  == 1, "yes", "no"))

null_cb <- cb |>
  specify(response = has_cabin, explanatory = has_barn, success = "yes") |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "diff in props", order = c("yes", "no"))

obs_cb <- cb |>
  specify(response = has_cabin, explanatory = has_barn, success = "yes") |>
  calculate(stat = "diff in props", order = c("yes", "no"))

get_p_value(null_cb, obs_stat = obs_cb, direction = "two-sided")
```

***

### Understanding hypothesis tests (vocabulary)

::: {.callout-note collapse="true" title="EX9.18 — show question"}
**EX9.18 (★★)** Define and order the six core terms of a hypothesis test (null, alternative, statistic, null distribution, p-value, decision).
:::

**Solution** *(reference: Section 9.3 "Understanding hypothesis tests")*.

| Term | One-sentence definition |
|------|--------------------------|
| (a) **Null hypothesis $H_0$**  | The "boring" claim — no effect, no difference, the status quo. |
| (b) **Alternative hypothesis $H_A$** | The claim you'd accept if the data are inconsistent with $H_0$ — "there *is* an effect". |
| (c) **Test statistic** | A single number computed from the data that summarizes the evidence (e.g., $\bar{x}_1 - \bar{x}_2$, $\hat{p}$). |
| (d) **Null distribution** | The distribution the test statistic would follow *if $H_0$ were true* — generated by simulation (permutations / bootstrap / draws) or by theory. |
| (e) **p-value** | The probability, *under $H_0$*, of seeing a test statistic at least as extreme as the one you observed. |
| (f) **Decision** | At a chosen $\alpha$, either *reject $H_0$* (data inconsistent with the null) or *fail to reject* (insufficient evidence). |

**Order in a typical workflow:** (a) state $H_0$ → (b) state $H_A$ → (c) compute the *observed* test statistic → (d) build the *null distribution* of that statistic → (e) read off the *p-value* by locating the observed statistic on the null distribution → (f) make a *decision* by comparing the p-value to $\alpha$.

A handy mnemonic: **"Hypothesize, Calculate, Distribute, P-value, Decide."** The whole chapter — proportions, means, two-sample, permutation, theory-based — is the same six-step skeleton with different *test statistics* and different *null-distribution generators* plugged in. That's why Section 9.4.3 says **"there is only one test."**

***

### Conducting hypothesis tests (`infer` workflow)

::: {.callout-note collapse="true" title="EX9.19 — show question"}
**EX9.19 (★★★)** p-value vs $n$.
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests", $n$-sensitivity)*. Larger $n$ ⇒ smaller p-value, *holding the effect size fixed*. With enough data, even tiny effects become "statistically significant" — which is why **practical significance** must be checked separately.

```{r 09-ex-19}
set.seed(76)
test_n <- function(n_sub) {
  sub <- bob_ross |> slice_sample(n = n_sub) |>
    mutate(has_tree = if_else(tree == 1, "yes", "no"))
  obs <- mean(sub$tree == 1)
  null_d <- sub |>
    specify(response = has_tree, success = "yes") |>
    hypothesize(null = "point", p = 0.5) |>
    generate(reps = 200, type = "draw") |>
    calculate(stat = "prop")
  get_p_value(null_d, obs_stat = obs, direction = "greater") |>
    mutate(n = n_sub)
}
purrr::map_dfr(c(50, 200, 400), test_n)
```

***

::: {.callout-note collapse="true" title="EX9.20 — show question"}
**EX9.20 (★★★)** Visualize null + observed for any test.
:::

**Solution** *(reference: Section 9.5.1 "infer package workflow", `shade_p_value()`)*.

```{r 09-ex-20, fig.height=3}
visualize(null_diff) +
  shade_p_value(obs_stat = obs_diff, direction = "greater") +
  labs(title = "Athletics Height: M − F vs permutation null")
```

***

::: {.callout-note collapse="true" title="EX9.21 — show question"}
**EX9.21 (★★★)** "There is only one test."
:::

**Solution** *(reference: Section 9.5.2 "There is only one test")*. Every test follows the same template: (1) compute an observed statistic, (2) build the null distribution under $H_0$, (3) compare. The **only difference** between tests is the choice of statistic and the choice of null-generation procedure (permutation, bootstrap, theory).

***

::: {.callout-note collapse="true" title="EX9.22 — show question"}
**EX9.22 (★★★)** Theory-based t-test.
:::

**Solution** *(reference: Section 9.7.1 "Theory-based approach for two-sample hypothesis tests")*.

```{r 09-ex-22}
t.test(Height ~ Sex, data = ath, alternative = "less")  # F < M
```

The simulation-based and theory-based p-values agree closely with $n$ in the thousands.

***

### Errors and alpha

::: {.callout-note collapse="true" title="EX9.23 — show question"}
**EX9.23 (★★★)** Type I and Type II errors in EX9.10.
:::

**Solution** *(reference: Section 9.6.2 "Types of errors")*. **Type I**: declaring "M is taller than F" when in truth $\mu_M = \mu_F$. **Type II**: failing to declare a difference when in truth $\mu_M \neq \mu_F$.

***

::: {.callout-note collapse="true" title="EX9.24 — show question"}
**EX9.24 (★★★)** $\alpha = 0.001$ vs $0.05$.
:::

**Solution** *(reference: Section 9.6.3 "How do we choose alpha?")*. A stricter $\alpha$ requires more extreme evidence to reject ⇒ fewer **false positives** (Type I) but the same data is more likely to fall short of the threshold ⇒ more **false negatives** (Type II). Always a tradeoff.

***

::: {.callout-note collapse="true" title="EX9.25 — show question"}
**EX9.25 (★★★)** "p = 0.03 ⇒ definitely true" — push back.
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests")*. p = 0.03 means the data are *unusual* under $H_0$, not that $H_A$ is "definitely true." It is evidence *against* $H_0$, not proof. Failing to consider sample size, multiple testing, or model assumptions can mislead this interpretation.

***

### Misuses and critical thinking

::: {.callout-note collapse="true" title="EX9.26 — show question"}
**EX9.26 (★★★)** "p = 0.06 ⇒ no effect" — refute.
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests")*. p = 0.06 means **failure to reject** $H_0$ at $\alpha = 0.05$ — *not* evidence that the effect is zero. The effect could be real but underpowered, or close to but just under the threshold. "Absence of evidence is not evidence of absence."

***

::: {.callout-note collapse="true" title="EX9.27 — show question"}
**EX9.27 (★★★)** Multiple testing.
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests", multiple comparisons)*. Expected number of "significant" results under all-true-nulls = $20 \times 0.05 = 1$. Apply a **multiple-comparisons correction** like Bonferroni ($\alpha / m$) or Benjamini-Hochberg.

***

::: {.callout-note collapse="true" title="EX9.28 — show question"}
**EX9.28 (★★★)** Full statistical conclusion for Athletics test.
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests", complete write-up)*. **$H_0$**: $\mu_M = \mu_F$. **$H_A$**: $\mu_M > \mu_F$. **Statistic**: difference in sample mean Heights, ~`r round(coef(lm(Height ~ Sex, ath))[2], 1)` cm. **p-value**: ~0 (extreme right tail). **Decision at $\alpha = 0.05$**: reject $H_0$. **Plain-English**: Olympic male athletics athletes are taller on average than female ones; this conclusion is highly unlikely to be due to chance.

***

::: {.callout-note collapse="true" title="EX9.29 — show question"}
**EX9.29 (★★★)** When testing is unnecessary.
:::

**Solution** *(reference: Section 9.7 "When inference is not needed")*. If your data **are** the population (a census), then sample statistics *are* the parameters — no inference needed. Testing $\mu = 1500$ on the entire `volcanoes` table is misleading: you already *know* $\mu$ exactly.

***

::: {.callout-note collapse="true" title="EX9.30 — show question"}
**EX9.30 (★★)** Statistical vs practical significance.
:::

**Solution** *(reference: Section 9.6 "Interpreting hypothesis tests", practical significance)*. With $n = 100{,}000$ volcanoes, a *1 m* difference in mean elevation might be statistically significant (tiny p-value) yet **practically irrelevant** for any geological question. Significance and meaningfulness are different.

***

::: {.callout-note collapse="true" title="EX9.31 — show question"}
**EX9.31 (★★★)** Theory vs simulation disagreement.
:::

**Solution** *(reference: Section 9.7.1 "Theory-based approach for two-sample hypothesis tests")*. They diverge with **small $n$**, **heavy-tailed populations**, or **violated assumptions**. They agree closely with large $n$ and a roughly normal sampling distribution — exactly when the CLT applies.

***

::: {.callout-note collapse="true" title="EX9.32 — show question"}
**EX9.32 (★★★)** Open exploration.
:::

**Solution** *(reference: Chapter 9 — synthesis)*. *Answers vary.* Example: "Are *Italian* episodes of *Rick Steves' Europe* rated higher than the rest? p ≈ 0.08 — borderline; can't confidently reject at $\alpha = 0.05$."

```{r 09-ex-32}
set.seed(76)
ep_it <- steves::episodes |>
  filter(!is.na(imdb_rating)) |>
  mutate(group = if_else(primary_country == "Italy", "italy", "other"))
null_it <- ep_it |>
  specify(response = imdb_rating, explanatory = group) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "diff in means", order = c("italy", "other"))
obs_it <- ep_it |>
  specify(response = imdb_rating, explanatory = group) |>
  calculate(stat = "diff in means", order = c("italy", "other"))
get_p_value(null_it, obs_stat = obs_it, direction = "two-sided")
```

***

::: {.callout-note collapse="true" title="EX9.33 — show question"}
**EX9.33 (★★★)** More-eruptions for taller volcanoes.
:::

**Solution** *(reference: Section 9.2 "Music popularity activity")*. Median-split volcanoes by elevation, count eruptions per volcano, then permutation-test the difference in mean eruption count.

```{r 09-ex-33}
set.seed(76)
elev_med <- median(volcanoes$elevation, na.rm = TRUE)

erupt_counts <- eruptions |>
  count(volcano_number, name = "n_eruptions") |>
  inner_join(volcanoes |> select(volcano_number, elevation),
             by = "volcano_number") |>
  filter(!is.na(elevation)) |>
  mutate(group = if_else(elevation > elev_med, "tall", "short"))

null_e <- erupt_counts |>
  specify(response = n_eruptions, explanatory = group) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "diff in means", order = c("tall", "short"))

obs_e <- erupt_counts |>
  specify(response = n_eruptions, explanatory = group) |>
  calculate(stat = "diff in means", order = c("tall", "short"))

get_p_value(null_e, obs_stat = obs_e, direction = "greater")
```

***
