Solutions for the end-of-chapter exercises in *Chapter 10 — Inference for Regression*. Datasets: `olympic_athletes` (`olympicAthletes`) and `planets` (`exoplanets`).

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                                                | # of exercises |
|-------------------------------------------------------------|---------------:|
| 10.2 Theory-based inference for simple linear regression     | 7              |
| 10.3 Simulation-based inference for simple linear regression | 5              |
| 10.5 Theory-based inference for multiple linear regression   | 4              |
| 10.6 Simulation-based inference for multiple linear regression | 2            |
| Visualizing uncertainty (cross-section)                      | 3              |
| Practical / interpretation (cross-section)                   | 4              |
| Critical thinking / synthesis                                | 11             |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                                | Exercises                                                    |
|-----------------------------------------------------------|--------------------------------------------------------------|
| 10.2.5 The regression table in R                           | EX10.1                                                       |
| 10.2.3 Confidence intervals for LS estimators              | EX10.2                                                       |
| 10.2.4 Hypothesis test for population slope                | EX10.3, EX10.4                                               |
| 10.2.6 Model fit and assumptions (LINE)                    | EX10.5, EX10.6, EX10.7                                       |
| 10.3 Simulation-based — bootstrap CI / permutation test    | EX10.8, EX10.9, EX10.10, EX10.11, EX10.12                    |
| 10.5 Theory-based for multiple regression                  | EX10.13, EX10.14, EX10.15, EX10.16                           |
| 10.6 Simulation-based for multiple regression               | EX10.17, EX10.18                                             |
| Visualizing uncertainty (CI band, prediction band)         | EX10.19, EX10.20, EX10.21                                    |
| Practical / interpretation (cross-section)                 | EX10.22, EX10.23, EX10.24, EX10.25                           |
| Critical thinking / synthesis                              | EX10.26, EX10.27, EX10.28, EX10.29, EX10.30, EX10.31, EX10.32, EX10.33, EX10.34, EX10.35, EX10.36 |

**A note on coverage.** *Sections 10.1 (the simple-LR model) and 10.4 (the multiple-LR model) appear with 0 exercise rows above*: they are **conceptual setups** for the inference machinery that follows, and EX10.25 ("What's *the population* in regression here?") plus the practical/interpretation block effectively quizzes those setup ideas. Exercises now follow chapter content order: theory-based simple LR → diagnostics → simulation-based simple LR → theory-based multiple LR → simulation-based multiple LR → visualizing uncertainty → practical interpretation → critical thinking.

Difficulty: ★ warm-up · ★★ standard · ★★★ critical thinking.
:::

```{r 10-ex-setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(infer)
library(moderndive)
library(broom)
library(olympicAthletes)
library(exoplanets)

bball <- olympic_athletes |>
  filter(Sport == "Basketball", !is.na(Height), !is.na(Weight))
m_b <- lm(Weight ~ Height, data = bball)

planets_lite <- planets |>
  filter(!is.na(radius_earth), !is.na(mass_earth), !is.na(discovery_method)) |>
  mutate(log_mass = log10(mass_earth), log_radius = log10(radius_earth))
top2 <- planets_lite |> count(discovery_method, sort = TRUE) |>
  slice_head(n = 2) |> pull(discovery_method)
pl2 <- planets_lite |> filter(discovery_method %in% top2)
```

### Theory-based inference for simple regression

::: {.callout-note collapse="true" title="EX10.1 — show question"}
**EX10.1 (★★)** Basketball regression table.
:::

**Solution** *(reference: Section 10.2.5 "The regression table in R")*.

```{r 10-ex-1}
get_regression_table(m_b)
```

***

::: {.callout-note collapse="true" title="EX10.2 — show question"}
**EX10.2 (★★)** Interpret slope CI.
:::

**Solution** *(reference: Section 10.2.3 "Confidence intervals for the least-squares estimators")*. We are 95% confident that, for each additional centimeter of height, the *population* mean Weight of Olympic basketball athletes increases by between [lower] and [upper] kilograms.

***

::: {.callout-note collapse="true" title="EX10.3 — show question"}
**EX10.3 (★★★)** What does the slope's p-value test?
:::

**Solution** *(reference: Section 10.2.4 "Hypothesis test for population slope")*. **$H_0$: population slope = 0** — i.e., Height has no linear relationship with Weight in the population. Tiny p-value ⇒ reject; basketball Height *does* predict Weight.

***

::: {.callout-note collapse="true" title="EX10.4 — show question"}
**EX10.4 (★★★)** t = estimate / SE.
:::

**Solution** *(reference: Section 10.2.4 "Hypothesis test for population slope", t-statistic)*.

```{r 10-ex-4}
tab <- get_regression_table(m_b)
tab
tab$estimate[2] / tab$std_error[2]
```

The computed ratio matches `statistic` in the table.

***

### Model fit and assumptions

::: {.callout-note collapse="true" title="EX10.5 — show question"}
**EX10.5 (★★)** L.I.N.E. assumptions.
:::

**Solution** *(reference: Section 10.2.6 "Model fit and model assumptions")*. **L**inearity of the relationship, **I**ndependence of residuals, **N**ormality of residuals, **E**qual variance (homoscedasticity).

***

::: {.callout-note collapse="true" title="EX10.6 — show question"}
**EX10.6 (★★★)** Diagnostic plots.
:::

**Solution** *(reference: Section 10.2.6 "Model fit and model assumptions", residual diagnostics)*. Look for systematic curvature (linearity), Q-Q deviations (normality), and trumpet shapes (equal variance). Olympic basketball is mostly clean.

```{r 10-ex-6, fig.height=4}
par(mfrow = c(2, 2)); plot(m_b); par(mfrow = c(1, 1))
```

***

::: {.callout-note collapse="true" title="EX10.7 — show question"}
**EX10.7 (★★★)** "Trust t over bootstrap" — push back.
:::

**Solution** *(reference: Section 10.3 "Simulation-based inference for simple linear regression")*. The t-test relies on assumptions (L.I.N.E.) that may not hold. The bootstrap is **assumption-light** — it just resamples the data — making it more reliable when residuals are skewed, when $n$ is small, or when nonlinear structure violates the t-test's premises.

***

### Simulation-based inference for simple regression

::: {.callout-note collapse="true" title="EX10.8 — show question"}
**EX10.8 (★★★)** Bootstrap distribution of slope.
:::

**Solution** *(reference: Section 10.3.1 "Confidence intervals for the population slope using infer")*.

```{r 10-ex-8, fig.height=3}
set.seed(76)
slope_boot <- bball |>
  specify(Weight ~ Height) |>
  generate(reps = 500, type = "bootstrap") |>
  calculate(stat = "slope")
visualize(slope_boot)
```

***

::: {.callout-note collapse="true" title="EX10.9 — show question"}
**EX10.9 (★★★)** Bootstrap percentile 95% CI; compare theory.
:::

**Solution** *(reference: Section 10.3.1 "Confidence intervals for the population slope using infer", method comparison)*.

```{r 10-ex-9}
slope_boot |> get_confidence_interval(level = 0.95, type = "percentile")
get_regression_table(m_b)
```

The bootstrap CI matches the theory-based CI very closely (large $n$, well-behaved residuals).

***

::: {.callout-note collapse="true" title="EX10.10 — show question"}
**EX10.10 (★★★)** Permutation test of slope = 0.
:::

**Solution** *(reference: Section 10.3.2 "Hypothesis test for population slope using infer")*.

```{r 10-ex-10}
set.seed(76)
null_slope <- bball |>
  specify(Weight ~ Height) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  calculate(stat = "slope")
obs_slope <- bball |>
  specify(Weight ~ Height) |>
  calculate(stat = "slope")
get_p_value(null_slope, obs_stat = obs_slope, direction = "two-sided")
```

***

::: {.callout-note collapse="true" title="EX10.11 — show question"}
**EX10.11 (★★★)** Visualize null + p-value.
:::

**Solution** *(reference: Section 10.3.2 "Hypothesis test for population slope using infer")*.

```{r 10-ex-11, fig.height=3}
visualize(null_slope) +
  shade_p_value(obs_stat = obs_slope, direction = "two-sided")
```

***

::: {.callout-note collapse="true" title="EX10.12 — show question"}
**EX10.12 (★★★)** Why simulation and theory agree.
:::

**Solution** *(reference: Section 10.3 "Simulation-based inference for simple linear regression")*. With $n$ in the thousands and reasonably normal residuals, both procedures converge on the same answer — the CLT applies to the sampling distribution of the slope estimator.

***

### Theory-based inference for multiple regression

::: {.callout-note collapse="true" title="EX10.13 — show question"}
**EX10.13 (★★★)** `pl2` parallel-slopes table.
:::

**Solution** *(reference: Section 10.5 "Theory-based inference for multiple linear regression")*.

```{r 10-ex-13}
m_par <- lm(log_mass ~ log_radius + discovery_method, data = pl2)
get_regression_table(m_par)
```

***

::: {.callout-note collapse="true" title="EX10.14 — show question"}
**EX10.14 (★★★)** Interpret 95% CI for `log_radius`.
:::

**Solution** *(reference: Section 10.5.3 "Confidence intervals", multiple-regression CIs)*. Holding discovery method fixed, a 10× increase in radius is associated with a $\beta$-unit change in log10(mass) — equivalently, a multiplicative change of $10^{\beta}$ in mass. The CI is narrow because $n$ is large.

***

::: {.callout-note collapse="true" title="EX10.15 — show question"}
**EX10.15 (★★★)** Discovery-method p-value tests what?
:::

**Solution** *(reference: Section 10.5.4 "Hypothesis test for a single coefficient")*. $H_0$: the difference between the non-baseline method and the baseline method, *holding radius fixed*, is 0. Equivalently: parallel slopes share a common intercept.

***

::: {.callout-note collapse="true" title="EX10.16 — show question"}
**EX10.16 (★★★)** Interaction p-value.
:::

**Solution** *(reference: Section 10.5.4 "Hypothesis test for a single coefficient", interactions)*. Tests whether the slope on `log_radius` *differs* between the two methods. A small p-value means slopes really do differ — the interaction is needed.

```{r 10-ex-16}
m_int <- lm(log_mass ~ log_radius * discovery_method, data = pl2)
get_regression_table(m_int)
```

***

### Simulation-based inference for multiple regression

::: {.callout-note collapse="true" title="EX10.17 — show question"}
**EX10.17 (★★★)** Bootstrap CI for `log_radius`.
:::

**Solution** *(reference: Section 10.6.1 "Confidence intervals for the partial slopes using infer")*. With multi-predictor models, easiest path is to extract coefficients per bootstrap replicate via `tidy()`.

```{r 10-ex-17}
set.seed(76)
boot_log_radius <- replicate(500, {
  s <- pl2 |> slice_sample(n = nrow(pl2), replace = TRUE)
  coef(lm(log_mass ~ log_radius + discovery_method, data = s))["log_radius"]
})
quantile(boot_log_radius, c(0.025, 0.975))
```

***

::: {.callout-note collapse="true" title="EX10.18 — show question"}
**EX10.18 (★★★)** Bootstrap CI for the **partial slope** on `log_radius` from the parallel-slopes model in EX10.13. Compare to the theory-based CI; under what condition would they disagree?
:::

**Solution** *(reference: Section 10.6 "Simulation-based Inference for multiple linear regression")*.

```{r 10-ex-18, message=FALSE, fig.height=3}
set.seed(76)

# Bootstrap the partial slope on log_radius
n_reps <- 500
boot_slopes <- replicate(n_reps, {
  idx <- sample(seq_len(nrow(pl2)), nrow(pl2), replace = TRUE)
  m_b <- lm(log_mass ~ log_radius + discovery_method, data = pl2[idx, ])
  coef(m_b)["log_radius"]
})

ggplot(data.frame(slope = boot_slopes), aes(x = slope)) +
  geom_histogram(bins = 30) +
  labs(x = "Bootstrap partial slope (log_radius)")

# (b) 95% percentile CI
boot_ci <- quantile(boot_slopes, c(0.025, 0.975))
boot_ci

# Compare with theory-based CI from EX10.14
m_par <- lm(log_mass ~ log_radius + discovery_method, data = pl2)
get_regression_table(m_par)
```

**(b) Comparison.** The bootstrap percentile CI and the theory-based CI on `log_radius` should overlap heavily — both centered on the OLS partial slope, both about the same width.

**(c) When would they disagree?** The theory-based CI assumes the four LINE conditions hold (Linearity, Independence, Normality of residuals, Equal variance). The bootstrap CI relaxes the **Normality** and **Equal-variance** assumptions — it directly resamples the joint $(x, y)$ distribution. So you'd expect them to disagree when:

1. **Residuals are heavy-tailed** (e.g., a few extreme outliers) — the theory CI under-states uncertainty; the bootstrap captures it.
2. **Heteroscedasticity is strong** — variance grows with the predictor; theory's "homoscedastic σ̂" assumption gets the SE wrong.
3. **Sample is small** — the t-distribution approximation degrades; the bootstrap, being non-parametric, fares better.

For *well-behaved* multi-thousand-row datasets (which this is), the two answers are usually within a few percent of each other — the bootstrap is *insurance* for the cases where theory's assumptions wobble.

***

### Visualizing uncertainty

::: {.callout-note collapse="true" title="EX10.19 — show question"}
**EX10.19 (★★★)** Regression line + 95% CI band.
:::

**Solution** *(reference: Section 10.2.6 "Model fit and model assumptions", uncertainty visualization)*.

```{r 10-ex-19, message=FALSE, fig.height=3}
ggplot(bball, aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm")
```

***

::: {.callout-note collapse="true" title="EX10.20 — show question"}
**EX10.20 (★★★)** Confidence band vs prediction band.
:::

**Solution** *(reference: Section 10.2.6 "Model fit and model assumptions", confidence vs prediction)*. **Confidence band**: uncertainty about the *mean* response at a given X — narrow. **Prediction band**: uncertainty about a *single new observation* at that X — much wider, because individual variation gets added to estimation uncertainty.

***

::: {.callout-note collapse="true" title="EX10.21 — show question"}
**EX10.21 (★★★)** Parallel-slopes with CI bands.
:::

**Solution** *(reference: Section 10.5 "Theory-based inference for multiple linear regression", visualizing parallel slopes)*.

```{r 10-ex-21, message=FALSE, fig.height=3}
ggplot(pl2, aes(x = log_radius, y = log_mass, color = discovery_method)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm")
```

***

### Practical questions

::: {.callout-note collapse="true" title="EX10.22 — show question"}
**EX10.22 (★★★)** `Height ~ Year` for basketball.
:::

**Solution** *(reference: Section 10.2.4 "Hypothesis test for population slope")*.

```{r 10-ex-22}
get_regression_table(lm(Height ~ Year, data = bball))
```

A small p-value confirms heights have grown over time in basketball.

***

::: {.callout-note collapse="true" title="EX10.23 — show question"}
**EX10.23 (★★★)** "Olympic basketball *trains* taller players" — refute.
:::

**Solution** *(reference: Section 5.3.1 "Correlation is not necessarily causation")*. Same as Ch 5: the slope reflects **selection** (the Olympics has progressively recruited taller athletes due to NBA pipeline, rule changes), not biology. Inference doesn't establish causation; only experiments and randomization do.

***

::: {.callout-note collapse="true" title="EX10.24 — show question"}
**EX10.24 (★★★)** Repeat for gymnastics; compare.
:::

**Solution** *(reference: Section 10.2.4 "Hypothesis test for population slope")*. Gymnastics likely has a *negative* slope (modern gymnasts are shorter than mid-century ones, especially female).

```{r 10-ex-24}
gym <- olympic_athletes |>
  filter(Sport == "Gymnastics", !is.na(Height))
get_regression_table(lm(Height ~ Year, data = gym))
```

***

::: {.callout-note collapse="true" title="EX10.25 — show question"}
**EX10.25 (★★★)** "The population" in regression — sampling frame.
:::

**Solution** *(reference: Section 10.1 "The simple linear regression model", scope of inference)*. The "population" is implicitly **all Olympic basketball athletes**, past and future, who *could* be in this dataset under the same selection process. Inference applies to that population — *not* to the general human population. This is why CIs from Olympic data tell you about Olympic athletes, not basketball players in general.

***

### Critical thinking and synthesis

::: {.callout-note collapse="true" title="EX10.26 — show question"}
**EX10.26 (★★★)** Tiny p-values vs practical significance.
:::

**Solution** *(reference: Section 10.5.4 "Hypothesis test for a single coefficient", practical vs statistical)*. **No.** With $n \approx 1000$, even effects measured in fractions of a kilogram or millimeters become statistically significant. Always check the **estimate's magnitude** in domain units before celebrating.

***

::: {.callout-note collapse="true" title="EX10.27 — show question"}
**EX10.27 (★★★)** Why removing `Sport` changes `Sex` coefficient.
:::

**Solution** *(reference: Section 10.5.2 "Model dependency of estimators", omitted-variable bias)*. **Sport** is correlated with both Sex (uneven gender ratios across sports) and Height (sport-specific selection on height). Omitting it makes the Sex coefficient absorb that confounding — coefficients are *not invariant* to which controls are included.

***

::: {.callout-note collapse="true" title="EX10.28 — show question"}
**EX10.28 (★★★)** Demonstrate for two specific sports.
:::

**Solution** *(reference: Section 10.5.2 "Model dependency of estimators")*.

```{r 10-ex-28}
two_sports <- olympic_athletes |>
  filter(Sport %in% c("Athletics", "Swimming"),
         !is.na(Height))
get_regression_table(lm(Height ~ Sex, data = two_sports))
get_regression_table(lm(Height ~ Sex + Sport, data = two_sports))
```

The `SexM` coefficient typically *changes* once `Sport` is added — the per-sport male-female gap differs from the pooled gap.

***

::: {.callout-note collapse="true" title="EX10.29 — show question"}
**EX10.29 (★★★)** CI widths at three sample sizes.
:::

**Solution** *(reference: Section 10.2.3 "Confidence intervals for the least-squares estimators")*. Width shrinks as $\sqrt{n}$ grows.

```{r 10-ex-29, fig.height=3}
set.seed(76)
ci_at_n <- function(nn) {
  s <- if (nn == "full") bball else slice_sample(bball, n = as.integer(nn))
  tab <- get_regression_table(lm(Weight ~ Height, data = s))
  tab |> filter(term == "Height") |>
    transmute(lower_ci, upper_ci, n = nn)
}
out <- bind_rows(ci_at_n("30"), ci_at_n("100"), ci_at_n("full"))
out
ggplot(out, aes(y = n, xmin = lower_ci, xmax = upper_ci)) +
  geom_errorbarh(height = 0.2)
```

***

::: {.callout-note collapse="true" title="EX10.30 — show question"}
**EX10.30 (★★★)** R², adjusted R², F-statistic.
:::

**Solution** *(reference: Section 10.2.6 "Model fit and model assumptions", F-test)*. The **F-statistic** tests $H_0$: all slopes are zero (the *whole* model has no explanatory power) vs $H_A$: at least one slope is non-zero. Tiny p-value ⇒ at least one predictor matters.

```{r 10-ex-30}
glance(m_b) |> select(r.squared, adj.r.squared, statistic, p.value, df)
```

***

::: {.callout-note collapse="true" title="EX10.31 — show question"}
**EX10.31 (★★★)** Adjusted R² decreasing on useless predictors.
:::

**Solution** *(reference: Section 10.5.5 "Model fit and diagnostics", adjusted R²)*. Adjusted R² penalizes extra predictors. Plain R² always rises when adding any predictor (even noise). Adjusted R² rises only when the new predictor's contribution exceeds what would be expected from chance — protecting against overfitting.

***

::: {.callout-note collapse="true" title="EX10.32 — show question"}
**EX10.32 (★★★)** Try to demonstrate Simpson's paradox in `planets`.
:::

**Solution** *(reference: Chapter 10 — synthesis, model dependency)*. *Answers vary.* With `mass_earth ~ radius_earth` colored by discovery_method, the *pooled* slope can differ from each *within-method* slope when methods cluster in different parts of the (radius, mass) plane.

```{r 10-ex-32}
pooled <- lm(log_mass ~ log_radius, data = pl2)
within_methods <- pl2 |>
  group_by(discovery_method) |>
  summarize(slope = coef(lm(log_mass ~ log_radius))[2])
list(pooled_slope = coef(pooled)[2], within = within_methods)
```

The slopes can differ; if they have opposite signs and the pooled has the wrong direction, that *would* be Simpson's paradox.

***

::: {.callout-note collapse="true" title="EX10.33 — show question"}
**EX10.33 (★★★)** R² = 0.95 + curvature: trust inference?
:::

**Solution** *(reference: Section 10.2.6 "Model fit and model assumptions")*. **No.** The high R² doesn't validate the *model form*. Curvature in residuals violates linearity, biasing inference (CIs/SEs misstate uncertainty) and producing predictions that systematically err in different parts of the X range.

***

::: {.callout-note collapse="true" title="EX10.34 — show question"}
**EX10.34 (★★★)** CI (0.001, 0.005): tiny but above 0.
:::

**Solution** *(reference: Section 10.2.3 "Confidence intervals for the least-squares estimators", practical significance)*. The slope is *statistically* different from 0 (CI excludes 0), but **practically** tiny — for any X-range you care about, the predicted Y barely moves. Don't write a press release.

***

::: {.callout-note collapse="true" title="EX10.35 — show question"}
**EX10.35 (★★★)** Open exploration.
:::

**Solution** *(reference: Chapter 10 — synthesis)*. *Answers vary.* Example: `Weight ~ Height` for swimming — strong positive slope, narrow 95% CI, p ≈ 0.

```{r 10-ex-35}
swim <- olympic_athletes |>
  filter(Sport == "Swimming", !is.na(Height), !is.na(Weight))
get_regression_table(lm(Weight ~ Height, data = swim))
```

***

::: {.callout-note collapse="true" title="EX10.36 — show question"}
**EX10.36 (★★★)** Synthesis: multiple regression on `eq_temp_k`.
:::

**Solution** *(reference: Chapter 10 — synthesis)*. *Answers vary.* Robustly supported: (i) insolation is a strong, statistically significant predictor of equilibrium temperature; (ii) host star temperature has a smaller but significant additional effect; (iii) selection-bias caveats apply — these conclusions describe the *catalog*, not all exoplanets.

```{r 10-ex-36}
planets_temp <- planets |>
  filter(!is.na(eq_temp_k), !is.na(star_temp_k), !is.na(insolation_earth)) |>
  mutate(log_insol = log10(insolation_earth))
get_regression_table(lm(eq_temp_k ~ log_insol + star_temp_k, data = planets_temp))
```

***
