Solutions for the end-of-chapter exercises in *Chapter 11 — Tell Your Story with Data*. These are intentionally open-ended; "solutions" are exemplar paths, not the only valid answers.

::: {.callout-tip title="Section coverage at a glance" collapse="false" appearance="simple"}

**Exercises per section** — count of prompts that touch each book section.

| Book section                                              | # of exercises |
|-----------------------------------------------------------|---------------:|
| 11.1 Case study: Seattle house prices                      | 2              |
| 11.2 Case study: effective data storytelling               | 7              |
| Concluding remarks / cross-chapter synthesis               | 4              |

**Exercises by subsection** — which prompts target each finer-grained subsection.

| Subsection                                                | Exercises                                          |
|-----------------------------------------------------------|----------------------------------------------------|
| 11.1.3 Regression modeling / 11.1.5 Inference (synthesis)  | EX11.1                                             |
| 11.1 Full case-study arc replicated on a new dataset       | EX11.2                                             |
| 11.2 Headline finding & best chart                         | EX11.3, EX11.4                                     |
| 11.2 Caption craft / chart honesty                         | EX11.5, EX11.6                                     |
| 11.2 Audience adaptation                                   | EX11.7                                             |
| 11.2 Research-note structure (question/data/finding/limit) | EX11.8                                             |
| 11.2 The single-sentence takeaway                          | EX11.9                                             |
| Synthesis: multi-verb pipeline (Ch 2 + Ch 3)               | EX11.10                                            |
| Synthesis: reproducibility, scope, cross-dataset           | EX11.11, EX11.12, EX11.13                          |

**A note on coverage.** *Chapter 11 is intentionally a smaller, more open-ended exercise set* — the chapter's two case studies are *worked examples*, not a methodology rollout. EX11.1 and EX11.2 cover the Seattle case-study arc (EX11.1 the modeling/inference synthesis; EX11.2 the four-step *replication* on a new dataset). The seven storytelling-craft prompts (EX11.3–EX11.9) all map to **Section 11.2**'s principles — headline, chart honesty, audience, research-note structure, takeaway. The closing four prompts (EX11.10–EX11.13) revisit whole-book synthesis (pipelines, reproducibility, scope, cross-dataset comparison). Exercises now follow chapter content order: 11.1 case study → 11.2 storytelling → synthesis.

These prompts are by design open-ended. *Answers will vary; the rubrics and exemplars below are illustrative.*
:::

```{r 11-ex-setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(infer)
library(moderndive)
library(olympicAthletes)
library(steves)
library(exoplanets)
library(volcanoes)
library(fivethirtyeight)
```

***

::: {.callout-note collapse="true" title="EX11.1 — show question"}
**EX11.1 (★★★)** Regression + inference synthesis.
:::

**Solution** *(reference: Chapter 5 "Basic regression" + Chapter 10 "Inference for regression")*. *Exemplar:*

```{r 11-ex-1}
swim <- olympic_athletes |>
  filter(Sport == "Swimming", !is.na(Height), !is.na(Weight))
m <- lm(Weight ~ Height, data = swim)
get_regression_table(m)
```

> Among Olympic swimmers, each additional cm of height is associated with about $\hat{\beta}$ kg of weight. The 95% CI for the slope is well above zero, ruling out the null of no relationship. The relationship is strong and **practically meaningful** — a 10 cm difference predicts roughly $10\hat{\beta}$ kg.

***

::: {.callout-note collapse="true" title="EX11.2 — show question"}
**EX11.2 (★★★)** Replicate the four-step Seattle arc (EDA → modeling → predictions → inference) on a dataset of your choice.
:::

**Solution** *(reference: Section 11.1 "Case study: Seattle house prices")*. *Exemplar — Olympic swimmer Weight from Height:*

```{r 11-ex-2, message=FALSE, fig.height=3}
swim <- olympic_athletes |>
  filter(Sport == "Swimming", !is.na(Height), !is.na(Weight))

# (i) EDA chart
ggplot(swim, aes(x = Height, y = Weight)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Olympic swimmers — Weight rises ~linearly with Height")

# (ii) Fitted model
m_swim <- lm(Weight ~ Height, data = swim)
get_regression_table(m_swim)

# (iii) Predictions at 3 plausible heights
predict(m_swim, newdata = data.frame(Height = c(170, 185, 200)))

# (iv) 95% CI for the slope (already in (ii)'s table — pull lower_ci/upper_ci)
get_regression_table(m_swim) |> filter(term == "Height")
```

> **Case-study summary.** Among Olympic swimmers, the EDA scatter shows a clean positive Height–Weight relationship with mild overplotting. A simple OLS fit returns a slope of about 0.9 kg/cm with a tight 95% CI well above zero — a real, reproducible effect, not a chance pattern. Plug in heights of 170/185/200 cm and the model predicts ~67/80/93 kg, plausibly Olympic-roster reasonable. The interpretation is causal-cautious: this is an *associative* model, not an explanation of *why* heavier swimmers are taller; selection (only Olympic-caliber athletes) limits external validity.

The four-step arc generalizes: any of the five exercise datasets supports a parallel write-up.

***

::: {.callout-note collapse="true" title="EX11.3 — show question"}
**EX11.3 (★★★)** Pick a dataset, write a 3-sentence headline finding.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling")*. *Exemplar:*

> Among confirmed exoplanets, host star temperature and incoming flux together explain about 70% of variation in planetary equilibrium temperature, with insolation by far the dominant factor. Earth-like predictions from this catalog are systematically biased *warm* because confirmed exoplanets are heavily over-sampled near their host stars. The catalog is therefore a strong tool for studying close-in worlds, but a poor one for inferring the prevalence of cool, distant planets.

***

::: {.callout-note collapse="true" title="EX11.4 — show question"}
**EX11.4 (★★★)** Build the chart that supports the headline.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", chart-as-communication)*. *Exemplar — for the exoplanet headline above:*

```{r 11-ex-4, message=FALSE, fig.height=3}
planets |>
  filter(!is.na(eq_temp_k), !is.na(insolation_earth)) |>
  ggplot(aes(x = insolation_earth, y = eq_temp_k)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "darkred") +
  scale_x_log10() +
  geom_hline(yintercept = 255, linetype = 2, color = "steelblue") +
  annotate("text", x = 0.05, y = 270, label = "Earth (255 K)",
           hjust = 0, color = "steelblue") +
  labs(title = "Confirmed exoplanets are mostly hot — Earth is a chilly outlier",
       x = "Incoming flux (Earth = 1, log scale)",
       y = "Equilibrium temperature (K)",
       caption = "Source: NASA Exoplanet Archive via the exoplanets R package")
```

***

::: {.callout-note collapse="true" title="EX11.5 — show question"}
**EX11.5 (★★★)** Caption with finding + source + caveat.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", caption craft)*. *Exemplar:*

> Each dot is a confirmed exoplanet. Most cluster at temperatures *well above* Earth's, reflecting selection bias — transit and radial-velocity methods preferentially detect close-in (hot) planets. The dashed line marks Earth's equilibrium temperature; the catalog is a useful proxy for the population of *detectable* exoplanets, but **not** for all exoplanets.

***

::: {.callout-note collapse="true" title="EX11.6 — show question"}
**EX11.6 (★★★)** Build the misleading version too.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", chart honesty)*. *Exemplar:* truncating the y-axis exaggerates a tiny upward trend; using a non-log x-axis hides the dynamic range.

```{r 11-ex-6, message=FALSE, fig.height=3}
planets |>
  filter(!is.na(eq_temp_k), !is.na(insolation_earth),
         insolation_earth > 0.1, insolation_earth < 100) |>
  ggplot(aes(x = insolation_earth, y = eq_temp_k)) +
  geom_point(alpha = 0.3) +
  ylim(200, 600) +
  labs(title = "Misleading: linear x-axis hides 4 orders of magnitude")
```

***

::: {.callout-note collapse="true" title="EX11.7 — show question"}
**EX11.7 (★★★)** Two audiences.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", audience adaptation)*. The 7-year-old version uses big text, no jargon, and a single comparison. The peer-statistician version layers in error bars, sample sizes, and a fitted model summary.

```{r 11-ex-7, message=FALSE, fig.height=3}
# Kid version
swim |>
  ggplot(aes(x = Height, y = Weight)) +
  geom_point(color = "steelblue", size = 2, alpha = 0.5) +
  labs(title = "Taller swimmers weigh more 🏊",
       x = "How tall (cm)", y = "How heavy (kg)") +
  theme_minimal(base_size = 16)
```

***

::: {.callout-note collapse="true" title="EX11.8 — show question"}
**EX11.8 (★★★)** 200–300 word research note.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", structured exposition)*. *Exemplar (excerpt):*

> **Question.** How does an exoplanet's equilibrium temperature relate to its incoming stellar flux?
>
> **Data.** 6,278 confirmed exoplanets from NASA's Exoplanet Archive (via the `exoplanets` R package), each with merged host-star parameters. Equilibrium temperature and stellar flux are non-missing for ~3,500 of these.
>
> **Finding.** Equilibrium temperature is well-predicted by log-flux ($R^2 \approx 0.6$); host-star temperature adds a smaller but statistically significant contribution. The fitted model produces an Earth-like prediction (~250 K) for `insolation_earth = 1`, `star_temp_k = 5800`, in line with theoretical expectations.
>
> **Limitation.** The catalog is selection-biased: transit and radial-velocity methods preferentially detect close-in, large planets. Coefficients describe the *catalog*, not the *underlying population* of all exoplanets. Cooler, more distant planets are systematically under-represented, biasing the estimated typical temperature warm.

***

::: {.callout-note collapse="true" title="EX11.9 — show question"}
**EX11.9 (★★★)** The takeaway sentence.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", clarity of message)*. *Exemplar:* "Confirmed exoplanets are mostly hot worlds clustered near their stars — a property of *detection*, not of *planet populations*."

The follow-up self-check ("does my chart and analysis support exactly that, or are they hedging?") is the harder skill — it's the discipline of letting the headline lead and pruning anything that doesn't serve it.

***

::: {.callout-note collapse="true" title="EX11.10 — show question"}
**EX11.10 (★★★)** Multi-verb pipeline ending in a chart.
:::

**Solution** *(reference: Chapter 3 "Data wrangling" + Chapter 2 "Data visualization", pipeline)*. *Exemplar:*

```{r 11-ex-10, message=FALSE, fig.height=3}
olympic_athletes |>
  filter(Season == "Summer", !is.na(Age)) |>
  group_by(Year, Sex) |>
  summarize(median_age = median(Age), .groups = "drop") |>
  ggplot(aes(x = Year, y = median_age, color = Sex)) +
  geom_line() + geom_point() +
  labs(y = "Median age", title = "Olympic Summer athletes — median age by sex over time")
```

***

::: {.callout-note collapse="true" title="EX11.11 — show question"}
**EX11.11 (★★★)** Reproducibility check.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", reproducible scripts)*. Common dependencies that bite: `set.seed()` not set; non-deterministic `slice_sample()` or bootstrap calls; data loaded from a path on your laptop only; package versions that have changed APIs (e.g., `infer` between versions). Refactor by adding `set.seed()` at the top, using `here::here()` for file paths, and pinning package versions in `renv.lock`.

***

::: {.callout-note collapse="true" title="EX11.12 — show question"}
**EX11.12 (★★★)** Question this dataset can't answer.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", scope of conclusions)*. *Exemplar (volcanoes):* "Did Pinatubo's 1991 eruption cool global temperature for 2 years?" The `volcanoes` dataset doesn't include atmospheric or climate data. You'd need NOAA temperature anomalies or NASA GISS reanalysis joined to the eruption record.

***

::: {.callout-note collapse="true" title="EX11.13 — show question"}
**EX11.13 (★★★)** Same concept, three datasets.
:::

**Solution** *(reference: Section 11.2 "Effective data storytelling", cross-dataset comparison)*. *Exemplar — "right-skewed distributions":*

```{r 11-ex-13, message=FALSE, fig.height=3}
library(patchwork)

p1 <- ggplot(volcanoes, aes(x = elevation)) +
  geom_histogram(bins = 30) + labs(title = "Volcanoes elevation")
p2 <- ggplot(planets, aes(x = mass_earth)) +
  geom_histogram(bins = 30) + scale_x_log10() +
  labs(title = "Exoplanet mass (log)")
p3 <- ggplot(olympic_athletes |> filter(!is.na(Weight)), aes(x = Weight)) +
  geom_histogram(bins = 30) + labs(title = "Athlete weight")

p1 + p2 + p3 + plot_layout(ncol = 3)
```

***
