Solutions for the end-of-chapter exercises in *Chapter 2: Data Visualization*. The dataset is `olympic_athletes` (≈`r round(nrow(olympicAthletes::olympic_athletes) / 1000)`k athlete-event rows, `r min(olympicAthletes::olympic_athletes$year)`–`r max(olympicAthletes::olympic_athletes$year)`), accompanied by `medal_table` (medal counts per Games × NOC for `r min(olympicAthletes::medal_table$year)`–`r max(olympicAthletes::medal_table$year)`) and `editions` (per-Games metadata for all `r nrow(olympicAthletes::editions)` editions, `r min(olympicAthletes::editions$year)`–`r max(olympicAthletes::editions$year)`), all from the [`olympicAthletes`](https://github.com/moderndive/olympicAthletes) package.

```{r}
#| label: ch2-coverage
#| results: asis
#| echo: false
#| message: false
helpers <- Find(file.exists, c("scripts/exercise_helpers.R", "../scripts/exercise_helpers.R", "../../scripts/exercise_helpers.R"))
source(helpers)
cat(render_coverage_callout(2))
```

```{r}
#| label: 02-ex-setup
#| message: false
library(dplyr)
library(ggplot2)
library(olympicAthletes)
```

```{r}
#| label: ch2-solutions
#| results: asis
#| echo: false
#| message: false
cat(render_solutions(2))
```
