Solutions for the end-of-chapter exercises in *Chapter 4 — Data Importing and Tidy Data*. The dataset is `bob_ross` from the [`fivethirtyeight`](https://fivethirtyeight-r.netlify.app/reference/bob_ross.html) package: `r {n_bin <- sum(vapply(fivethirtyeight::bob_ross, \(x) all(x %in% c(0, 1, NA)), logical(1))); nrow(fivethirtyeight::bob_ross)}` episodes × `r ncol(fivethirtyeight::bob_ross)` columns (`r ncol(fivethirtyeight::bob_ross) - n_bin` identifiers + `r n_bin` binary indicator columns).

```{r}
#| label: ch4-coverage
#| results: asis
#| echo: false
#| message: false
helpers <- Find(file.exists, c("scripts/exercise_helpers.R", "../scripts/exercise_helpers.R", "../../scripts/exercise_helpers.R"))
source(helpers)
cat(render_coverage_callout(4))
```

```{r}
#| label: 04-ex-setup
#| message: false
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(fivethirtyeight)
```

```{r}
#| label: ch4-solutions
#| results: asis
#| echo: false
#| message: false
cat(render_solutions(4))
```
